### 2011-02-04 MScDiplom_Lemenkova.tex
added 'Index of People'

\documentclass[11pt, a4paper]{article}
\usepackage{authblk}
%\usepackage{authorindex}
\usepackage{color}
\usepackage[stable]{footmisc }
\usepackage{setspace}
\usepackage[printonlyused,withpage]{acronym}
\usepackage[a4paper, left=30mm,right=25mm,top=30mm,bottom=35mm]{geometry}
%\usepackage{glossary}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage[numbers]{natbib}
\usepackage{calc}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage[svgnames,table]{xcolor}
\usepackage[subfigure]{tocloft}
%\usepackage[table]{xcolor}
\usepackage{paralist}
%\usepackage{makeidx}\makeindex
%\usepackage[makeindex,protected]{splitidx}
%\usepackage{index}
\usepackage{imakeidx}
\makeindex[name=concepts,title=Index of Concepts,intoc]
\makeindex[name=people,title=Index of People,intoc,columns=3,columnseprule]
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{fancyheadings}
\topmargin = 1pt
\headheight=10mm
\headsep=7pt
%\ensuremath
\usepackage{lastpage}
\usepackage{float}
\usepackage{subfig}
\usepackage{caption}
\usepackage{chngcntr}
\counterwithin{figure}{section}
\counterwithin{table}{section}
\usepackage{appendix}
\captionsetup{margin=10pt,font=small,labelfont=bf}
\usepackage{wrapfig}
\definecolor{darkblue}{rgb}{0.0,0.0,0.3}
\setcounter{tocdepth}{4}
\parindent 0pt
%\floatstyle{boxed} 
\restylefloat{figure}
\usepackage[a4paper=true, ]{hyperref}
\hypersetup{pdftitle={Seagrass mapping and monitoring along the coasts of Crete, Greece}, pdfauthor={Lemenkova Polina}, pdfsubject={MScThesis}, pdfcreator={Lemenkova Polina}, pdfproducer={Lemenkova Polina}, colorlinks=true,linkcolor=Navy, citecolor=Crimson, filecolor=magenta, urlcolor=blue}

%%%%%%%%%%%%%%%%%%%%%%%% End Document Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% Begin Lemenkova Polina's MSc Thesis Document%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\pagenumbering{roman}

%\newcommand{\listpeoplename}{}
%\newlistof{people}{peo}{\listpeoplename}
%\newcommand{\people}[1]{\refstepcounter{people}}

%\newindex[List of People]{person}
%\setindexpreamble[person]{...}
%\newcommand{\person}[2]{\sindex[person]{#2, #1}#2}% "Given Surname" in text

\phantomsection	
\addcontentsline{toc}{section}{Front Page}
\begin{titlepage}
	\centering
	\title{Seagrass mapping and monitoring \\ along the coasts of Crete, Greece}
	\author[1]{by Lemenkova Polina}
	\affil[1]{MSc student at the \textit{University of Twente, Faculty of Geo-Information Science and Earth Observation (ITC)}, The Netherlands}
	\date{\today}
	\maketitle	
\end{titlepage}
\pagebreak

\section*{}
\addcontentsline{toc}{section}{Course Title}
\textbf{Course Title} \\
{Geo-Information Science and Earth Observation for Environmental Modelling and Management}\\

\textbf{Level} \hfill{Master of Science (MSc)}\\

\vspace{1em}

\textbf{Course Duration:} \hfill{September 2009 – March 2011}\\

\vspace{1em}

\textbf{Consortium partners:} \\

\begin{itemize}
	\item[*] \href{http://www.southampton.ac.uk/}{University of Southampton}, \href{http://www.southampton.ac.uk/geography/}{School of Geography} (UK)\\
	\item[*] \href{http://www.lunduniversity.lu.se/}{Lund University}, \href{http://www.giscentrum.lu.se/english/index.htm}{GIS Centre} (Sweden)\\
	\item[*] \href{http://www.uw.edu.pl/en/}{University of Warsaw} (Poland)\\
	\item[*] \href{http://www.utwente.nl/en}{University of Twente}, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)} (The Netherlands)
\end{itemize}

\begin{figure}[H]
	\centering
	\subfloat{\includegraphics [width=0.2\textwidth] {Logo-Sou.jpg}}
	\hspace{3mm}
	\subfloat{\includegraphics [width=0.1\textwidth] {Logo-Lun.jpg}}
	\hspace{3mm}
	\subfloat{\includegraphics [width=0.1\textwidth] {Logo-War.jpg}}
	\hspace{3mm}
	\subfloat{\includegraphics [width=0.5\textwidth] {ITC-logo.jpg}}	
\end{figure}

\pagebreak

\section*{}
\addcontentsline{toc}{section}{Supervisory Committee}
\begin{center}
\textbf{Seagrass mapping and monitoring along the coasts of Crete, Greece}\\
\vspace{1em}
by Lemenkova Polina\index[people]{Lemenkova, P.}
\end{center}

Thesis submitted to the University of Twente, Faculty ITC, in partial fulfillment of the requirements
for the degree of Master of Science in Geo-information Science and Earth Observation for
Environmental Modelling and Management.\\

\textbf{Thesis Assessment Board}:
\begin{itemize}
	\item Name Examiner 1 (Chair)\dotfill
	\item Name Examiner 2\dotfill
	\item External Examiner \dotfill \\ \hfill{Dr.{\textit{Petter Pilesj\"o}\index[people]{Pilesj\"o, P.}} (Lund University, GIS Centre)}
\end{itemize}

\vspace{2em}

\textbf{Member Supervisors}: \\
\begin{itemize}
	\item[* ] Mr. \textit{Valentijn Venus}\index[people]{Venus, V.} \dotfill \\ (Department of Natural Resource, ITC)
	\item[* ] Dr. \textit{Albertus G. Toxopeus}\index[people]{Toxopeus, A.G.} \dotfill \\ (Department of Natural Resource, ITC)
\end{itemize}

\begin{figure}[h]
	\begin{center}
	\includegraphics[height=2cm,width=8cm]{ITC-logo.jpg}
	\end{center}
\end{figure}
\pagebreak

\section*{}
	\markright{Table of Content}
	\addcontentsline{toc}{section}{Table of Content}
	%\setlength{\textwidth}{6in}
%	\addtolength{\voffset}{+5pt}
%	\marginparwidth = 30pt
%	\textheight = 630pt
	\tableofcontents 
\pagebreak

\section*{Acknowledgements}
	\markright{Acknowledgements}
	\addcontentsline{toc}{section}{Acknowledgements}

This research was carried out at the University of Twente, \href{http://www.itc.nl/}{Faculty of Geo-Information Science and Earth Observation (ITC)}, Enschede, the Netherlands, under the financial support of the EU \href{http://ec.europa.eu/education/external-relation-programmes/doc72_en.htm}{Erasmus Mundus Scholarship}. 
\vspace{1mm}\\
I thank my scientific supervisors, Dr. \textit{Albertus G.Toxopeus}\index[people]{Toxopeus, A.G.}  and Mr. \textit{Valentijn Venus}\index[people]{Venus, V.} from the ITC, Department of Natural Resources, for our successful scientific collaboration, help and discussions of my research work. \\
I also express them my most sincere appreciations for our fieldwork on Crete. \\ Besides scientific routine, we had an unforgettable time spent together. \\ Personally, I would like to mention that I have visited Greece the first time in my life, and I will always remember my Cretan period-2010 and Ligaria beach.\vspace{1mm}\\
I extend my gratitude to the staff of the Museum of Natural History, the University of Crete, especially to Dr.\textit{Petros Limberakis}\index[people]{Limberakis, Petros} and his colleagues, for their assistance during our fieldwork, providing us with all necessary equipment (e.g. SCUBA gear diving, instruments for underwater measurements, etc). I also appreciate their very warm, personal attitude and interest towards our research. 
\vspace{1mm}\\
My sincere gratitude goes again to Mr.\textit{Valentijn Venus}\index[people]{Venus, V.} for his generous providing me with a lot of spatial data   relevant to my research (e.g. such as digital vector layers of Crete island, raster georeferenced maps for the whole island for my GIS project, Excel spreadsheets, etc). Without these vital data the implementation of this project would have hardly been possible. I am also thankful to him for the technical help in Google Earth imagery capturing, using GDAL technologies and assistance during the fieldwork (videometric underwater footage, consultancies in seagrass observation and monitoring, driving boat, etc). \\ My special thank to his personal, friendly support in various situations on Crete.\vspace{1mm}\\
I am thankful to Dr.rer.nat.\textit{Peter Gege}\index[people]{Gege P.}, the software developer from the Earth Observation Centre of the German Space Agency, for his generous consultancies concerning the technical question about WASI that arose during my work. I am thankful for all his explanations and articles given to me via e-mail \vspace{1mm}\\
My special thanks are addressed to my elder brother, Mr.\textit{Peter Lemenkov}\index[people]{Lemenkov, P.}, the software developer from the ipPort, for his priceless numerous advices and constant professional support throughout my career. He encouraged me to use various very useful software and tools, always gave me necessary recommendations and precious help, thus highly contributing to my professional growth. \\ My special thank for his help with Python scripting. \\
I am also heartfelt thankful to him for his personal friendship and support. 
\vspace{1mm}\\
I am grateful to Mr.\textit{Willem Nieuwenhuis}\index[people]{Nieuwenhuis, W.}, the software developer from the ITC, Department of Natural Resources, for shared technology of Google Earth aerial imagery grabbing and for his consultancy with ASTER imagery. 
\vspace{1mm}\\
I thank Dr. Ir. \textit{Cornelis A.J.M. de Bie}\index[people]{de Bie C.A.J.M.} from the Department of Natural Resource, ITC, for his patient sharing of iPAQ knowledge during my Polish fieldwork in summer 2010. I used this skills on Crete. 
\vspace{1mm}\\
Many thanks for Dr. \textit{Dingtian Yang}\index[people]{Yang D.} from the South China Sea Institute of Oceanology, our colleague in problem of seagrass monitoring and research, for his handing over the data of spectral reflectance of Thalassia seagrass, which enabled to compare spectra of various seagrass types.

\pagebreak

\section*{Disclaimer}
	\markright{Disclaimer}
	\addcontentsline{toc}{section}{Disclaimer}
		This document describes work undertaken as part of a programme of study at the University of
		Twente, Faculty ITC. All views and opinions expressed therein remain the sole responsibility of
		the author, and do not necessarily represent those of the university.\\

\vspace{3em}

		I herewith certify that although I conferred with other researchers while preparing this document, and used a range of various research and literature 			resources, cited in the text, the content of this MSc Thesis document remains my own, original work, written by me personally. \\

\vspace{1em}

		Signed................................(Lemenkova Polina) \\

\vspace{1em}

		Enschede, The Netherlands. \\ 18th February 2011.
\pagebreak

\section*{Abstract}
	\markright{Summary Abstract of the MSc Thesis}
	\addcontentsline{toc}{section}{Abstract}

The seagrasses, a unique group of aquatic plants, create complex, extremely diversified and productive ecological systems in the littoral coastal zones. The only flowering plant in the world that is able to live completely submerged, seagrasses play vital role in the marine ecosystems of the World Ocean. Seagrasses are the most important component in the environmental food chain of the coastal ecosystems, being a vital food source for various marine species (e.g. fish, dugongs, turtles, swans), and a producer of organic matter, which is the very basis of the food web.\\
The \textit{P.oceanica} seagrass is an endemic for the Mediterranean region, and a main species in the marine coastal environment of Greece. Meadows of \textit{P.oceanica} are subjected to the human activities, because they occur in coastal areas, where they are affected both by anthropogenic and by climatic and environmental factors. Nowadays \textit{P.oceanica} is in the alarming state of regression, because of the deterioration of the environment in the Mediterranean Sea. Due to these reasons, \textit{P.oceanica} is a protected species since 1988 in some European countries (France). Monitoring \textit{P.oceanica} is therefore an important contribution to the saving and protecting the environment of Mediterranean region.\\
The current MSc thesis focuses on the monitoring of seagrass \textit{P.oceanica} along the northern coasts of Crete Island, Greece, and investigates the application of the remote sensing techniques for the seagrass mapping. 
This research was articulated in two parts, where the first one involves an ecological approach to the seagrass distribution in various regions around the globe and the experience of seagrass monitoring nowadays. The second part of this work has technical character and investigates the application of the remote sensing techniques towards seagrass mapping. It, furthermore, focuses on the optical properties of the \textit{P.oceanica} and other seafloor cover types, and studies distinguishability of various seafloor cover types. Studies of the optical characteristics of separate seafloor cover types were made with purpose to clarify, whether their spectral properties change with varying environmental conditions. \\
Special attention has been drawn on the role of environmental factors on the distribution of \textit{P.oceanica} along the coasts of Crete, and in particular, how the optical properties of the seafloor cover types, i.e. spectral reflectance, are being changed under varying external conditions, e.g. water column, amount of suspended particles and sediments in the seawater, and water temperature. For this purpose we studied differences in the spectral reflectance of \textit{P.oceanica} and other bottom cover types at three distinct depths. The diverse spectral values entail variations in optical properties of the seafloor cover types at changing environmental conditions. We applied \ac{WASI} simulation techniques for the modelling of the optical parameters of various seafloor cover types by various spaceborne imaging spectrometers (\ac{MERIS}, \ac{SeaWiFS}, \ac{CZCS} and \ac{MODIS}), in order to understand their suitability and possible limitations for the seagrass mapping.\\
Fieldwork research sites were presented by separate locations on the northern coast of Crete region (Ligaria, Agia Pelagia, Xerocampos). The additional measurements of the reflectance spectra of the seawater with and without sediments have been made in aquarium tank in 2009 by means of Trios-RAMSES spectroradiometer. Parallel to the collection of spectra signatures, we captured the imagery for the seagrass mapping, which consists of the aerial images from the Google Earth website and the satellite Landsat TM and ETM+ scenes.

\pagebreak

\section*{Document Outline}
	\markright{Structure of the MSc Thesis}
	\addcontentsline{toc}{section}{Document Outline}
\textit{Chapter 1}: \nameref{sec:1} is an introductory section. It describes general background of the research problem, outlines the need and actuality of the problem (including the environmental vulnerability of the seagrasses), and highlights limitations and possibilities of the remote sensing application for the current work. In this chapter we also set up research objectives, put research questions for hypothesis testing and sketch research approach for the proposed work.
\newline
In \textit{Chapter 2}: \nameref{sec:2} we review the existing literature and reported research experience on the similar problem: studies of spectral separability of various seafloor types, limitations and advantages of the remote sensing techniques applied for seagrass mapping seagrass ecology and mapping seagrass environment. We considered not only the Mediterranean environment, but also papers from Australian and Chinese scientists, because seagrass monitoring is most actively developed in the southern regions of our planet. The review of the existing RTM algorithms for the retrieval of the optical parameters, as well as description of various tools for the spectrometric measurements - are given in the same Chapter as well. More close attention has been given to the imaging hyperspectral radiometer Trios RAMSES, used for the data collection.\\
\textit{Chapter 3}: \nameref{sec:3} deals with the methods and materials used in current work. We start our discussion from the fieldwork area location and describe tools and instruments used during the fieldwork, as well as sampling design. Procedures of the data pre-processing and capturing imagery from the Google Earth are also presented in the same chapter. \\
\textit{Chapter 4}: \nameref{sec:4} presents to the reader the main results of the current MSc work, obtained during the processing fieldwork and other collected data. It starts from the review of the collected data from different sources, then describes the modelling component of the research, namely \ac{WASI} colour simulation and plotted resulting graphs. We also discuss here the particularities of various sensors, suitable enough for deriving radiometric information to discriminate sand from the seagrass. Furthermore, this chapter includes analysis of the spectral signatures of various seafloor cover types in conditions of changing environment, possibility and limitations for \textit{P.oceanica} spectral discrimination, statistical analysis of the data sets and mapping based on images processing.\\
\textit{Chapter 5}: \nameref{sec:5} summarizes and briefly discusses again the main results, reported in detail in a previous chapter. This chapter proposes some final discussions to the reader, and comes to the main research outcomes.\\
Finally, we suggest some recommendations for any further research focusing on the seagrass environment along the coasts of Crete (or other Mediterranean areas) in the final \nameref{sec:6}.\\
A lot of plotted graphs, statistical outcomes, tables, auxiliary (yet relevant to our research) pictures and images, illustrating our work - are collected in the \nameref{app}. Due to the standard editorial limitations of the current MSc work, it was not possible to include all them to the main chapters. However, we tried to make the structure of the Appendices most clear and easy-readable as possible, by dividing the Appendices into 10 various sub-sections and referencing to them from the main text where necessary. \\
The \nameref{bib} section includes literature and internet resources used for the current work. The cross-referencing and web hyper-referencing are used in the whole document to make reading more quick, effective and informative.\\
An Index is placed in the end of the Document to help the reader find what he is looking for.

\pagebreak

\section*{List of Acronyms\footnote{The page is given where the glossary is first entered and defined}}
	\markright{Acronyms}
	\addcontentsline{toc}{section}{List of Acronyms}
%\singlespacing
\begin{acronym}
	\acro{4AOP}{\href{http://www.noveltis.fr/4AOP/interface/index.php}{Automatized Atmospheric Absorption Atlas Operational Release}} 
	\acro{6SV1}{Second Simulation of a Satellite Signal in the Solar Spectrum Version1} 
	\acro{ASCII}{\href{http://en.wikipedia.org/wiki/ASCII}{American Standard Code for Information Interchange}}
	\acro{ANOVA}{\href{http://en.wikipedia.org/wiki/Anova}{ANalysis of VAriance}} 
	\acro{ASTER}{\href{http://asterweb.jpl.nasa.gov/}{Advanced Spaceborne Thermal Emission and Reflection Radiometer}} 
	%\acro{BRDF}{Bidirectional Reflectance Distribution Function}
	\acro{CASI}{\href{http://www.borstad.com/casi.html}{Compact Airborne Spectrographic Imager}}
	\acro{CNES}{\href{http://www.cnes.fr/}{Centre National d'Etudes Spatiales (of France)}}
	\acro{CRTM}{\href{http://www.jcsda.noaa.gov/projects_crtm.php}{Community Radiative Transfer Model}}
	\acro{CZCS}{\href{http://oceancolor.gsfc.nasa.gov/CZCS/}{Coastal Zone Color Scanner}}
	\acro{DISORT}{\href{ftp://climate1.gsfc.nasa.gov/wiscombe/Multiple_Scatt/DISORT1.2/}{Discrete Ordinates Radiative Transfer Program for a Multi-Layered Plane-Parallel Medium}}
	\acro{ENVISAT}{\href{http://envisat.esa.int/instruments/meris/}{ENVIronmental SATellite}}
	\acro{ESA}{\href{http://www.esa.int/}{European Space Agency}}
	%\acro{ERS}{European Remote Sensing satellite}
	\acro{GIS}{Geographic Information System}
	%\acro{GNU}{GNU's Not Unix}
	\acro{GloVis}{\href{http://glovis.usgs.gov/}{Global Visualization Viewer}}
	\acro{GPS}{Global Positioning System}
	\acro{GRASS}{\href{http://grass.fbk.eu/}{Geographic Resources Analysis Support System}}
	%\acro{GRETL}{GNU Regression, Econometrics and Time-series Library}
	\acro{HyMap}{\href{http://www.hyvista.com/}{Hyperspectral Mapper}}
	\acro{HyperOCR}{\href{http://www.satlantic.com/Hyperspectral_Radiometers/Hyperspectral_Radiometers.asp}{Hyperspectral Ocean Colour Radiometer}}
	\acro{ILWIS}{\href{http://www.ilwis.org/}{Integrated Land and Water Information System}}
	\acro{iPAQ}{\href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{Internet Pocket PC by CompAQ}}
	\acro{KOPRA}{\href{http://www.imk-asf.kit.edu/english/312.php}{Karlsruhe Optimized and Precise Radiative transfer Algorithm}}
	\acro{LAI}{\href{http://en.wikipedia.org/wiki/Leaf_Area_Index}{Leaf Area Index}}
	\acro{Landsat TM}{\href{http://landsat.gsfc.nasa.gov/}{Landsat Thematic Mapper}}
	\acro{Landsat ETM+}{\href{http://landsat.gsfc.nasa.gov/}{Landsat Enhanced Thematic Mapper Plus}}
	\acro{MERIS}{\href{http://envisat.esa.int/instruments/meris/}{MEdium Resolution Imaging Spectrometer}}
	\acro{MIPAS}{\href{http://envisat.esa.int/instruments/mipas/}{Michelson Interferometer for Passive Atmospheric Sounding}}
	\acro{MODIS}{\href{http://modis.gsfc.nasa.gov/}{Moderate Resolution Imaging Spectroradiometer}}
	\acro{NASA}{\href{http://www.nasa.gov/}{National Aeronautics and Space Administration}}
	\acro{NIR}{Near-Infra-Red}
	\acro{NOAA}{\href{http://www.osdpd.noaa.gov/ml/index.html}{National Oceanic and Atmospheric Administration}}
	\acro{OBC}{On Board Calibration}
	\acro{Olympus ST}{\href{http://www.olympusamerica.com/}{Olympus Stylus Tough}}
	\acro{RRTM}{\href{http://rtweb.aer.com/}{Rapid Radiative Transfer Model}}
	\acro{RTTOV-9}{\href{http://research.metoffice.gov.uk/research/interproj/nwpsaf/rtm/}{Radiative Transfer for TIROS (Television and InfraRed Observation Satellite) Operational Vertical Sounder}}
	\acro{RTM}{Radiative Transfer Model}
	\acro{SAM}{Scanning Acoustic Microscope}
	\acro{SCUBA}{\href{http://www.padi.com/scuba/}{Self Contained Underwater Breathing Apparatus}}
	\acro{SeaWiFS}{\href{http://oceancolor.gsfc.nasa.gov/SeaWiFS/}{Sea-viewing Wide Field-of-view Sensor}}
	\acro{SPOT}{\href{http://www.spotimage.fr/}{Satellite Pour l'Observation de la Terre}}
	\acro{SPSS}{ \href{http://www.spss.com/}{Statistical Package for the Social Sciences}}
	\acro{STDEV}{Standard Deviation}
	\acro{SWIR}{Short Wave InfraRed}
	\acro{TIR}{Thermal InfraRed}
	\acro{Trios-RAMSES}{\href{http://trios.de/index.php?option=com_content&view=category&layout=trios&id=47&Itemid=76&lang=en}{RAdiance Measurement SEnSor}}
	\acro{UML}{Unified Modelling Language}
	\acro{USGS}{\href{http://www.usgs.gov/}{U.S. Geological Survey}}
	\acro{UV}{UltraViolet (10 nm to 400 nm)}
	\acro{UVA}{UltraViolet A, or long waves (315 nm-400 nm)}
	\acro{UVB}{UltraViolet B, or medium waves (280 nm-315 nm)}
	\acro{VIS}{VISible}
	\acro{VNIR}{Visible and Near-InfraRed}
	\acro{WASI}{\href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}}
\end{acronym}
\pagebreak

\singlespace
\section*{List of Tools}\label{sec:0}
	\markright{List of Tools}
	\addcontentsline{toc}{section}{List of Tools}\index[concepts]{Software used}

	\vspace{2em}

	\textbf{Software and utilities used in the current work}:\\

	\vspace{1ex}
	\textbf{Main Tools}:
	\begin{itemize}
		\item \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0}: mapping, data integration and spatial analysis
		\item \href{http://www.ilwis.org/}{ILWIS}, \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: imagery processing and classification
		\item \href{http://www.filewatcher.com/b/ftp/ftp.dfd.dlr.de/pub/WASI.0.0.html}{WASI Water Color Simulator}, for modelling optical water 					properties under changing environment
		\item \href{http://welcome.hp.com/country/us/en/prodserv/handheld.html}{iPAQ} and \ac{GPS} for data capture during the fieldwork
		\item \href{http://www.bibtex.org/de/}{BibTeX}, for management of bibliography and reference database
		\item \href{http://www.latex-project.org/}{\LaTeX}, for document typesetting, formatting and performing pdf output
		\item \href{http://www.python.org/}{Python language}, for writing auxiliary scripts (e.g. raw data interpolation)
		\item \href{http://www.gnuplot.info/}{Gnuplot} for plotting graphs
		\item \href{http://www.openoffice.org/}{Open Office}: document typesetting, data pre-processing and preliminary statistical evaluation

		\item \href{http://www.spss.com/}{SPSS (PASW Statistics 18)}: statistical data edition and analysis 
		\item \href{http://gretl.sourceforge.net/}{Gretl: Gnu Regression, Econometrics and Time-series Library} for statistic analysis
		\item \href{http://staruml.sourceforge.net/en/}{StarUML}, \href{http://live.gnome.org/Dia}{Dia}, \href{http://argouml.tigris.org/}						{ArgoUML}: \ac{UML} diagram tool for creating flowcharts of work algorithms
		\item \href{http://inkscape.org/}{Inkscape}: drawing
	\end{itemize}

	\vspace{1ex}
	\textbf{Auxiliary Tools}:
	\begin{itemize}
		\item \href{http://www.lyx.org/}{LyX} for exporting \LaTeX document to the Word .doc format via HTML
		\item \href{http://calc2latex.sourceforge.net/}{Calc2LaTeX}: converting spreadsheets into \LaTeX environment
		\item \href{http://openev.sourceforge.net/}{OpenEV}: used for raster Google Earth imagery viewing, visualisation and preliminary analysis
		\item \href{http://www.notepad-plus-plus.org/}{Notepad++}: auxiliary text editing and raw data formatting
	\end{itemize}

\pagebreak

\section*{}
\setcounter{lofdepth}{2}
\markright{List of Figures}
\addcontentsline{toc}{section}{List of Figures}
\onehalfspacing
\listoffigures
\pagebreak

\section*{}
\markright{List of Tables}
\addcontentsline{toc}{section}{List of Tables}
\listoftables
\pagebreak

\pagenumbering{arabic}
\rfoot{MSc Thesis\\ \emph{"Seagrass mapping and monitoring along the coasts of Crete, Greece"}}
\section{Introduction}\label{sec:1}
%%%%%%%%%%%%%%%%%%  Section 1   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\counterwithin{figure}{section}
\counterwithin{table}{section}

\markright{I. Introduction}

\subsection{Summary}
The seagrasses, a unique group of aquatic plants growing submerged in the sea water, with root-like
structures (rhizoms) buried in the sediments and vertical elongate leaves. A flowering plant,
completely adapted to the marine environment, they are however, more closely related to the lily family
(\textit{Liliaceae}) than to the true grass, despite their name \textit{seagrass}, caused by the ribbon-like, grassy leaves
\cite{McKenzie08}\label{McKenzie08}. Seagrasses\index[concepts]{Seagrasses} create unique, complex, extremely diversified and productive ecological systems in the
littoral coastal zones between 0-50 meters in shallow waters all over the world \cite{Hogarth07}\label{Hogarth07}, and
serve as valuable environmental indicators for the marine ecosystems health. Seagrasses are closely
connected and linked with complex interactions to other vegetation types, e.g. mangroves, coral reefs,
etc. An important constructing component of littoral ecosystems, seagrass contributes significantly to
their structure and functioning.\\
The adaptation to the salt waters is evidently influenced the global distribution of the seagrasses,
limiting it to shallow coastal areas. A number of critical conditions determine growth of the seagrass,
including general climatic characteristics of the area, i.e.temperature, day length, geological and
geomorphological conditions, e.g. soft type of sediments (sand or mud), shallow depths, as well as
chemical and physical parameters of the water: salinity, waves \cite{McKenzie02}\label{McKenzie02}.
The seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}) is a key species to inhabit littoral of the
Mediterranean Sea, see Fig. \ref{fig:1}

\begin{figure}[h]
	\centering
	\subfloat {\label{fig:1-1a}\includegraphics[width=0.4\textwidth]{Fig-1-1a.jpg}}
	\hspace{2mm}
	\subfloat {\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-1-1b.jpg}}
	\caption{Seagrass \textit{Posidonia Oceanica}}
	\label{fig:1}
\end{figure}

and is widely spread along the coasts of Crete \cite{Dumay02} \label{Dumay02}. 
It plays an important role in a number
of geomorphological and ecological processes. Namely, it is a source of food for herbivorous fauna as
well as shelter zones for fish and other marine organisms; it contributes to the nutrient recycling; it
provides sediments stability by reducing the degree of water movements, etc \cite{Francour99}\label{Francour99}.\\
The purpose of current MSc research work aims to apply methods of remote sensing analysis, including \ac{RTM}, 
\ac{GIS}-based spatial analysis, processing and classification of satellite and aerial photos, as well as videometric underwater footage, towards mapping and environmental monitoring of seagrass \textit{P. oceanica} along the selected locations of the northern coasts of Crete island, Greece. The technical implementation is based on \ac{WASI} RTM software, \ac{GIS} (ILWIS, Erdas Imagine and ArcGIS), using aerial and satellite images and the results of the underwater videometric measurements.

\subsection{Background}
\subsubsection{Global distribution of the seagrasses}
Globally, there are 58 recognized and described seagrass species\index[concepts]{Seagrasses!global species} (Fig. \ref{fig:2}), belonging to two orders
(\textit{Hydrocharitales} and \textit{Najadales}), four families (\textit{Hydrocharitaceae, Posidoniaceae, Cymodoceaceae
and Zosteraceae}), and 12 genera (\textit{Enhalus, Thalassia, Halophila, Posidonia, Syringodium, Halodule,
Cymodocea, Amphibolis, Thalassodendron, Zostera, Heterozostera} and \textit{Phyllospandix}) \cite{Kuo89}\label{Kuo89}.

\begin{wrapfigure}{r}{0.4\textwidth}
	\centering
	\includegraphics[scale=0.30]{Fig-1-2.jpg}
	\caption{Distribution of seagrasses in the world.\\ Source:\cite{Green03}\label{Green03} 
	\label{fig:2}} 
\end{wrapfigure}

The distribution of the seagrasses is strongly influenced by several environmental factors, which
include climate (mostly, tropical and temperate areas\index[concepts]{Seagrasses!distribution of}), bathymetry (shallow shelf zones), hydrological
particularities (chemical content of water, nutrient availability and turbidity of waves), and geological 
characteristics - sedimentation and cover types of the seafloor \cite{McKenzie06}\label{McKenzie06}.\\
There are four European seagrass species \index[concepts]{Seagrasses!Mediterranean species} in Mediterranean area \cite{Borum04}\label{Borum04}: \textit{Zostera marina,
Zostera noltii, Cymodocea nodosa} and \textit{P.oceanica}. In Greece the common species are \textit{P.oceanica}
\textit{(L.) Delile, Cymodocea nodosa (Ucria) Ascherson, Zostera noltii Hornemann} and \textit{Halophila
stipulacea} \cite{Amoutzopoulou-Schina05}\label{Amoutzopoulou-Schina05}. 
These species differ in morphological and phenological features (Fig. \ref{fig:3}) as well as in structure and dynamics. Thus, \textit{Cymodocea nodosa} is
considered the pioneer species of \textit{P.oceanica} beds, the latter species forming the last stage. When \textit{P.oceanica} beds regresses, \textit{C.nodosa} often replaces them \cite{DenHartog77}\label{DenHartog77}; as a result, \textit{P.oceanica},
\textit{C.nodosa}, and \textit{Z.noltii} do not form mixed persistent stands \cite{Buia91}\label{Buia91}.

\subsubsection{Ecological significance of the seagrasses}
Seagrass plays vital role in the marine ecosystems of the world ocean. 
Seagrasses are the only flowering plant in the world that is able to live completely submerged. Seagrass is a habitat for
numerous marine fish species \cite{Nagelkerken00}\label{Nagelkerken00}, source of primary production and food for
fish, turtles and other organisms, which gives them special environmental value \cite{Noralez10}\label{Noralez10}.\\
Seagrass meadows produce enormous quantities of organic matter (leaves, epiphytes), which
constitutes the basis of the food web both within and outside the ecosystem \cite{Gobert06}\label{Gobert06}.
Finally, seagrasses are an important component in the environmental "food chain" of the coastal
ecosystems, being the food source for dugongs, turtles, swans and various fish \cite{Cappo95}\label{Cappo95}.

\begin{wrapfigure}{r}{0.4\textwidth}
	\centering
	\includegraphics[scale=0.25]{Fig-1-3.jpg}
	\caption{Morphology of different types of seagrasses. Source: \url{http://www.ozcoasts.org.au}}
	\label{fig:3}
\end{wrapfigure}

Due to their wide distribution, meadows size, easy collection and abundance, sensitivity to the
modifications of the coastal zone and their important role in maintaining coastal water quality
and clarity, seagrass is perfect indicator and descriptor of the environmental health of marine
ecosystems, and is highly suitable for the environmental monitoring \cite{Pergent-Martini05}\label{Pergent-Martini05}.\\
Being often confused with marine “algae”, “seagrasses” are vastly different from them. There are
fundamental differences between both marine organisms, the major of them should be briefly
mentioned: first, seagrasses are true plants with root system and leaves which photosynthesis, have
complex structure and create landscape-similar vast formations on the colonized seafloor areas with
soft sediment, whereas algae are simple organisms that can only holdfast; secondly, seagrasses are
complex vascular plants with reproductive mechanism such as fruits, seeds and spores, while algae
have simple few cell structure with spores and gametes; finally, seagrasses uptake nutrient through
root system, while algae nourish directly from the water column \cite{Dixon05}\label{Dixon05}. There are other
differences between “seagrass” and “algae” but they go beyond the scope of this work.

\subsubsection{Environmental vulnerability of the seagrasses}
Meadows of \textit{P. oceanica} are subjected to the human activities\index[concepts]{Seagrasses!ecological vulnerability}, as they occur in coastal areas, where
they can be affected both directly \cite{Meinesz91}\label{Meinesz91} or indirectly, through the impact on the
quality of waters and sediments \cite{Duarte02}\label{Duarte02}. As \textit{P.oceanica} is a long-living plant with a slow
growth rate, the anthropogenic modifications of the coastal zone, happening more rapidly than the
capacity of the plant to adapt to these changes, reduce its distribution area \cite{Micheli05}\label{Micheli05}.

\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.25]{Fig-1-4.jpg}
	\caption{Distribution of seagrass in relation to mean ocean temperature. \\
	Source: \cite{Orth06}\label{Orth06}}
	\label{Fig.4}
\end{wrapfigure}

One of the main drivers of seagrass decline is, for instance, the location of the fish farming near the
seagrass meadows. The negative effects of the sedimentation of waste particles in the farm vicinities
on \textit{P.oceanica} meadows are diverse and complex, and may cause benthic deterioration, accumulation
of organic matter and seagrass decline \cite{Holmer08}\label{Holmer08}. Seagrasses are subject to anthropogenic
nutrient (N and P) loading, which may occasionally casue morphological (e.g. leaf length) and
physiological (e.g. chlorophyll and nitrogen content of the leaves) responses towards changed
environmental conditions \cite{Leoni06,Leoni07}\label{Leoni06} \label{Leoni07}. 
The detailed research of the fish
farm-induced decline of the seagrass meadows \cite{Diaz-Almela06}\label{Diaz-Almela06}reports the relationships of
fish farm organic and nutrient content in the sediments with dynamics of the key seagrass species (\textit{P.
oceanica}) in the Mediterranean Sea. Nowadays \textit{P.oceanica} is in the alarming state of regression due
to the deterioration of the environment in the Mediterranean Sea \cite{Ribed02}\label{Ribed02}. Due to these reasons,
\textit{P.oceanica} is a protected species since 1988 in some European countries \cite{Francour99}\label{Francour99}, and
its presence serves as an indicator of a stable healthy environment. Among other negative factors,
affecting both growth and status of the seagrasses the environmental contaminants can be mentioned,
e.g. thermal, sewage, dredging and chemical pollution\index[concepts]{Seagrasses!anthropogenic impacts on}  as well as any other kind of maritime works,
e.g. trawling and anchoring of boats \cite{Ribed02}\label{Ribed02}. Other human activities that cause degrading of the
seagrass are recreational boating, commercial overexploitation of coastal resources, eutrophication
\cite{McKenzie06}\label{McKenzie06}.\\
Besides anthropogenic factors, various biochemical, climatic and environmental processes\index[concepts]{Seagrasses!natural impacts on} can cause
negative influence on seagrass distribution. Seagrass is exposed to threats from the global climate and
environmental change, i.e. increases in sea surface temperature; sea level rise; increased frequency
and intensity of storms and waves; local decrease of water quality, increased sedimentation,
contamination and nitrification; desiccation; salinity fluctuations; nutrient changes; suspended
sediments \cite{Blake00}\label{Blake00}. These stress-drivers can alone result in large-scale seagrass
degradation, but often seagrass undergo simultaneous affects from several of these factors together.
It naturally increases the environmental pressing  to drastic loss of very large areas of
seagrass globally \cite{Orth06}\label{Orth06}.
In tropical areas, where most of seagrasses are located (Fig.\ref{Fig.4}), seagrasses are subjectto catastrophic
extinction and loss, due to the cyclones, typhoons, storms, regular floods and increased rainfalls. Recovery from such
events can take up to several years and often it is only possible by means of the seed reserves from the local
environmental surveys \cite{McKenzie07}\label{McKenzie07}. Other threats for seagrasses in tropical areas are 
increased nutrient availability in the coastal zones, increased
eutrophication and invasive macroalgae. These processes have strong affect on the status of the
seagrass meadows, and often lead to their complete disappearance \cite{Holmer09}\label{Holmer09}.\\
Other environmental threat for the seagrasses arises as a result of the environmental struggle and
competition for existence among species. Thus, meadows of P.oceanica in the Mediterranean Sea are
presently facing invasion by alien algal species, particularly in areas where \textit{P.oceanica} is already
degrading, stressed, have gaps and patchy structure in meadows and show other signs of regression
\cite{Montefalcone10}\label{Montefalcone10}.
Seagrasses are vulnerable fragile species, important for the marine coastal ecosystems, especially for
the protection of the beach structure. However, the facts about seagrass global degrading sound
worrying: about 54 \% of the total seagrass meadows have lost any part of their area; the areas, where
the seagrass ecosystems are degrading or lost, are not located in a specific area or continent, but
registered globally; since 1980s global losses of the seagrasses on our planet is equal to two football
fields per hour \cite{Mellors09b}\label{Mellors09b}.
\pagebreak

\subsubsection{General characteristics of \textit{P.oceanica}}\index[concepts]{Seagrasses!\textit{P.oceanica}}
The endemic Mediterranean seagrass \textit{Posidonia oceanica} (further \textit{P.oceanica}), is a main species in
marine coastal environment of Greece, forming, despite its slow growth, the largest, most widespread,
homogeneous and dense meadows in the Mediterranean between 5 and 40 m depth \cite{DenHartog70}\label{DenHartog70}.

\begin{wrapfigure}{r}{0.4\textwidth}
	\centering
	\includegraphics[scale=0.20]{Fig-1-5.jpg}
	\caption{Geographical distribution of
	\textit{P.oceanica}. Source: \cite{Borum04}\label{Borum04}}
	\label{fig:5}
\end{wrapfigure}

The dominant and most productive coastal ecosystem of the Mediterranean, \textit{P.oceanica} is
spatially restricted to the Mediterranean area\index[concepts]{Seagrasses!\textit{P.oceanica}!distribution} (Fig.\ref{fig:5}), 
with its extension limited by the western part of the Mediterranean Sea where cold Atlantic waters enter Gibraltar and mix with warm
Mediterranean waters, thus decreasing its temperature.
Morphologically\index[concepts]{Seagrasses!\textit{P.oceanica}!morphology}, \textit{P.oceanica} consists of long, 5-12 mm broad, “hairy-like” leaves, 3-4 mm thick roots and short rhizomes (0.5-2.0 mm). 
The leaves are, perhaps, the most particular characteristics of \textit{P.oceanica}, making it highly recognizable and distinguishable from other seagrasses (Fig.\ref{fig:8}(a)): having usual length of 20-40 cm, in some cases they can reach up to 1 m \cite{Borum04}\label{Borum04}  (Fig.\ref{fig:8}(c)).
Growing \textit{P.oceanica} make a meadows which, in turn, consist of smaller patches, called \textit{matte}
(Fig.\ref{fig:8}(b)), a monumental construction made by the growth of rhizomes and leaves with entangled roots
and entrapped sediment \cite{Francour06}\label{Francour06}.

\begin{figure}[H]
	\centering
	\subfloat [\textit{P.oceanica}. Source \cite{Luque04}\label{Luque04}]{\includegraphics[width=0.25\textwidth]{Fig-1-6.jpg}}
	\subfloat [Scheme of matte structure of \textit{P.oceanica}. Source\cite{Pergent90}\label{Pergent90}]{\includegraphics[width=0.3\textwidth]{Fig-1-7.jpg}}
	\subfloat [Structure and components of \textit{P.oceanica}. Source\cite{DiCarlo04}\label{DiCarlo04}]{\includegraphics[width=0.3\textwidth]{Fig-1-8.jpg}}
	\caption{Morphology of \textit{P.oceanica}}\label{fig:8}
\end{figure}

Representing one of the most productive Mediterranean ecosystems, \textit{P.oceanica} usually serves as a
perfect biological indicator for the assessment of the quality of waters and environmental health
\cite{Boudouresque89}\label{Boudouresque89}. Some authors \cite{Guidetti08,Montefalcone09}\label{Guidetti08} \label{Montefalcone09} used status and population dynamics of \textit{P.oceanica} as indicators for the evaluation of the meadow health status. 
There are several environmental factors, determining the growth of the \textit{P.oceanica}. 
The adaptation\index[concepts]{Seagrasses!\textit{P.oceanica}!environmental adaptation}
to dry-summer subtropical climate reduces its extension to Mediterranean area only (Fig.\ref{fig:5}).
Besides, the distribution of the seagrasses changes with water depth: it is noticed \cite{Dural10}\label{Dural10} that
the highest flowering density is usually in the 4-7 m depth. 
\textit{P.oceanica} flowers appeared in shallow
stands in September while in November only in stands deeper than 15 m. This time delay is caused by
the different maximum summer temperatures at those depths \cite{Buia91}\label{Buia91}.
The phenology\index[concepts]{Seagrasses!\textit{P.oceanica}!phenology} of \textit{P.oceanica} is also affected by the coastal bathymetry: in the isolated meadows in shallow waters plants have shorter and falciform leaves, compared to ones in the deeper and central
areas \cite{Dural10}\label{Dural10}. In \textit{P.oceanica} flower abundance is related to the structure of the meadow with
the maximal flower density in the densest stands, while the occurrence of flowering is regulated by
environmental factors \cite{Buia91}\label{Buia91}. Phenology of the \textit{P.oceanica} undergoes 
modifications with varying seasons during the year: during the flowering period (ca 3 months long)
the number of leaves on the flowering shoots decrease. Changes of the leaf growth also appear in the
flowering shoots with longer oldest leaves and shorter and narrower leaves induced during the
flowering \cite{Gobert01}\label{Gobert01}.

\subsection{Research problem}\index[concepts]{Research}
Monitoring\index[concepts]{Research!problem} of the marine benthic ecosystems of seagrasses is essential for the environmental
assessment of the coastal zones. It increases our knowledge of the seagrass ecology, highlights 
threats to the seagrass and preventing them from possible losses and degrading and improves
techniques and methods of the underwater-based observations. Mapping the seagrass contributes to
the evaluating of the seagrass current distribution, analysis of its dynamics and changes over time, as
well as estimations of the degrading of seagrass meadows for the purpose of the coastal management.\\
Precise, correct and up-to-date information about the distribution of \textit{P.oceanica} is necessary for the
sustainable conservation of the marine environment and ecosystems in Mediterranean area, being an
important contribution to the environmental coastal zone management \cite{Pergent-Martini06}\label{Pergent-Martini06}.
However, mapping the seagrass has limitations due to the specific location and characteristics of the
research object. The remote sensing techniques have traditionally been widely used for the seagrass
monitoring.

\begin{wrapfigure}{r}{0.4\textwidth}
	\centering
	\includegraphics[scale=0.25]{Fig-1-9.jpg}
	\caption{Spectra of the seagrass on different depths (0 - 15 m). Source \cite{Edwards00}\label{Edwards00}}
	\label{fig:9}
\end{wrapfigure}

The general overview of the application of various remote sensing data types (colour,
infrared, and black and white images) for the seagrass monitoring shows its high suitability and
potential as a research method \cite{Pasqualini01,Matarrese06}\label{Pasqualini01}\label{Matarrese06}. The using of the aerial
photographs as base maps for the seagrass meadows mapping is, perhaps, the most traditional
application \cite{McKenzie03,Kendrick00,Pasqualini99}\label{McKenzie03} \label{Kendrick00} \label{Pasqualini99}. The results of the image
processing of colour aerial photographs for the monitoring of littoral environment with seagrass beds
have been reported by several authors \cite{Kelly80,Walker89,Green96}\label{Kelly80} \label{Walker89} \label{Green96}.
Satellite imagery processing has also being used for the seagrass monitoring, due to their accuracy,
repeatability and information value as a source of data \cite{Dekker05b}\label{Dekker05b}, enabling regular temporal
coverage over the large remote areas and providing a cost-effective approach for the mapping of the
remotely located feature, such as underwater vegetation \cite{Jensen95}\label{Jensen95}. Satellite images provide
with detailed information on seagrass canopy and other environmental indicators \cite{Fyfe03}\label{Fyfe03}.
Various research papers report successful application of the image processing for the
seagrass mapping \cite{Calvo03,Dekker05a,Fornes06,Green96,Jackson07,Jensen95,Kendrick00,Lyzenga81,Malthus03,Matarrese08,Mount03,Pasqualini01,Pasqualini99,Pasqualini98a,Pergent-Martini06,Ralph05,Ribed02,Short01,Walker89}\label{Calvo03}\label{Dekker05a}\label{Fornes06}\label{Green96}\label{Jackson07}\label{Jensen95}\label{Kendrick00} \label{Lyzenga81}\label{Malthus03}\label{Matarrese08}\label{Mount03}\label{Pasqualini01}\label{Pasqualini99}\label{Pasqualini98a}\label{Pergent-Martini06}\label{Ralph05}\label{Ribed02}\label{Short01}\label{Walker89}. The application of the remote sensing data towards seagrass mapping is based on the spectral reflectance
 characteristics of the \textit{P.oceanica} seagrass, which enable its spectral discrimination from spectra of other seafloor types. 
It is proved \cite{Thorhaug07} \label{Thorhaug07} that the spectral signatures of different species of tropical seagrasses are well
distinguishable from each other. The application of the methods of images classification for seagrass
mapping is based on the classifying the pixels on the image according to their spectral reflectance
values \cite{Fornes06} \label{Fornes06}, so that the seafloor can be divided into several types: sand, rock, P.
oceanica, other vegetation, etc.
Other example of monitoring of \textit{P.oceanica} using remote sensing techniques \cite{Calvo03}\label{Calvo03}
reports the application of the \ac{CZCS} images towards the case study of the Italian coast and shows
successful results of the neural-based classification using Isodata method of supervised classification.
In case of \textit{P.oceanica} meadows aerial and satellite images are particularly suited for the surveying
shallow waters \cite{Pasqualini98b}\label{Pasqualini98b} enabling to distinguish seagrass formations and dynamics of
the temporal evolution of seagrass meadows over the research area \cite{Pasqualini01}\label{Pasqualini01}.
However, using space borne satellite imagery for the seagrass mapping has certain limitations, due to
the uncertainties of the spectral signature of the seagrass at higher depths  (Fig.\ref{fig:9}), as well as some
optical particularities, e.g. light refraction under water, unevenness of the water surface, depths, etc.
Some problems can also arise in the images interpretations, as quite different objects may have
similar spectral reflectance, e.g. seagrass, dark-coloured bottom patches (mud), macroalgae.
The \textit{in-situ} fieldwork including underwater videographic measurements is an important part of the
seagrass monitoring, and has been successfully applied towards seagrass mapping \cite{Haag08}\label{Haag08}.
The underwater measurements are used to validate the results and to receive detailed, accurate and
precise data for the selected locations. The underwater measurements cannot be applied for the whole
research area, however it provides with detailed monitoring along the route of the boat. Therefore, in
the selected locations it becomes a useful tool for the assessment of the distribution, density and
coverage of the seagrass along the track log. Besides, the underwater observations using scuba diving
equipment have been conducted for the measurements of depths.

\subsection{Research objective}\index[concepts]{Research!objective}
The current MSc research aims to explore the environmental conditions for the spatial distribution of \textit{P.oceanica} seagrass along the northern coast of Crete Island, based on the remote sensing and \ac{GIS} techniques, knowledge about the coastal environment in Crete and integration of various data
from the following sources \index[concepts]{Data}:
	\begin{itemize}
		\item[] i) spectra of \textit{P.oceanica}, carbonate sand, silt and other seafloor types
		\item[] ii) satellite imagery: Landsat TM, Landsat ETM+
		\item[] iii) aerial photos: Google Earth
		\item[] iv) \textit{in-situ} fieldwork data of underwater videographic measurements
		\item[] v) vector GIS layers.
	\end{itemize}
The research aims to explore the limitations of the application of the \ac{RTM} and remote sensing techniques 
towards the study of the environmental properties of the \textit{P.oceanica} meadows distribution
over the selected locations of northern Crete. The main research objective is monitoring the seagrass \textit{P.oceanica} in selected areas of Crete and analyzing the environmental conditions for the growth of \textit{P.oceanica}. \\
This research is supported by the \textit{in-situ} measurements in two selected locations of the northern coast
of Crete (Ligaria beach in Agia Pelagia district and Xerocampos), using following methods of the
remote sensing techniques: spectral modelling by means of the\ac{RTM} \ac{WASI},  underwater videometric measurements made by \ac{Olympus ST}camera, aerial Google Earth and satellite images from different sources, spatial \ac{GIS} and statistical analysis.

\subsubsection{General objective}\index[concepts]{Research!objective!general}
The main objective of this study is to analyse the optical properties of the seagrass \textit{P.oceanica} and other seafloor types, and to apply the remote sensing techniques towards the investigation of the seagrass distribution in the selected locations along the northern coast of Crete.
General objectives:
	\begin{itemize}
		\item[*]Analysing spectral reflectance of the \textit{P.oceanica} and other seafloor cover types means of radiative transfer model tools (RTMs), using WASI.
		\item[*]Mapping spatial distribution of the seagrass \textit{P.oceanica} over selected locations along the northern coasts of Crete Island 
	\end{itemize}

\subsubsection{Specific objectives}\index[concepts]{Research!objective!specific}
	\begin{itemize}	
		\item[]1.	To study narrow-band spectral reflectance properties of \textit{P.oceanica} and other seafloor cover types (sand and silt) using WASI water colour simulation software 
		\item[]2.	To use methods of the in situ diving observations and underwater videometric measurements by \ac{Olympus ST}camera in order to receive large-scale imagery of the P.oceanica mattes 
		\item[]3.	To apply remote sensing data (Google Earth aerial images, \ac{Landsat TM} and ETM+ satellite images) for the monitoring of the seagrass meadows distribution 
		\item[]4.	To perform supervised images classification for the thematic mapping of the \textit{P.oceanica} seagrass distribution along the selected locations over the coasts of northern Crete. 
	\end{itemize}

\subsection{Research questions}\index[concepts]{Research!questions}
	\begin{itemize}	
		\item Is \textit{P.oceanica} spectrally distinct from other seafloor cover types with varying \textit{in-situ} conditions ?
		\item Whether broadband and hyperspectral remote sensing data can be used for the mapping of \textit{P.oceanica} ?
	\end{itemize}

\subsection{Hypotheses testing}\index[concepts]{Research!questions!hypothesis}
A statistical testing will be used to compare between the spectral responses of the different seafloor cover types (i.e. sand and \textit{P. oceanica}),
 whether it is spectrally distinct and at least one pair is statistically different at every spectral
band.
Hypothesis Ho: seagrass types are not spectrally distinct from other seafloor types with
varying \textit{in-situ} conditions, which means
Ho: μ1 =μ2 =μ3=...= μn.
The alternative Hypothesis Ha claims the opposite statement: seagrass is spectrally
distinct with varying \textit{in-situ} conditions, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn.
The distribution of the spectral responses at every spectral band is assumed to be normal as well as
the equality of the statistical variances.
The hypothesis testing is suggested to be carried out using the \ac{ANOVA} statistical test. The purpose of
ANOVA test is to visualize in an effective and quick way the spectral differences between seagrass
species and their spatial distribution. Thus, the key hypotheses of the research will be tested to prove
whether the results of the research are accurate, reasonable and correct.

\subsection{Research approach}\index[concepts]{Research!approach}
Seagrass consistent monitoring and mapping is necessary and important for the sustainable coastal
development and conservation measures. Earlier, many seagrass meadows have been destroyed by
human activities in the coastal zone, mainly due to the ignorance of their existence, because
information on the seagrass bed exact location was not available \cite{Choo06}\label{Choo06}. Well-time seagrass
observations and mapping enables precise control of its spatial distribution, detection of any changes
in the seagrass landscapes, highlights potential environmental threats in the coastal zone (e.g.
declining of meadows) before they become unmanageable for the coastal management services.
Choosing the right and most effective approach method for the seagrass monitoring is essential.
Remote sensing methods alone, though having evident advantages, are insufficient, because satellite
images of underwater habitats are notoriously difficult to identify and interpret. The best research
method should be based on the integrated approach, well described in various scientific works \cite{Brown02,Montefalcone06,Kirkman96}\label{Brown02} \label{Montefalcone06} \label{Kirkman96}, which includes combination of various
techniques of the seagrass monitoring, i.e. remote sensing imagery classification of aerial and satellite
images, GIS-based spatial analysis and ground \textit{in-situ} surveys.
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.48]{UML_ClassDiagram.jpg}
	\caption{General methodological approach for the analysis of spectral signatures. \href{http://argouml.tigris.org/}{ArgoUML}}
	\label{fig:1.10}
\end{figure}
The current study is based on the application of the remote sensing data, broadband satellite imagery,
aerial images and the results of the underwater videographic measurements towards seagrass
mapping  (Fig.\ref{fig:1.11}). Image classification is based on the principle of the differentiation between the spectral
signatures of various seafloor cover types  (Fig.\ref{fig:1.10}). The spectrum of light coming up from the ocean surface in
shallow waters keeps information on the optical properties of the seawater components and benthic
substrate which can be read from their spectral signatures \cite{Werdell03}\label{Werdell03}. \\
The spectral reflectance, distinct for each seafloor cover type, enables their discrimination on the image. The pre-processing
of the images includes imagery corrections for atmospheric noises and effects of the water
column. Reflectance spectra of the seagrass canopy at different depths of the water-column are
analysed for the discrimination of their spectral signatures, enabling to separate various seafloor types
during classification. The results of the of imagery classification are analysed for the detection of the
dynamics in \textit{P.oceanica} seagrass distribution along the northern coasts of Crete. Aerial imagery from
Google Earth with high spatial resolution, suitable for the large-scale detailed mapping of seagrass
mattes, is used for the improvement of the accuracy of large seagrass meadows and separate mattes
within the meadows. The \textit{in situ} underwater videometric measurements of the seafloor are collected
during the fieldwork in Crete, for the validation of the classification results and to determine the exact
current distribution of the \textit{P.oceanica} meadows. The image processing includes steps of the remote
sensing techniques, i.e. calibration, masking from land and cloud, atmospheric correction, sea surface
glint and depth effects correction as recommended \cite{Matarrese08}\label{Matarrese08}. During the image classification
working step the training sites for the supervised classification methods are designed, as well as its
control and trials of different classification approaches (Unsupervised, K-means or Isodata;
Supervised, Maximum Likelihood).

%%%%%%%%%%%%%%%%%  Section2   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Overview of literature]{Seagrass monitoring: overview of literature and research resources} \label{sec:2}
\markright{II. Seagrass monitoring: overview of literature and research resources}
\renewcommand{\headrulewidth}{0.4pt}
\counterwithin{figure}{section}
\counterwithin{table}{section}

\subsection{Seagrass global monitoring: history and perspectives}\index[concepts]{Seagrasses!global monitoring}
Mapping and monitoring the seagrass is important for the environmental assessment of the marine
ecosystems in coastal areas. Regular tracking of current distribution of seagrass meadows, based on
correct information and cartographic visualization of seagrasses, is a preventive environmental
management, which helps to analyse potential environmental risks of coastal areas, decrease of the
number of species, loss of meadows and patches of the seagrasses.
The tradition of global seagrass mapping though has not a very long history comparing to the
terrestrial cartography, due to the technical difficulties of underwater observations, yet nowadays is
has become a successful, rapidly developing and increasingly popular and challenging research branch.\\
Seagrasses can be mapped using various approaches, with the most popular method of the remote
sensing application, enabling to detect and discriminate the substratum and various vegetation types
growing on the seafloor.\\
Several global seagrass surveys\index[concepts]{Seagrasses!global surveys} organize and provide regular monitoring of the seagrass distribution,
health and species sustainability, amongst which the most world-known are, perhaps, the following
ones: \href{http://www.seagrassnet.org/}{Global Seagrass Monitoring project}, \href{http://www.seagrasswatch.org/}{Seagrasswatch}, 
located in Australia, two European-located institutions - \href{http://mediterranean.seagrassonline.org/}{the Mediterranean association Seagrass-2000} and \href{http://www.imedea.uib.es/index.php}{the Mediterranean Institute for Advanced Studies}, US American 
\href{http://www2.fiu.edu/~seagrass/}{Seagrass Ecosystems Research Laboratory in South Florida} and
\href{http://www.flseagrass.org/}{Florida Seagrass organisation}, the \href{http://www.seagrasses.org/}{Seagrasses.org} and the \href{http://wsa.seagrassonline.org/}{World Seagrass Association}.
These organizations aim at the global seagrass monitoring, providing with
research results and reporting guidelines and manuals with standardized methods and
recommendations, specific for the seagrass research and monitoring. There are also university marine
centres and research institutes conducting seagrass monitoring and as a particular part of their
research. Their reports and guidelines were used for references in the current research.\\
Regular observations and monitoring of the seagrasses are known since 1960s, mainly in tropical
regions (Australia). Since that time traditional methods of the seagrass monitoring and common
recommendations are being elaborated. The development of the underwater SCUBA diving
equipment and devices enabled to conduct underwater detailed measurements and observations
largely contributed to the improvement of the traditional \textit{in-situ} observations of seagrass. From the
other side, development of the remote sensing methods and data acquisition from space contributed to
the new methods of seagrass mapping, using distance approach and generally based on images
classification.

\subsection[Measuring water optical properties]{Measuring water optical properties: \\hyperspectral radiometers}
The application of the remote sensing data for seagrass mapping is based on our knowledge of the
spectral reflectance properties of the target objects, and using it for the classification of these objects
on the image. In case of seagrasses it is spectral reflectance of the seafloor cover types, which can be
analyzed using measurements of optical properties of sea water: radiance and irradiance.
Optical remote sensing methods can get through the clear waters to approximately 15–30 m \cite{Mumby04}\label{Mumby04}.
When sunlight enters the waters and goes down into the water column, parts of the
electromagnetic energy are absorbed and scattered, which is determined by the optical and physical
properties of the water, e.g. concentration of suspended particles, chlorofill, colored dissolved
organic matter (gelbstoff) that make up the water content \cite{Ackleson2003}\label{Ackleson, 2003}. Besides, light is strongly
dependent on wavelengths, i.e. it is greater in blue wavelengths (400 nm) than in others.

\begin{figure}
	\centering
	\subfloat [RAMSES-ACC-UV - Hyperspectral UVA/UVB Irradiance Sensor: 280-500 nm]{\label{fig:1-1a}\includegraphics[width=0.4\textwidth]{Fig-2-1.jpg}}
	\subfloat [RAMSES-ARC - Hyperspectral UV-VIS Radiance Sensor: 320-950 nm.]{\label{fig:1-1b}\includegraphics[width=0.4\textwidth]{Fig-2-2.jpg}}
	\caption{RAMSES Hyperspectral Sensors. Source: \href{http://www.trios.de}{Trios}}
	\label{fig:2.1}
\end{figure}

The optical properties\index[concepts]{Seawater optical properties} of the sea water vary with different environmental conditions and reflect
current chemical content and physical specifics of the water, revealing the variability and distribution
of color of the sea waters, determined by the material in the water, e.g. chlorophyll, as well as its
physical properties, e.g. water absorption, attenuation, backscattering \cite{Maffione01a}\label{Maffione01a}.
Shallow waters generally contain more dissolved substances and suspended particles, which directly
influences the transparency and color of the waters of shelf zones \cite{Jerlov52}\label{Jerlov52}. Being highly
dynamic environments, coastal waters experience a variety of processes which alter their optical
properties incessantly. The effects of these processes influence application of the hyperspectral
remote sensing\index[concepts]{Seawater optical properties!hyperspectral measurements} and reinforce other processes \cite{Maffione01b}\label{Maffione01b}. Thus, waves and tides increase
sedimentation processes, which in turn, may change micro relief properties and topology. \\
The optical properties of the water are best reflected in the values of its radiance and irradiance,
which can be converted into spectral reflectance, or reflectivity. The spectral irradiance (E) is a
radiant flux of the electromagnetic solar radiation energy, received per surface unit area in a given
time ($W*m^{−2}*nm^{−1}$), while radiance (L) characterizes total emission or reflection that passes through
or is emitted from a particular area ($W*sr^{−1}*m^{−2}*nm^{−1}$). Therefore, the spectral reflectance, or the
reflectivity of the object, can be estimated by the direct mathematical division of these first two
characteristics, and is expressed in percentage.
The irradiance\index[concepts]{Seawater optical properties!irradiance} and radiance\index[concepts]{Seawater optical properties!radiance} of the water thus should be measured, in order to estimate spectral
reflectance of the various seafloor cover types. \\
The optical measurements of the irradiance and radiance of the sea water and bottom cover types of the seafloor can be received by the means of the
optical sensors spectroradiometers\index[concepts]{Seawater optical properties!spectroradiometers}. There are several companies producing radiometers with various
characteristics and adjusted for different purposes, e.g. portable and miniature spectrometers from
\href{http://www.stellarnet.us}{StellarNet}, Hyperspectral Ocean Colour Radiometer (HyperOCR) sensor by
\href{http://www.satlantic.com}{Satlantic}, traceable spectroradiometers by \href{http://www.orboptronix.com}{Orbotronix} 
and lots of others. Among other radiometers there are ones designed by the \href{http://www.trios.de/}{Trios company} producing
optical sensors, GER-Series Field portable spectroradiometers from
\href{http://www.spectrapartners.nl/}{SpectraPartners}, etc. \ac{Trios-RAMSES}\index[concepts]{Seawater optical properties!spectroradiometers!Trios-RAMSES} hyperspectral radiometers
(\ac{Trios-RAMSES} Hyperspectral \ac{UVA}/\ac{UVB} Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral \ac{UV}-\ac{VIS} Radiance Sensor) are small-sized, low power-consuming, flexible for fieldwork yet with high
level of precision, specially calibrated for air and for water application as well as color
measurements  (Fig.\ref{fig:2.1}). These products have been used for the radiance and irradiance
measurements in Agia Pelagia bay, Crete Island, 2009.

\subsection[Radiative Transfer Models (RTM)]{Radiative Transfer Models (RTM) for the simulation of water optical properties: a brief review of existing tools}

Understanding radiance distribution within a water colum\index[concepts]{Seawater optical properties!RTM}n is necessary for the studies of the underwater visibility, because scattering properties of the water body naturally vary with changing depth, wavelength and environmental conditions.  Simulation of the radiance quantities for natural water bodies enables to analyse seafloor color remote sensing properties. Therefore, the artificial modelling of the seawater optical properties by means of the \ac{RTM} is used when water optics is studied under changing environment (e.g., depths, sun angle, suspended particles in water column). \\ In such cases a retrieval of water optical parameters from the remote measurements should be tested and analyzed. 
\newline
We briefly discuss below the most effective RTM software and algorithms from a range of various best-known up-to-date tools. 
	\begin{itemize}
		\item [*] The \ac{6SV1} \index[concepts]{Seawater optical properties!RTM!6SV1} is a US atmospheric correction algorithm, developed in the University of Maryland; adjusted for the \ac{MODIS} satellite imagery lookup tables. The 6S algorithm has been \href{http://grass.osgeo.org/grass64/manuals/html64_user/i.atcorr.html}{implemented} by the \ac{GRASS} software for the atmospheric correction 				 
		\item [*] Another example of the RTM adjusted for a specific imagery is a German \ac{KOPRA}\index[concepts]{Seawater optical properties!RTM!KOPRA}, fitted for \ac{MIPAS} and \ac{ENVISAT} imagery.
		\item [*]  The \href{http://www.sequoiasci.com/products/Hydrolight.aspx}{\textit{HydroLight RTM}} commercial software is an advanced model for oceanic radiative transfer calculations, developed by the Sequoia Scientific, Inc., USA. It is designed to solve a wide range of problems in optical oceanography and limnology.   
		\item [*] The British \ac{RTTOV-9}\index[concepts]{Seawater optical properties!RTM!RTTOV-9} is a radiative transfer model, developed for nadir viewing atmospheric sounders and imagers. It includes a number of useful tolls, e.g. can compute sea-surface emissivity for each channel, enables, to specify cloudiness for radiance calculations, etc.  This program runs under Unix/Linux platforms and is open source. 
		\item [*] The \ac{RRTM}\index[concepts]{Seawater optical properties!RTM!RRTM} is one more RTM freeware, designed in the USA, Massachusetts. 
		\item [*] The \href{http://www.modtran.org/}{\textit{MODTRAN}}\index[concepts]{Seawater optical properties!RTM!MODTRAN}, a commercial software from Spectral Sciences Inc., is another example of the RTM. Among its newest updates is generation of atmospheric correction data; it also has an option to write spectral output in binary, and convert it to \ac{ASCII}.
		\item [*] A \ac{DISORT}\index[concepts]{Seawater optical properties!RTM!DISORT}  is another RTM freeware , developed as an algorithm for calculation radiance of the scattering and absorbing Earth's atmosphere and transmissivity computations within the waveband range from UV to radar.  
		\item [*] The \ac{CRTM} \index[concepts]{Seawater optical properties!RTM!CRTM} is designed by the  (\href{http://www.jcsda.noaa.gov/}{JCSDA}, \ac{NASA}) and includes Surface Emissivity/Reflectivity Models, Cloud Absorption/Scattering Model and Gaseous Absorption Model.
		\item [*] An \ac{4AOP}\index[concepts]{Seawater optical properties!RTM!4AOP} is a French RTM freeware, designed in cooperation by \href{http://www.noveltis.fr/fr/index.php}{Noveltis} and \href{http://www.cnes.fr/}{the French Space Agency} (\ac{CNES}). Being a fine program which enables accurate computations of the radiance, defining other emissivity functions and other set up, it is however, particularly designed for the infrared waves (thus, not quite suitable in our case). To the advantages of this freeware we would count its user friendly interface and open source accessibility for the academic purposes. 
	\end{itemize}
For our purpose we have chosen the \ac{WASI}\index[concepts]{Seawater optical properties!RTM!WASI} \ac{RTM} software, due to its effectiveness, adaptability for the Mediterranean environment, open source availability, coverage of necessary wavebands and clear, user friendly interface, enabling to adjust various environmental parameters.

\subsection[\textit{In-situ} observations of the seagrass]{The \textit{in-situ} observations of the seagrass}
The traditional methods of \textit{in-situ} seagrass monitoring include in general the following standard
scheme \cite{McKenzieetal03} \label{McKenzieetal03}. The seagrass is being sampled on the selected sites using transect lines,
quadrant frame, single point markers, markers, \ac{GPS} and other equipment. The seagrass sampling is
taken on the regular way with observation points covering the study area with normal distribution.
During the measurement process, the vertical photograph of the measurements frame is taken, and the
following points are traditionally estimated: percentage of the seagrass cover within the quadrate,
species composition, sediment composition, canopy height, epyphyte abundance, algae percent cover,
count of microfauna and a specimen of seagrass is being taken. This scheme, well described by
McKenzie \cite{McKenzieetal03}\label{McKenzieetal03} is widely used and well-known among the marine biologists and
seagrass researchers. Applications of the \textit{in-situ} seagrass observations of the structuring epiphyte
community composition in the \textit{P.oceanica} ecosystems in Mediterranean Sea is, for example,
described by Villegas \cite{Villegas06}\label{Villegas06}. Realization of traditional methods for mapping seagrass usually
involves intensive and time-consuming \textit{in-situ} observations during the fieldwork, as, for example,
reported by Iverson and Bittaker \cite{Iverson86}\label{Iverson86}. \\
Other methods of seagrass \textit{in-situ} monitoring are based on
using the active hydro-acoustic sonar sensors that send towards a sea floor a signal of energy and then
collect the return echoes for the analysis, yet this requires specialized equipment and is mostly used in
the depth waters, combined with bathymetric measurements. Another limitation of the acoustic
techniques is that they are not effective at identifying the composition or even the presence of
submerged aquatic vegetation such as seagrasses and seaweeds \cite{Werdell03}\label{Werdell03}, because
initially acoustic methods were intended for the bathymetric surveying, i.e. they are not adjusted for
the benthic habitat discrimination. The results of the \textit{in-situ} measurements and observation are usually
managed and treated using integrated \ac{GIS} approach, as, e.g. reported by Schmieder \cite{Schmieder97}\label{Schmieder97}.

\subsection[Application of the remote sensing...]{Application of the remote sensing data towards seagrass mapping}\index[concepts]{Remote sensing}
Various methods and approaches have been applied towards mapping of the seagrasses, based on
digitized aerial photographs, \ac{GPS} data, remote sensing and \ac{SCUBA}-based fieldwork measurements.
\ac{SCUBA}-based \textit{in-situ} observations, though providing high resolution and accuracy results in seagrass
mapping, is limited in application, because of their time consumption, weather-dependency and
unsuitability for the case of monitoring large areas of water for small-scale mapping. 
The underwater videography\index[concepts]{Remote sensing!underwater videography} with a \ac{GPS}\index[concepts]{Remote sensing!GPS} is a tool of seagrass monitoring which has certain advantages, i.e.
high spatial and visual resolution, non-destructive sampling, effectiveness at all depths and rapid data
collection in the field \cite{Schultz08}\label{Schultz08}.

\begin{wrapfigure}{r}{0.4\textwidth}
	\centering
	\includegraphics[scale=0.30]{Fig-11.jpg}
	\caption{\textit{In situ} optical reflectance spectra of seagrass. Shaded areas - \% of spectra lying within the range of reflectance. 
		White lines - mean spectra. Source: \cite{Hochberg03b}\label{Hochberg03b}}
	\label{fig:13}
\end{wrapfigure}

However, it cannot cover large areas for small-scale mapping.
Remote sensing techniques offer clear advantages over other methods of \textit{in-situ} field measurements
and seagrass observations, mentioned above. Preference of the remote sensing methods consists in
their weather-independence, cost-effectiveness, accuracy and spatial coverage, which enables
periodic monitoring of the seagrass meadows and gives access to the distant and unapproachable
areas. Integrated together with \ac{GIS} vector layers and maps, remote sensing data enable historical
mapping \cite{Carter08,Ardizzone06}\label{Carter08} \label{Ardizzone06} and assessment of change detection.
However, application of the remote sensing techniques for mapping of submerged vegetation,
seafloor cover types and benthic vegetation, \textit{inter alia} seagrasses, are still in their development.
Approaches and methods for the seagrass protection and monitoring still remain location-specific or, at
least, nation-specific, depending to large extent on the tools available for the researchers \cite{Mellors09a}\label{Mellors09a}.

\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.25]{Fig-12.jpg}
	\caption{Difference between broadband multispectral and hyperspectral resolution of
		spectral signatures. Source: \cite{Clark99}\label{Clark99}}
	\label{fig:2.3}
\end{wrapfigure}

Universal, international, standardized methods for seagrass directly for seagrasses
as such still should be developed. Various case studies have been performed, yet their mostly report
methods adjusted for particular areas, without evaluating standard general algorithms that could be
extrapolated towards other regions. \\ Application of the remote sensing towards seagrass mapping is
generally based on the assumption that various types of the seafloor bottom have different
characteristics of the reflectivity, which is visually expressed in distinct colors of the objects. In its
turn, reflectivity of the sediments is affected by the water optical properties and content. For
example, Stephens \cite{Stephens03}\label{Stephens03} prove that microalgae biomass and community structure affect
hyperspectral reflectance of sediments, which enable to estimate total microalgae biomass from
measurements of hyperspectral reflectance. 
Spectral measurements\index[concepts]{Remote sensing!spectral measurements} of the target objects are made
by means of the radiometers, which receive and register the amounts of energy (radiance and
irradiance) from the objects. Measuring optical properties of of the seawater allows to calculate spectra of the
objects and to discriminate them on the aerial and satellite images. Thus, various scientists report
success in spectral discrimination of submerged vegetation and other seafloor cover types on imagery
using hyperspectral optical properties of the sea water for the assessment of benthic habitats \cite{Lewis01,Louchard03,Dehouck08,Werdell03}\label{Lewis01}\label{Louchard03}\label{Dehouck08}\label{Werdell03}.
Studies of spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the different seagrass species comparing to the spectra of sand and
other seafloor cover types \cite{Vahtmae06}\label{Vahtmae06} prove that spectra of green, brown and red benthic
macroalgae differ from each other, as well as from sand and deep water reflectance spectra. These
differences are well detectable by the means of the remote sensing research methods. Comparing to
the terrestrial plants, aquatic vegetation, \textit{inter alia} seagrass, cannot be detected using red edge of the
spectrum, as these wavelengths are significantly absorbed by water \cite{Kirk94}\label{Kirk94}, as well as by
scattering and absorption by phytoplankton. Some authors \cite{Dierssen03}\label{Dierssen03} also use
spectrally based radiative transfer approach to quantitatively estimate shallow-water bathymetry and
leaf area index (\ac{LAI}) of the seagrass. The spectral reflectance in general is the
result of the spectral absorption in different bands, typical for each target object.

\begin{table}[htbp]
	\caption{Characteristics of selected ocean-colour sensors}
	\begin{center}
		\rowcolors{1}{OldLace}{FloralWhite}
		\begin{tabular}{|p{2cm}|p{20mm}|p{15mm}|p{17mm}|p{15mm}|p{15mm}|p{15mm}|}
			\hline\hline
			\textbf{Satellite Sensor:} & {\textbf{Landsat}} & {\textbf{SPOT}} & {\textbf{ENVISAT}} & {\textbf{Terra}} & {\textbf{GeoEye}} & {\textbf{Nimbus 7}} \\ \hline\hline
				\textbf{Origin} & USA & France & Europe (ESA) & USA (NASA) & USA & USA \\ \hline
				\textbf{Instrument} & \textit{TM, ETM+} & \textit{Pan, XS} & \textit{MERIS} & \textit{MODIS} & \textit{SeaWiFS} & \textit{CZCS}\\ \hline
			\textbf{Number of channels} & 8 & 6 & 15 & 36 & 8 & 5 \\ \hline
			\textbf{Wavelength coverage} & 185-1700 & 500-1750 & 412.5- 900 & 469-1640 & 412-865 & 443-750 \\ \hline
			\textbf{Ground resolution (nadir), m} & 15-pan,30(MS),60 (TIR) & 2.5, 5, 10,20 & 1.2 km/300 m & 1.0 km & 1.13 km & 825 m \\ \hline
			\textbf{Band Width (nm)} & 0.45-10.4 & 10,20 & 2.5,7.5,9,10 14,20 & 0.18,0.3,0.5,10 - 50 & 20,40 & 20,100 \\ \hline
			\textbf{Launched} & 1972, 1999 & 1986 & 2002 & 1999 & 1997 & 1978 \\ \hline
			\textbf{Recurrent period, days} & 16 & 24 & 35 & 16 & 16 & 16 \\ \hline
		\end{tabular}
	\end{center}
	\label{tab:1}
\end{table}

Spectral reflectance\index[concepts]{Seawater optical properties!spectral reflectance} of the seagrass  (Fig.\ref{fig:13}) is largely influenced by the water
depth where it is located, and is generally decreasing in values by increasing depths.
The most important spectral diapason\index[concepts]{Seawater optical properties!spectral reflectance!spectral diapason} for marine mapping of submerged vegetation
and, particularly, for seagrass, is 350-800 nm \cite{Barille09}\label{Barille09}.
Using airborne imagery for retrospective data (before 1970s) together with the most
recent imagery allows to detect changes in seagrass distribution on over different years
and to analyse dynamics of the seagrass distribution \cite{Ball09}\label{Ball09}. \\
Another important advantage of the application of the remote sensing data for mapping of submerged aquatic vegetation has
commercial nature: using remote sensing data and methods enables more low-cost and up-to-date
seagrass mapping \cite{Mumby99}\label{Mumby99}, and is especially useful for the areas where the fieldwork data
capturing is unavailable.
Seagrass meadows may reach spatial scales from several up to hundreds of metres, therefore they are
susceptible by the means of satellite imagery from remote sensors, both with moderate resolution
(e.g., \href{http://landsat.gsfc.nasa.gov/}{Landsat MSS, Landsat TM, Landsat ETM+}, \href{http://envisat.esa.int/instruments/meris/}{MERIS} from \ac{ESA}, \href{http://asterweb.jpl.nasa.gov/}{ASTER}, \href{http://modis.gsfc.nasa.gov/}{MODIS}) and high resolution as
well (e.g., \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}, \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird}, \href{http://www.spotimage.fr/}{SPOT}, \href{http://www.itres.com/products/imagers/casi550}{CASI}). The possibility of their application towards seagrass mapping varies and is limited by the technical characteristics (Table \ref{tab:1}) and resolution of these sensors  (Fig.\ref{fig:2.3}).
In the next paragraphs (\nameref{sec:2.4.1} and \nameref{sec:2.4.2}) we will briefly discuss research experience of the application of satellite imagery towards seagrass mapping, its advantages and limitations, reported by various researchers.

\subsubsection[Review of multispectral imagery...]{Review of multispectral imagery used for seagrass mapping}\label{sec:2.4.1}\index[concepts]{Remote sensing!multispectral imagery} 
Seagrass mapping using remotely sensed data from multispectral sensors is based on the classification
and discrimination of the seafloor types using their spectral characteristics in different wavebands.
Perhaps, the most known imagery, widely used in the remote sensing mapping, is received from the
group of Landsat sensors, known for its historical, pioneer role in the satellite industry.
Being the longest running satellite system, lunched in 1972, Landsat is the only source of archival
data going back to 1984 at a sufficient spatial resolution \cite{Dekker05b}\label{Dekker05b}, which makes its data
desirable for historical mapping or environmental analysis of change detection of the seagrass
landscapes.

\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.35]{Fig-13.jpg}
 	\caption{Band coverage of ASTER and Landsat channels on the e/m spectrum. Source:\cite{Kalinowski04}\label{Kalinowski04}}
	\label{fig:2.4}
\end{wrapfigure}

The \ac{Landsat TM}\index[concepts]{Remote sensing!multispectral imagery!Landsat TM} and \ac{Landsat ETM+}\index[concepts]{Remote sensing!multispectral imagery!Landsat ETM+} data, with the most recent from sensor Landsat 7 - an advanced
and multispectral scanning, launched in 1999, prove to be feasible and useful for the mapping of
submerged vegetation, such as seagrasses or coral reefs. The successful applications of the imagery
\ac{Landsat TM} towards the seagrass mapping, were reported in numerous research works \cite{Palandro03,Gullstroom06,Wabnitz08,Bierwirth93,Ferguson97,Rasib97}\label{Palandro03}\label{Gullstroom06}\label{Wabnitz08}\label{Bierwirth93}\label{Ferguson97}\label{Rasib97}.\\ The Landsat data are particularly suitable for the
case of change detection of seagrass landscapes at a decade scale, because being the main sensor
onboard the Landsat satellites, the Thematic Mapper (TM) provides the longest time series
available for change detection analysis over submerged vegetation \cite{Palandro03}\label{Palandro03}.
Another well-known multispectral sensor, \ac{SPOT}\index[concepts]{Remote sensing!multispectral imagery!SPOT} provides multispectral imagery with a spatial
resolution of 10 m, covering covers a surface area of 3600 km² (60*60 km swath), 26-day
orbital repeat cycle for nadir viewing and imagery with a spatial resolution 20 - 2.5 m
\cite{SPOT}\label{SPOT}. \ac{SPOT} imagery was used for mapping beds of \textit{P.oceanica} in the
Mediterranean Sea \cite{Pasqualini05} \label{Pasqualini05}. 
The \href{http://www.satimagingcorp.com/gallery-ikonos.html}{IKONOS}\index[concepts]{Remote sensing!multispectral imagery!IKONOS}, offering multispectral and panchromatic imagery, was the first to collect publicly
available high-resolution imagery at 1- and 4-meter resolution from \href{http://www.geoeye.com}{Geoeye}. IKONOS
imagery has been applied for the seagrass mapping due to its high resolution and accessibility.

\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.30]{Fig-14.jpg}
	\caption{Multispectral vs. hyperspectral band coverage. Source: \cite{Sabins97}\label{Sabins97}}
	\label{fig:2.5}
\end{wrapfigure}

Thus, the results of image classification in case study of shallow-water marine environment \cite{Mumby02}\label{Mumby02}
made using IKONOS, \ac{Landsat TM}, and \ac{CASI}\index{concepts}{Remote sensing!multispectral imagery!CASI}, show that in the blue part of the
spectrum, the best results are achieved by the IKONOS and \ac{CASI}, while \ac{Landsat TM} has not high
enough resolution. 
It may be caused to some extent by the loss of the radiance contrast, atmospheric
Rayleigh scattering and defects of scattering. However, comparing between CASI and IKONOS, the
same authors prove that \ac{CASI} enable to receive still more accurate results of the classification
than IKONOS \cite{Mumby02}\label{Mumby02}.
Another comparative analysis of the application of \href{http://www.itres.com/products/imagers/casi550}{CASI}, Landsat and \href{http://www.digitalglobe.com/index.php/85/QuickBird}{Quickbird} imagery \cite{Phinn08}\label{Phinn08}
 demonstrates high suitability of \ac{CASI} images for the fine-scale mapping of the seagrass
landscapes. Thus, \ac{CASI} and Quickbird-2\index[concepts]{Remote sensing!multispectral imagery!Quickbird} images enable to identify even separate seagrass species
with small width and heterogeneous nature of the seagrass patches, which could not be detected using
Landsat TM images with their 30*30m resolution.
\ac{ASTER}\index[concepts]{Remote sensing!multispectral imagery!ASTER}), launched in
1999 on board Terra sensor\index[concepts]{Remote sensing!multispectral imagery!Terra sensor}, provides high-resolution images of the Earth in 15 different bands of the
electromagnetic spectrum, ranging from visible to thermal infrared light  (Fig.\ref{fig:2.4}). It has diverse
subsystems for the \ac{VNIR} with 15 m resolution, shortwave infrared \ac{SWIR}, and
\ac{TIR} wavelength regions \cite{Kalinowski04}\label{Kalinowski04}. For each channel there is separate
on board calibration \ac{OBC} system, telescope with independent pointing and different detector technology \cite{Arai05}\label{Arai05}. 
This makes \ac{ASTER} imagery especially suitable for the application towards detailed
mapping of surface temperature, emissivity and reflectance of objects and bathymetric elevations as
well. Successful application of the \ac{ASTER} imagery towards mapping of submerged vegetation
reported, for example, by Hirose \cite{Hirose04}\label{Hirose04}. 
Other multispectral images have also been used for the
seagrass interpretation. For example, the success of using imagery from the multispectral airborne
scanner Daedalus AADS1268 is reported by Heege et al \cite{Heege03}\label{Heege03} where they aim at classification of
macrophytes in shallow waters of the Lake Constance.

\begin{wrapfigure}{r}{0.5\textwidth}
	\centering
	\includegraphics[scale=0.15]{Fig-15.jpg}
	\caption{Reflectance spectra of seagrass; Rrs(0-) - subsurface RS reflectance; Rrs b - the bottom
		reflectance. Source:\cite{Yang10}\label{Yang10}}
	\label{fig:2.6}
\end{wrapfigure}

In regard to the methods chosen for the image interpretation, the supervised classifications proves to
be the most worthy in a majority of case studies \cite{Palandro03,Peneva08}\label{Palandro03}\label{Peneva08}. While
often methods of the unsupervised classification are used as a tool for classifying submerged object
and features on the multispectral images and aerial imagery \cite{Fletcher09}\label{Fletcher09}, it still does not
provide common algorithms that can be applied to other images and regions. Thus, the strength of the
optical signal coming from the various types of the seafloor is strongly influenced by the effects of the
water column, its depth and chemical properties. Yet methods of the unsupervised classification
overleap mixed spectral effects of the water column which shift the real values of the spectra, with the
pure reflectance from the benthos; as a result, it may cause significant errors \cite{Dierssen03}\label{Dierssen03}.
Therefore, for the accurate assessments of various seafloor cover types, water
column depth and its optical properties, methods of supervised classification should be preferably
used as being more suitable in classification and interpretations of imagery.
While this and other studies \cite{Phinn08}\label{Phinn08} demonstrated the advantages and success of the
application of multispectral imagery for the spectral discrimination of the seafloor cover types and
mapping the submerged landscapes on the basis of pixels classification, the application of data from
hyperspectral sensors has better potential due to their higher resolution (Fig.\ref{fig:2.5}).

\subsubsection[Review of hyperspectral imagery...]{Review of hyperspectral imagery used for seagrass mapping}\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery}
The application of the hyperspectral sensors is most effective and provide more accurate classification
results  (Fig.\ref{fig:2.5}), due to their higher spectral resolution \cite{Bharathi03}\label{Bharathi03} with interval
narrows to 10 nanometers, while broadband sensors are limited to the spectral width of ca
150 nm  (Fig.\ref{fig:2.3}). 
Hyperspectral imagery is acquired through the simultaneous acquisition of images in many
narrow, contiguous spectral bands from hyperspectral scanners (mostly) cover the 400-
to 2500-nm spectral bands \cite{Schmidt03}\label{Schmidt03}.
Perhaps, the most advantageous and general characteristics of hyperspectral imagery is its
high spectral resolution, desirable for the case of seagrass monitoring. 
The most suitable
scanner fit for detailed seagrass mapping would cover bands of 550-750 nm and have a spectrum
resolution of 5-15 nm \cite{Fyfe04}\label{Fyfe04}, which exactly characterizes typical airborne and
hyperspectral satellite scanners. Important features may be detected in the narrow
wavelengths of hyperspectral imagery, while this information can be lost in the broader
wavelengths of other sensors. With its 126 spectral bands, \ac{HyMap} imagery enables to
distinguish features of interest, i.e. seagrass types \cite{Peneva07}\label{Peneva07}, which is the major
advantage of the of hyperspectral data for mapping landscapes of the seagrasses.
While comparing multispectral imagery with airborne hyperspectral, the last one showed higher
overall accuracies \cite{Phinn08}\label{Phinn08}. Several, regular and narrow (10 nm) spectral bands\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery!spectral bands}, specific for
hyperspectral imagery, are strong tool enabling to detect even slight and subtle differences in the
spectral reflectance between various seafloor types, e.g. different seagrass species at diverse depths,
algae, corals, dark-coloured sands or other types of sediments \cite{Hochberg03a}\label{Hochberg03a}. 
Therefore, there is great potential of the application of hyperspectral remote sensing imagery towards seagrass
mapping at species level, as long as they are distinguishable spectrally (Fig.\ref{fig:2.6}) which, for example, has been
tested in the case study of Australian marine ecosystems by Fyfe \cite{Fyfe04}\label{Fyfe04}. 
Application of various classifications methods, \textit{inter alia} maximum likelihood, minimum distance and means, towards
hyperspectral imagery \cite{Peneva08}\label{Peneva08}, combined with fieldwork measurements ensures accurate
mapping results with the maximum likelihood methods producing the best results. 
Therefore, accurate mapping of the seagrass landscapes and other seafloor types using remote sensing
approaches requires application of high-spatial resolution (higher than 5 m) or hyperspectral imagery.
Comparative analysis of the application of hyperspectral\label{sec:2.4.2}\index[concepts]{Remote sensing!hyperspectral imagery!application} and multispectral imagery towards the
seafloor types classification \cite{Hochberg03a}\label{Hochberg03a} demonstrates that coral, seagrasses and
sand very well distinguishable in their spectra with an overall classification accuracy of 98 \%.
However, the use of data from various sensors, both hyperspectral and multispectral, is possible and
reasonable, as soon as it meets the research specific objective. 
Thus, the use of multispectral imagery
with high spatial resolution is preferable to using hyperspectral medium resolution data in case of
mapping benthic vegetation in areas where the spatial heterogeneity is very high \cite{Vahtmae07}\label{Vahtmae07}.

%%%%%%%%%%%%%%%%%  Section  3  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and methods}\label{sec:3}
\counterwithin{figure}{section}
\counterwithin{table}{section}

In the section \nameref{sec:2} we briefly discussed main existing cartographic methods and tools for seagrass mapping. In the current  research we try to combine various techniques from mentioned above for the assessment of the seagrass distribution in the Mediterranean environment, for the case study of Crete. We mentioned in the \nameref{sec:0} the software that we used for logically separated research sections, which required various approach. Thus, for example,  spectral analysis and assessment of spectral signatures is technically based on \ac{WASI}, statistical analysis of series of experimental raw data has been made by means of statistical software, such as \href{http://gretl.sourceforge.net/}{Gretl} and \href{http://www.spss.com/}{SPSS}; mapping has been made using  \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS}; images processing and classification were made using \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine} software by means of Google Earth and  \href{http://landsat.gsfc.nasa.gov/}{Landsat} imagery.
\href{http://www.esri.com/software/arcgis/index.html}{ArcGIS 10.0} software is used for the spatial analysis, general mapping and cartographic layout
presentation; the raster processing techniques are applied for the detection of seagrass spatial
distribution using supervised classification.
The research data include Google Earth aerial images and scenes from the \ac{Landsat TM} and \ac{Landsat ETM+}
covering research period of 10 years in the same year time, taken from \ac{USGS}, \ac{GloVis}, \ac{NOAA} and \href{http://glcfapp.glcf.umd.edu:8080/esdi/index.jsp}{The Earth Science Data Interface}. 
The satellite imagery provides
vital information of the most recent changes in \textit{P.oceanica} within the coastal areas, as well as the
condition (poor or destroyed).
The raster processing includes making mosaic-like covering for the whole research area. The Google
Earth images are most appropriate for the detailed mapping than satellite than \href{http://landsat.gsfc.nasa.gov/}{Landsat} or \href{http://asterweb.jpl.nasa.gov/}{ASTER} enabling to produce accurate maps with correct results. Therefore, the main set of images for the
current work is Google Earth aerial images. The  \href{http://landsat.gsfc.nasa.gov/}{Landsat} images are used for the general overview.
Besides aerial and satellite photographs, data acquired during the fieldwork (three weeks in September - October) are
necessary addition to the mapping helping to solve problems of interpretation during the images
classification \cite{Pasqualini98a}\label{Pasqualini98a}. Therefore, this work includes sampling of the \textit{in-situ}
measurements of the seagrass distribution.
The sampling stations were located in two candidate places on the northern (Ligaria) and southern
(Xerocampos) parts of Crete island, as these regions are well suitable for the seagrass, due to the
annual mean water temperatures and geological factors, i.e. seafloor conditions and sediments. \\ The
field campaign has been carried out during the September-October period 2010. 
The information about the location of the seagrass (mostly represented by the 	extit{P.oceanica} species) is useful for the
understanding of the relationship between the spatial distribution of the seagrass and the environment
of the selected areas of Cretan beaches. The results of the videographic measurements are used for the
seafloor types detection, because the objects represented on the photos can be well distinguished and
classified according to the following well-known characteristics \cite{Butler87}\label{Butler87}: size (yet some
measurements are necessary for the similar-looking objects); shape (the general form is most reliable
evidence for identification); colour (common and reliable object indicator); texture (when changes in
tone are too small to be distinguishable, texture may assist identification, e.g., stippled, granular,
rough, smooth, etc.); associated features (those usually found near other objects, e.g., rocks and soil).

\subsection{Study area}
General area: Island of Crete, Greece (Fig.\ref{fig:3.1}).\\
The study area of the current MSc research is located in the shallow areas of the Ligaria beach, Agia
Pelagia and Xerocampos, Crete island, Greece (Fig.\ref{fig:3.2}). These shelf areas have maximal depth of four meters.
Seagrass sampling has been performed at two stations at a depth of 4 meters, in the following selected
areas:
\begin{itemize}
	\item [1.] Ligaria beach (Agia Pelagia district), $36\,^{\circ}20\textquoteright N, 22\,^{\circ}59\textquoteright E$ 
	\item [2.] Xerokampos, $35\,^{\circ}12\textquoteright N, 26\,^{\circ}18\textquoteright E$
	\item [3.] Agia Pelagia, $36\,^{\circ}20\textquoteright N, 22\,^{\circ}59\textquoteright E$.
\end{itemize}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.10]{Fig-16.jpg}
	\caption{Study area: Crete island}
	\label{fig:3.1}
\end{figure}

\subsection{Fieldwork data collection}\index[concepts]{Data!collection}
Seagrass sampling has been performed at two research stations at Crete Island - Ligaria beach
($36\,^{\circ}20\textquoteright N, 22\,^{\circ}59\textquoteright E$) and Xerocampos ($35\,^{\circ}12\textquoteright N, 26\,^{\circ}18\textquoteright E$), at depths lesser than 4 meters.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.25]{Fig-17.jpg}
	\caption{Study area: Ligaria beach, Crete island}
	\label{fig:3.2}
\end{figure}

The Ligaria Beach is a narrow, sandy and pebble beach (Fig.\ref{fig:3.2}), located ca 15 km north-west from Heraklion.
The \textit{in-situ} measurements were conducted during the fieldwork in the period 21.09.2010-11.10.2010.
The fieldwork on seagrass monitoring included visual estimations and photo- and video footage of the
above-ground seagrass patches, sediment seafloor cover types, species compositions, water depth and
geographic locations recorded using \ac{GPS}.

\subsubsection{Fieldwork equipment}
The research materials and equipment were provided by \href{http://147.52.130.151/MFIK_en/}{the Natural History Museum of Crete}, the \href{http://www.uoc.gr/Department/index.html}{University of Crete} and the ITC, and included the following items (Fig.\ref{fig:3.3}): 1) three iPAQs (Fig.\ref{fig:3.3a}); 2)
three GPS (Fig.\ref{fig:3.3c}); 3) Three underwater video cameras, \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), suitable for
photographing up to up to 33 foot depths and high-resolution 12.0-Mpixel image sensor; 4)
Markers and cords for the depths measurements; 5) Waterproof plastic Otterbox  (Fig.\ref{fig:3.3b})for keeping the
iPAQ dry; 6) \ac{SCUBA} Diving equipment, taken from the Ligaria Diving Center; 7) Boat (Fig.\ref{fig:3.4});

\begin{figure}[H]
	\centering
	\subfloat [iPAQ, HP]{\label{fig:3.3a}\includegraphics[width=0.3\textwidth]{Fig-19a.jpg}}
	\subfloat [Waterproof Otterbox for iPAQ]{\label{fig:3.3b}\includegraphics[width=0.3\textwidth]{Fig-19b.jpg}}
	\subfloat [GPS]{\label{fig:3.3c}\includegraphics[width=0.15\textwidth]{Fig-19c.jpg}}
	\subfloat [Olympus waterproof camera]{\label{fig:3.3d}\includegraphics[width=0.25\textwidth]{Fig-19d.jpg}}
	\caption{Fieldwork equipment}
	\label{fig:3.3}
\end{figure}

A \ac{GPS} and \ac{iPAQ} have been used for detection of the geodetic coordinates and keeping the tracklogs
along the boat route for \ac{GIS} project.

\subsubsection{Sampling design}
The sampling design of the fieldwork was aimed at surveying of the spatial distribution of the
meadows of \textit{P.oceanica}, and spatial pattern of the seagrass meadows consisted from separate
patches. The fieldwork included several routes of the boat in the Ligaria beach sampling site, nine
routes in total, in the directions parallel to the coastline, ca 180-200 m long each one, thus enabling the
course plot to cover the area of growing seagrass: shelf areas not deeper than four meters.
The measurements of the seafloor cover types have been made using underwater video cameras
Olympus ST 8010, mounted under the boat to capture video footage and imagery. The data records
were made along each path using these cameras. A videographic approach, tested previously \cite{Norris97}\label{Norris97},
 was applied during the fieldwork on Crete island, for collecting information on benthic cover types and
distribution of the seagrass patches from photo transects, in order to use for the calibration of mapping approaches.
Three underwater video cameras, located on the bottom of the boat, provided videometric measurements of the
seafloor during the track, and resulted in a series of consequent overlapping images of the sea bottom under the
boat path. The general locations of the sampling sites and routes were selected on the basis of the visual examination
of the seagrass beds during snorkeling and \ac{SCUBA} diving, recommendations from the Greek collaborators of the \textit{Natural History Museum of Crete}, and available maps covering the research area. To ensure even and objective selection of sampling sites
we used transect sampling method, i.e. photographs were taken along the research path.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.70]{Fig-18.jpg}
	\caption{Boat used for the fieldwork measurements on Crete Island. 2010}
	\label{fig:3.4}
\end{figure}

Transect method is a common sampling technique in studies of the seagrass monitoring \cite{Shortis07}\label{Shortis07},
enabling the occurrences of the seagrass meadows to be recorded and counted in a systematic and
accurate way, detect spatial distribution of the single seagrass mattes to be properly identified,
without location bias. 
The videographic measurements and photos, enabling to detect types of the
seafloor, were captured by means of three underwater digital video cameras \ac{Olympus ST} 8000  (Fig.\ref{fig:3.3d}), at
five minutes interval along the tracklog path (Fig.\ref{fig:3.6}).
research sampling design also included measurements of depths (Fig. \ref{fig:A.4}), because bathymetry is one of the
most determining factors for the seagrass locations. Measurements of depths were performed during
the fieldwork using cord, \ac{iPAQ}, and markers in order to assure that the videometric measurements
are taken at depths not more than four meters, in the shelf area.
As a result, nine transects were established of one m wide and 20 m long track, to cover seagrass beds
with videographic measurements. The \ac{GPS} allowed to capture measurement locations on the \ac{iPAQ},
encapsulated in a plastic waterproof Otterbox (Fig \ref{fig:3.3b}). The camera were adjusted horizontally by a leveler and
 mounted under the bottom of the boat to capture video footage and imagery (Fig.\ref{fig:3.5}). The data
were taken at proper weather conditions: sunny, serene and cloud-free days with glassy sea state.

\begin{figure}[H]
	\centering
	\subfloat [view from below]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20a.jpg}}
	\subfloat [view \textit{en-face}]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20b.jpg}}
	\subfloat [view from above]{\label{fig:22a}\includegraphics[width=0.3\textwidth]{Fig-20c.jpg}}
	\caption{Scheme of placements of the Olimpus cameras during the measurements}
	\label{fig:3.5}
\end{figure}
The locations of the route were randomly selected in the areas of the Ligaria beach, to ensure most dense
coverage of the seagrass meadows in the research area. The underwater measurements of the seagrass
coverage were carried out by taking video footage and photos of ca 0.5 m2 size each (Fig \ref{fig:A.7}).
The results of the underwater videometric measurements include series of digital images helping to
classify seafloor cover types and seagrass meadows, according to the differences in the structure,
colour, texture and shapes of the depicted objects.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.40]{Fig-21.jpg}
	\caption{Locations of the video measurements and GPS tracklogs, Ligaria}
	\label{fig:3.6}
\end{figure}

Along the Ligaria beach there are several types of the seagrass landscapes, namely, seagrass meadows
continuously covering vast areas (Fig.\ref{fig:A.3} (a)), aggregated seagrass patches, represented by separate mattes with short irregular channels between them (Fig.\ref{fig:A.3} (b)) and isolated seagrass patches, or mattes,
which located separately from broad meadows (Fig.\ref{fig:A.3} (c)).
The results of the underwater videometric measurements of the \ac{Olympus ST} cameras made during the
ship route include nine total tracklog routes in the selected research area, including series of
consequent images, completely covering the area under the boat path. The received data contain
information on seagrass presence within the study area, distribution of seagrass \textit{P.oceanica} meadows
and nature of the seafloor cover types: rocks, sandy, mixed (Fig. \ref{fig:A5}). Seagrass species on the Ligaria and Agia
Pelagia beach consist of 	extit{P.oceanica}. The types of sediment on the Ligaria vary from coarse sand,
\textit{P.oceanica} patches, sand, mud, rock gravel and fine sand (Fig.\ref{fig:A.6}).

\subsection{Data pre-processing}\index[concepts]{Data!pre-processing}
A collection of visible spectra of two seafloor cover types - \textit{P.oceanica} and carbonate sand - consist of multiple measurement sets of \textit{P.oceanica} made by means of RAMSES hyperspectral radiometer: 1-350 for 15th October (fragment on Tab. \ref{tab:A.8}), 1-445 for 14th October for 	\textit{P.oceanica}; 1-106 for water without sediments, measured in aquarium tank, 1-27 for seawater with sediments measured in aquarium tank; 1-75 for carbonate sand (Tab. \ref{tab:17}). All these data have been analyzed and statistically tested. We measured mean, extreme values (min-max), median and statistical quartiles, in order to visualize distribution of the values at various spectral wavelengths. Carbonate sand was measured at wavelengths 402-750 nm, while \textit{P.oceanica} - at the interval 318-951 nm. We visualized the behavior of different spectra on relevant graphs (Fig. \ref{fig:4.4}, \ref{fig:A.35}, \ref{fig:A.28}). 
The available collected data were tested for the spectral variability and separability under varying conditions of different environmental
constituents (e.g. depth, water content, sun angle), in order to determine the
potential that it may have on the approaches for further images processing and classification. Behavior of the spectra of \textit{P.oceanica} have been tested using datasets for various depths: 0.5, 1.5 and 2.5 meters (Tab. \ref{tab:8}, Tab. \ref{tab:A.7}). The raw
initial data of measurements of the spectral reflectance have been preprocessed and statistically analyzed thereafter with graphs visualizing mean spectra and Q1-Q3 interval, instead of the series of single observations which is illustrated, for example, on Figures \ref{fig:27}, \ref{fig:4.10}, \ref{fig:28}). These data
have been measured using different step of the wavebands: some measurements
were made with 3 nm interval, while others – using 4 and 1 nm step.
Therefore, these data had to be interpolated (Fig. \ref{fig:A.10}) and standardized to one
format, which is values of spectral reflectance with one nm step. For the
interpolation we used script written on Python programming language that allowed receiving more 
detailed data by interpolating them from 3 and 4 nm step up to 2 nm (Fig.\ref{fig:A.8}).
The interpolated data contained spectral measurements of the seagrass 	\textit{P.oceanica} (measured at Agia
Pelagia beach, Crete), sand (measured at Agia Pelagia beach, Crete), silt and default artificial
spectrum of constant albedo at \ac{WASI} (Fig.\ref{fig:A.9}). 
We included measurements of the Thalassia seagrass
(measured at Southern Chinese Sea by C.Yang \cite{Yang10} \label{Yang10}) for the comparison of the spectral reflectance of
different seagrass species under various environmental conditions.
The most suitable wavebands for the seagrass monitoring usually lay between 400 and 700 nm,
which can be concluded by the visual examination, comparison and analysis of the different spectra of
the seagrasses. Therefore, we chosen the spectra 400-750 as the most appropriate range for
further research experiment. The results of the linear interpolation (Fig.\ref{fig:A.35}) demonstrate values of the
sand spectral reflectance with 1 nm interval covering the wavelength diapason of 400-750 nm.\\
The initial measured data were stored in raw-oriented format, so that re-formatting them into the
column-based layout was done using “transpose” command in Open Office or Excel (Fig. \ref{fig:A.11}). The next step included
calculation of the median, mean, quartiles and other statistical values at every data set (see \nameref{app}, Fig. \ref{fig:A.25}, Tab. \ref{tab:17}).
After the preliminary analysis, the measured data were visualized using Gnuplot program, which
enables fine plotting of various datasets together: Fig \ref{fig:4.4}, Fig.\ref{fig:27}, Fig.\ref{fig:28}. The most acceptable method of interpolation was B\'ezier curve (Fig. \ref{fig:A.18}, Fig. \ref{fig:A.16}),
as it has trend-friendly graph better showing the general behavior of the curves at different wavebands
comparing to splines (see \nameref{app} for more results). 
Therefore, after several experiments with
various interpolation techniques (Fig.\ref{fig:A.15}, Fig.\ref{fig:A.23}), we have chosen B\'ezier curves interpolation (Fig.\ref{fig:A.18}) which contains convex
hulls made on its control points, and therefore is best suitable for our case: analysis of optical
properties of seawater.

\subsection[Google Earth aerial imagery]{Google Earth aerial imagery for the seagrass mapping}
Apart from the satellite imagery, the aerial photographs from the Google Earth provide
a powerful tool for seagrass mapping, because they are important, reliable, detailed
and up-to-date source of imagery.
Perhaps, the clearest advantage of the Google
Earth imagery is its high resolution (15 m in land areas and lower in the oceans).
Obtained from the airborne platforms, Google Earth images have general spatial
resolution of several meters (though varying in different areas), which allows very
detailed habitat and seafloor types discrimination, comparing with images
acquired from the space-borne satellite platforms.\\
The spatial coverage of the Google imagery is lesser comparing to spaceborne images, but this can be solved as
well: while in general providing smaller area coverage than satellites images, the Google Earth
images can be stitched to the composite maps of the acceptable spatial extent, using Google grabbing
process which allows multiple overlapping of single images over the flight paths, and generates
mosaics.

%%%%%%%%%%%%%%%%%  Section  4  %%%%%%%%%%%%%%%%%%%%%%%
\section{Results}\label{sec:4}
\markright{IV. Results}
\counterwithin{figure}{section}
\counterwithin{table}{section}

\subsection{Review of the collected data}
The remote sensing imagery was collected for the Crete island area covering the research period and
area, enabling the field data to be used for the calibration and validation. The available aerial and
satellite imagery are commonly used for the mapping of seagrass landscapes and their application is
proven by various research papers. The imagery includes satellite multi-spectral imagery (Landsat-TM,
ETM+) and aerial imagery from the Google Earth.
The overview of the collected data enables to summarize their following types:
\begin{itemize}
	\item[] 1. Optical spectra of \textit{P.oceanica}, carbonate sand, seawater with sediments and seawater measured in aquarium tank, without sediments, at different environmental conditions (e.g., Fig. \ref{fig:A.23})
	\item[]2. Aerial imagery from the \href{http://www.google.com/earth/index.html}{Google Earth}
	\item[]3. Satellite images from various open sources (e.g. Landsat, Tab \ref{tab:20}) (previews: Fig. \ref{fig:A.37})
	\item[]4. Results of underwater videometric measurements of the \ac{Olympus ST} cameras made during the ship route
\end{itemize}
The available satellite imagery (Fig.\ref{fig:A.38}) and aerial images are read into the ArcGIS project.
The available broadband and hyperspectral remote sensing data are used for the mapping of the
seagrass in shelf areas no deeper than 4.0 meters, for the environmental monitoring in order to detect
the dynamics in changes of the seagrass distribution along Crete during the past 10 years using
different satellite images for 2000-2010.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.40]{UML_Dia.jpg}
	\caption{Work flow for the data acquisition, \href{http://live.gnome.org/Dia}{Dia} display}
	\label{fig:24}
\end{figure}

The spatial resolution of \ac{Landsat ETM+} image is 30 m in the visible and near infrared bands (bands
1-5 and 7); the spatial resolution of \ac{ASTER} 15 m for the visible and near-infrared bands. IKONOS
acquires data in 3 visible channels and \ac{NIR}, with spatial resolution of 1-4 meters. \ac{ASTER} and
IKONOS images are suggested to be included as soon as available, in addition to the Landsat. The
research includes fieldwork \textit{in-situ} measurements of the seagrass distribution along the northern and
south-eastern coasts of Crete in chosen locations.\\
The optical measurements of the irradiance and radiance of the sea water and bottom cover types of
the seafloor have been received by the means of the optical sensors \ac{Trios-RAMSES} Hyperspectral
UVA/UVB Irradiance Sensor and \ac{Trios-RAMSES}-ARC Hyperspectral UV-VIS Radiance Sensor (Fig. \ref{fig:2.1}), both adjusted for the measurements of the irradiance and radiance (see \nameref{app}: Tab. \ref{tab:A.1}, Tab. \ref{tab:A.2}). \\
The highest measured values are located in the diapason of 410-730 nm for the water irradiance (Fig. \ref{fig:A.18}) and 430-650 for the water radiance (Fig.\ref{fig:A.16}). 
Afterwards, the measured values of the radiance and irradiance, respectively, have been used for the computation of the spectral reflectance
properties of the sea water and bottom cover types (Fig.\ref{fig:A.15}). The spectral range of radiance
cover diapason of 320-950 nm, and irradiance measurements are covered in the interval of 280-500
nm, which is suitable for characteristics of seagrass reflectance.
Different curves on the reflectance, radiance and irradiance graphs for example, on Fig.\ref{fig:A.12}, or \ref{fig:A.14}, enlarged) represent several series of
the measurements. The values of the spectral reflectance are received from the
computations of these values using mathematical formulae. The graphs shown on Fig. \ref{fig:4.2}, Fig. \ref{fig:A.23}, Fig. \ref{fig:A.24}, Fig. \ref{fig:A.17} display values of the radiance and irradiance of the sea water in Agia Pelagia, with and without sediments and suspended particles, respectively. Graph on Fig \ref{fig:A.13} displays statistical analysis of the measured sets of observations, i.e. shows the mid-spread of the statistical quartiles (Q1-Q3), mean and extreme values.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{GNU-17.jpg}
	\caption{Optical properties of the sea water with sediments, measured in aquarium tank. Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}				{Gnuplot}}\label{fig:4.2}
\end{figure}
\pagebreak

%
\subsection{Analysis of spectral signatures}
The distinguishing spectral signatures for various seafloor types (e.g. seagrass species, coral reefs,
various types of sand, mud, other sediments, (Fig.\ref{fig:4.4})) exist in well-defined and narrow (10–20 nm)
wavelength ranges. 
The values of their spectral reflectance are accepted as constant. The results of
spectral measurements enable to analyze, whether P.Oceanica is spectrally distinct from other sea
floor types with changing environmental conditions, using the differences in their spectral signatures
on the graphs in a \ac{WASI} software. \\The \ac{WASI}, a
software tool for analyzing and simulating the most common types of spectra \cite{Gege05}\label{Gege05}, is highly
suitable for the seagrass spectral analysis.
There are several environmental characteristics, included in \ac{WASI} interface (Fig.\ref{fig:4.7}), which influence the
results of water spectral reflectance, e.g. different bottom depths, concentration of suspended particles
in water column, water temperature, sun angle, concentration of Gelbstoff (coloured dissolved organic
matter), concentration of phytoplankton, aerosol scattering, exponent of backscattering by small
particle \cite{Gege04}\label{Gege04}. The backward-scattering coefficient (bb), also included in WASI, is a
fundamental optical property which plays a central role in the ocean-color remote sensing, providing
the remotely sensed optical signal, as well as suspended particle distributions, water clarity, and
underwater visibility \cite{Maffione97}\label{Maffione97}. WASI enables simulation of backscattering of pure
water, large and small particles. The values of all these parameters can be redacted and changed
manually. However, the most important, major factors affecting the \textit{in-situ} conditions are water depth (Fig.\ref{fig:4.9})
and cover fraction of the seafloor types: \textit{P.oceanica} and carbonate sand.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.27]{GNU-18.jpg}
	\caption{Multiplot showing spectral reflectance of the seawater with sediments, measured in aquarium tank, Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis.}
	\label{fig:4.4}
\end{figure}

Seagrass \textit{P.oceanica} can be mapped using remotely acquired spectral imagery, but only if it has distinctive
reflectance signatures at different depths. Therefore, the depths of the shelf area are the first variable
condition for this research question. The depths values chosen for the current research are lesser than
3.5 meters, covering shelf zone, and providing the best environmental conditions for the seagrass
\textit{P.oceanica}: 0.5, 2.0 and 3.5 meters with an interval of 1.5 meters.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.3]{GNU-20-new.jpg}
	\caption{Statistical comparison of spectral reflectance of \textit{P.oceanica} and carbonate sand. Selected measurements sets. \href{http://www.gnuplot.info/}{Gnuplot} graph}
	\label{fig:28}
\end{figure}

The 3.5 m depth limit was chosen based on the analysis of the separability of seagrass reflectance signatures, received by the means of
previous \textit{in-situ} measurements (year 2009) of the radiance and irradiance of water in Agia Pelagia
bay, which indicated that \textit{P.oceanica} seagrass could be only well separated within depths of 3.5 m.
A statistical analysis of WASI-simulated spectral reflectance has been used in order to answer the first
research question: whether the 	extit{P.oceanica} spectra is spectrally distinct at varying environmental \textit{in-situ}
conditions, and if \textit{P.oceanica} remains spectrally distinct with the increasing water column depth.
To answer this research question, different seafloor cover types are discriminated using data of the
broadband remote sensing.

\begin{table}[htbp]
	\rowcolors{1}{LightSteelBlue}{Honeydew}
	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, with average values (for sets 1 - 350). Generalisation up to step 20 nm}\label{tab:4.1}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline
	Wavelength, nm & mean & Q1 & min & Q3 & max & median \\ \hline
		\textbf{318.23354} & 0.030467 & 0.038907 & 0.029764 & 0.069762 & 0.116100 & 0.048984 \\ \hline
		\textbf{338.29900} & 0.033560 & 0.043030 & 0.032847 & 0.076252 & 0.128290 & 0.054334 \\ \hline
		\textbf{358.37932} & 0.039897 & 0.051320 & 0.039131 & 0.091143 & 0.154409 & 0.066354 \\ \hline
		\textbf{378.47247} & 0.045786 & 0.058849 & 0.045039 & 0.105993 & 0.177150 & 0.078339 \\ \hline
		\textbf{398.57640} & 0.052346 & 0.066859 & 0.051625 & 0.122328 & 0.201038 & 0.091905 \\ \hline
		\textbf{418.68905} & 0.058942 & 0.074842 & 0.058240 & 0.138984 & 0.226028 & 0.106207 \\ \hline
		\textbf{438.80838} & 0.070690 & 0.089761 & 0.069999 & 0.167825 & 0.267712 & 0.130580 \\ \hline
		\textbf{458.93235} & 0.080107 & 0.102089 & 0.079392 & 0.191424 & 0.304933 & 0.149762 \\ \hline
		\textbf{479.05891} & 0.086278 & 0.110348 & 0.085546 & 0.207940 & 0.332934 & 0.162881 \\ \hline
		\textbf{499.18600} & 0.098498 & 0.126592 & 0.097689 & 0.239161 & 0.380603 & 0.188192 \\ \hline
		\textbf{519.31159} & 0.110222 & 0.143976 & 0.109022 & 0.265948 & 0.421370 & 0.210237 \\ \hline
		\textbf{539.43363} & 0.125694 & 0.166156 & 0.124310 & 0.302678 & 0.475842 & 0.240492 \\ \hline
		\textbf{559.55006} & 0.131844 & 0.175865 & 0.130257 & 0.319385 & 0.503907 & 0.253726 \\ \hline
		\textbf{579.65885} & 0.121915 & 0.166084 & 0.120186 & 0.302078 & 0.485661 & 0.237873 \\ \hline
		\textbf{599.75794} & 0.077236 & 0.111113 & 0.075669 & 0.200557 & 0.343876 & 0.154016 \\ \hline
		\textbf{619.84529} & 0.057038 & 0.085386 & 0.055588 & 0.154781 & 0.281737 & 0.117003 \\ \hline
		\textbf{639.91885} & 0.050841 & 0.077438 & 0.049432 & 0.141810 & 0.265181 & 0.106188 \\ \hline
		\textbf{659.97657} & 0.038118 & 0.059971 & 0.036997 & 0.110774 & 0.223366 & 0.082411 \\ \hline
		\textbf{680.01641} & 0.032892 & 0.052265 & 0.032012 & 0.095789 & 0.198389 & 0.071686 \\ \hline
		\textbf{700.03633} & 0.034756 & 0.058426 & 0.033643 & 0.101504 & 0.215413 & 0.075789 \\ \hline
		\textbf{720.03426} & 0.026287 & 0.047084 & 0.025162 & 0.082940 & 0.193535 & 0.061184 \\ \hline
		\textbf{740.00817} & 0.009740 & 0.022355 & 0.010145 & 0.042861 & 0.132244 & 0.029414 \\ \hline
		\textbf{759.95601} & 0.008651 & 0.020234 & 0.009191 & 0.039371 & 0.128165 & 0.026558 \\ \hline
		\textbf{779.87573} & 0.006019 & 0.014157 & 0.006397 & 0.028320 & 0.100314 & 0.019322 \\ \hline
		\textbf{799.76528} & 0.008607 & 0.020030 & 0.009050 & 0.037323 & 0.118123 & 0.026535 \\ \hline
		\textbf{819.62263} & 0.008777 & 0.020295 & 0.009288 & 0.037598 & 0.118055 & 0.026799 \\ \hline
		\textbf{839.44571} & 0.002562 & 0.008010 & 0.002864 & 0.017472 & 0.076515 & 0.011432 \\ \hline
		\textbf{859.23249} & 0.002653 & 0.006921 & 0.002897 & 0.016121 & 0.070377 & 0.009889 \\ \hline
		\textbf{878.98091} & 0.002327 & 0.008080 & 0.002821 & 0.018625 & 0.082462 & 0.011617 \\ \hline
		\textbf{898.68893} & 0.004077 & 0.013098 & 0.004709 & 0.031217 & 0.101146 & 0.019767 \\ \hline
		\textbf{918.35451} & 0.006215 & 0.020162 & 0.007229 & 0.045213 & 0.199711 & 0.029950 \\ \hline
		\textbf{937.97559} & 0.005783 & 0.026132 & 0.007307 & 0.065475 & 0.195072 & 0.040848 \\ \hline
		\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
	\end{tabular}
	\end{center}
\end{table}

The results enable to study reflectance properties of the seagrass and other
seafloor types. Application of the optical radiative transfer model WASI is suggested to simulate
remote sensing sensors (\ac{MODIS}, \ac{ASTER}, \ac{MERIS} and \ac{SeaWiFS}), (Fig.\ref{fig:4.6}).

\begin{figure}[h]
	\centering
	\subfloat [CZCS]{\label{fig:29a}\includegraphics[width=0.4\textwidth]{Fig-35a.jpg}}
	\subfloat [MERIS]{\label{fig:29b}\includegraphics[width=0.4\textwidth]{Fig-35b.jpg}}\\
	\subfloat [MODIS]{\label{fig:29c}\includegraphics[width=0.4\textwidth]{Fig-35c.jpg}}
	\subfloat [SeaWiFS]{\label{fig:29d}\includegraphics[width=0.4\textwidth]{Fig-35d.jpg}}
	\caption{Simulated remote sensing reflectance of \textit{P.oceanica} at various sensors}
	\label{fig:4.6}
\end{figure}

In order to focus on the factors of primary importance, other and less influencing factors are excluded,
i.e. sun angle, concentration of the suspected particles in the water column, content of gelbstoff, etc.
For these factors default values of WASI are accepted.
Under normal conditions by independent water colour sampling, values of the remote-sensing
reflectance can vary by 12–24 \% \cite{Toole00}\label{Toole00}, and these variations in the radiometric
determinations are mainly caused by the variety of the environmental factors.
Different factors influence colour and spectral reflectance of water, among which different bottom
depths, concentration of suspended particles in water column, water temperature, sun angle,
concentration of Gelbstoff (coloured dissolved organic matter), concentration of phytoplankton,
aerosol scattering, exponent of backscattering by small particles, cloudiness, viewing geometry and
wind speed (which is, however, not the major source of uncertainty). All these environmental
components increase the absorption and scattering of light which, in its turn, results in a complex
relationship between their concentrations and the radiance of water that finally influence its spectral
reflectance.\\
However, although there are a variety of environmental factors that contribute to the spectral
reflectance, the most important ones from the mentioned above are the water column (depth) and seafloor fraction.
It is because spectra of the seagrass \textit{P.oceanica} vary qualitatively over the depths interval of 0.5–4 m,
and secondly, the content and cover fraction of the seafloor have the most distinctive effect on the
spectral reflectance of the water. \\
Therefore, in the current research we focus on these two major factors, and study the response of the
water reflectance towards changing conditions of the water column depth and seafloor bottom cover
fraction (seagrass \textit{P.oceanica}, carbonate sand and silt).\\
The analysis of the spectral signatures of the seagrass \textit{P.oceanica} and sand clearly shows (Fig.\ref{fig:27}) that seagrass
has spectral reflectance much lesser than that of a carbonate sand, in general not increasing values of
10 \% reflectance in spectra of 500-600 nm, while sand has spectral reflectance approaching 33\% in its
highest values.

\subsubsection{Water color spectral simulation using WASI}\index[concepts]{Data!processing}
The main aim of this part of research is to clarify if the bottom reflectance of the different seafloor types including patches of
the seagrass \textit{P.oceanica} meadows, silt and carbonate sand differ and can be clearly discriminated while mapping. A study is
based on three seafloor types containing silt, carbonate sand and seagrass, as well as mixed
types, where the spectral signatures were examined. WASI software\index[concepts]{Seawater optical properties!RTM!WASI} (Fig.\ref{fig:4.7}) is used to simulate
spectral reflectance and colour discrimination of water, affected by presence of \textit{P.oceanica} and other factors, under various
environmental conditions which influence its colour.

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.30]{Fig-30.jpg}
	\caption{Main interface of WASI water color spectral simulation: \\ Spectral reflectance of \textit{P.oceanica} and silt at 0.5-4.0m}
	\label{fig:4.7}
\end{figure}

The data provided with the model was determined at freshwater of the Boddensee \cite{Gege05}\label{Gege05}, 
yet the model was adjusted for the marine environment, so that its parameters (concentration of chlorophyll, concentration of small particles,
yellow substance, etc) now perfectly simulate the Mediterranean water conditions.
The remote sensing reflectance has been compared under the conditions of different water depths and
cover fraction of the seafloor, in order to assess spectral signatures of the seagrass and carbonate sand
as major seafloor types. WASI enables to use forward or inverse calculation for the spectrum types at
a diapason of 350-800 nm with a 1nm spectral resolution. For the spectral analysis we applied
forward calculations, i.e. a computing and plotting of series of spectra according to specified
parameter settings, with exactly defined depths and cover fraction. \\
The specific parameters have been chosen for the simulation of the environmental conditions where seagrass
grow. The adaptations to life in salt sea water requires various physical and chemical parameters
which include salinity, temperature of $17-25\,^{\circ}C$, light requirements with 10-20 \% on average, ranging
from 4.4 \%  minimal up to 29 \% \cite{McKenzie09}\label{McKenzie09}, so that the zenith angle is taken as $35-45\,^{\circ}$
and reflection factor 0.0201. \textit{P.oceanica}. These values, simulating the environment of the Mediterranean Sea, are fixed (Tab.\ref{tab:3}) among WASI user-defined parameters.
The calculations are done for the spectrum 350-800 nm, covering the most important part of the RS spectrum: 1) Blue-green 0.45 - 0.5 μm; 2) Green 0.5 - 0.6 μm; 3) Red 0.6 - 0.7 μm; 4) Red-NIR 0.7 - 0.8 μm (Fig.\ref{fig:4.8}).

\begin{figure}[h]
	\centering
	\subfloat [Concentration of Gelbstoff at depths 0.5 – 8.0 m]{\label{fig:31a}\includegraphics[width=0.4\textwidth]{Fig-31a.jpg}}
	\subfloat [Bottom reflectance at depth 0.5-4.0 meters]{\label{fig:31b}\includegraphics[width=0.4\textwidth]{Fig-31b.jpg}}
	\caption{Sea water physical properties, modeled by WASI}
	\label{fig:4.8}
\end{figure}

\paragraph{Model parameters: depth and bottom cover fraction}
Although seagrass \textit{P.oceanica} can be found until depth limits down to 40 m depth (Hartog
1970), the most preferable limits of its distribution in the Mediterranean Sea are shallow waters until
4 meters of depth.

\begin{table}
	\rowcolors{1}{Aquamarine}{Azure}
	\caption{Model-specific parameters of water WASI adjusted to simulate environment of the Mediterranean Sea along Crete}
	\centering
	 \begin{tabular}{|  p{2cm} | p{6cm} | p{25mm} | }
	    \hline
		    \textbf{Parameter, WASI} & \textbf{Name and description} & \textbf{Values} \\ \hline\hline
		    C-L & Concentration of large suspended particles & 8 \\ \hline
		   C(i), i=0..5 & Concentration of Phytoplankton & 0.035 -0.089 ug -l \\ \hline
		    bbS & Specific backscattering for small particles & 0.005 m2g-1 \\ \hline
		   T-W & Temperature of water & 17-25 °C \\ \hline
		    n & Exponent of Backscattering by small particles & 0.005 m2g-1 \\ \hline
		    Q &Anisotropy factor of upwelling radiation ("Q-factor") & 5.00 \\ \hline
		    sigma-L & Reflection factor of sky radiance & 0.0201 \\ \hline
		    b1 & Backscattering coefficient of saline waters & 0.00144 m-1 \\ \hline
		    λ0 & Reference Wavelength for Gelbstoff absorption & 440 \\ \hline
		    sun & Sun zenith angle & 45.0 \\ \hline
		    zB & Bottom depth & 4.00 \\ \hline
		   f(i), i=0...5 & Areal fraction of bottom surface type number n & 01/10/10 \\ \hline
		    K0 & Coefficient of Attentuation & 1.0546 \\ \hline
		   view & Viewing angle (0 = nadir) & 0\\ \hline
		    C-X & Concentration of non-chlorophyllous particles (absorption at λ0) & 0 \\ \hline
		   λS & Reference wavelength for scattering of small particles & 500 \\ \hline
		    C-S & Concentration of small suspended particles & 0 \\ \hline
		   S & Exponent of Gelbstoff absorption & 0.0140 \\ \hline
		    C-Y & Concentration of Gelbstoff (absorption at λ0) & 0.400 \\ \hline
		    Bn & BDRF of bottom reflectance (sand) & 0.318 sr -1\\
	    \hline
	  \end{tabular}
	   \label{tab:3}
\end{table}

The increase of depths (zB) influences weakening of light and thus directly affects
production of chlorophyll, because when light passes through the water and suspended particles, it is
being largely modified through the absorption and scattering before it finally reaches plant canopy of
the seagrass. Therefore, the most healthy and suitable areas for the seagrass grow are located at depths
lesser than 4 meters (Fig.\ref{fig:4.9}).

\begin{figure}[H]
	\centering
	\subfloat [depth of 0.5 m]{\label{fig:32a}\includegraphics[width=0.4\textwidth]{Fig-32a.jpg}}
	\subfloat [depth of 2.0 m]{\label{fig:32b}\includegraphics[width=0.4\textwidth]{Fig-32b.jpg}}\\
	\subfloat [depth of 3.5 m]{\label{fig:32c}\includegraphics[width=0.4\textwidth]{Fig-32c.jpg}}
	\subfloat [depths of 0.5, 2.0 and 3.5 m]{\label{fig:32d}\includegraphics[width=0.4\textwidth]{Fig-32d.jpg}}
	\caption{Remote sensing reflectance of \textit{P.oceanica} at various depths}
	\label{fig:4.9}
\end{figure}

An analysis of the spectral reflectance of \textit{P.oceanica} is done using the WASI simulations\index[concepts]{Seawater optical properties!RTM!WASI} in order to
determine, which wavebands can be still used to identify \textit{P.oceanica}. The analysis of spectra shows
that the appropriate wavebands for seagrass mapping lay between 500 and 600 nm and has
also peaks at around 700 nm, ca between 680 and 710 nm (Fig.\ref{fig:4.10}). The highest values of the bottom
reflectance are at spectra of 500-600 nm. The most appropriate depths at which the spectral signatures of the seagrass
could still be discriminated are lesser than 2.5 meters.
The remote sensing properties of the broad-band satellite sensors are often limited and do not allow
easy and unique discrimination of seagrass from other seafloor types. Therefore, we focus on narrow-band
spectra with the resolution of 1 nm.

\paragraph{Spectral discrimination of \textit{P.oceanica}}
The patches of white sandy bottoms of the seafloor, studied in the fieldwork in Ligaria beach, are
much brighter than mattes of P.Oceanica (Fig.\ref{fig:A.6}, \ref{fig:A.5}): seafloor types in Ligaria), which can be clearly seen
at the graph comparing values of the spectral reflectance of the carbonate sand (Fig.\ref{fig:A.35}) and that of P.
oceanica (Fig.\ref{fig:A.28}).
The graph is received in excel spreadsheet using mean values of spectral reflectance of
sand (Fig.\ref{fig:A.27}) and seagrass (Fig.\ref{fig:A.35}), respectively, which have been calculated from measurements of radiance and
irradiance received in Agia Pelagia bay.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.40]{GNU-19.jpg}
	\caption{Statistical analysis of the spectral reflectance of \textit{P.oceanica}: min-max, average values (red bold points), Q1-Q3 areas (green vertical dashed) and 			measured values (dotted): multiplot of measurement sets 200-300. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:4.10}
\end{figure}
\begin{figure}[H]
	\centering
	\subfloat [Sand]{\label{fig:34a}\includegraphics[width=0.4\textwidth]{Fig-27a.jpg}}
	\subfloat [\textit{	extit{P.oceanica}}]{\label{fig:34b}\includegraphics[width=0.4\textwidth]{Fig-27b.jpg}}
	\caption{Bottom albedo of carbonate sand and \textit{P.oceanica}, Agia Pelagia}
	\label{fig:4.11}
\end{figure}
\pagebreak

\subsubsection{Statistical analysis of the observational data and hypothesis testing}\index[concepts]{Data!statistical analysis}
The total amount of measured data was large and included following datasets made using
hyperspectral radiometer Ramses: 350 measurement sets of \textit{P.oceanica} reflectance for 14th Oct (Tab.\ref{tab:A.29}), 400
sets of \textit{P.oceanica} reflectance for 15th Oct (Tab.\ref{tab:A.8}), 84 datasets for seawater reflectance with sediments, 105
datasets for seawater reflectance without sediments, 87 sets for spectral reflectance of carbonate sand.
A statistical approach is evidently necessary for the proper processing of such amounts of data. 
The schematic view of the statistical approach used for the data processing  can be seen on Fig. \ref{fig:4.12}.
Statistical pre-processing of large sets of serial data enables to generalize data by using the most typical and predicted values of
spectral reflectance for further calculations, and to get rid of the extreme values, noise and errors.
The statistical calculators were mainly made by means of Gretle\index[concepts]{Data!statistical analysis!Gretl} and SPSS\index[concepts]{Data!statistical analysis!SPSS}. 
The Open Office was used for preliminary data view and preprocessing, and included following computations
at each data set (for example, Fig.\ref{fig:A.11}) with summary outcome of mean, median, Q1 and Q3, standard deviation, min and max values (Tab.\ref{tab:4}). 
The statistical pre-processing and analysis were made to display the mean values of spectral reflectance,
which were used for comparison with reflectance data of various seafloor cover types as well as
seawater with and without sediment (e.g. Fig..\ref{fig:A.16}).
In the statistical analysis of the raw observed data pool\index[concepts]{Data!measurements pool} of the values of spectral reflectance (about 400 single measurements for \textit{P.oceanica}, e.g. Tab.\ref{tab:A.8}), we summarize their complexity by concentrating on some simple numerical characteristics that they possess, i.e. parameters. Examples are the mean and variance of a probability distribution of measurements dataset (for example, Tab.\ref{tab:4.1}). We divided the total data pool into datasets of 25 measurements for more convenience and ease of use, both for \textit{P.oceanica} and carbonate sand. The mean of a probability distribution of spectral reflectance describes the average value of the random variable over all of its possible realizations (\nameref{app}). Conceptually, there are an infinite number of such realizations, therefore parameters are not known to us. However, in the statistical analysis of the observed data our goal is to estimate these parameters using a finite amount of information available to us: fieldwork observations. \\

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5, angle=90]{UML_StatAn.jpg}
	\caption{Work Flowchart Diagram illustrating algorithm of the Statistical Analysis applied towards the observation dataset (\href{http://staruml.sourceforge.net/en/}{StarUML})}
	\label{fig:4.12}
\end{figure}
\pagebreak

Thus, we collect a number of realizations (a sample of 25 measurements for both sand and \textit{P.oceanica}) and then estimate the statistical parameters (\nameref{app}). \\
Finally we analyzed about 400 datasets totally (example of carbonate sand on Fig.\ref{fig:4.14}, \textit{P.oceanica} - on  Fig.\ref{fig:4.13}) made during different days. 
Since the actual values that measured variables of spectral reflectance take on are not actually known before we observed them, they are random. 
Thus, we analyzed the statistical distribution of the spectral reflectance values of seafloor cover types using common mathematical formulae, implemented in Gretl software (Tab.\ref{tab:19}). 
Their probability expresses uncertainty about the possible values of the spectral reflectance of \textit{P.oceanica} and sand, respectively.
There is a distinction to be made between variables whose values are not yet observed (random variables) and those whose values have been observed (observations). 
Each time we observe the outcome of a random variable\index[concepts]{Data!statistical analysis!variables} of spectral reflectance, we obtain an observation, which is hence no longer random. 
We applied various methods of statistical analysis towards data pool (some examples of the selected tables are in the \nameref{app}: Tab.\ref{tab:18}, Tab.\ref{tab:24}, Tab.\ref{tab:25}, Tab.\ref{tab:26}, Tab.\ref{tab:29}), in order to analyze the distribution of values of spectral reflectance of \textit{P.oceanica}. 
A probability distribution\index[concepts]{Data!statistical analysis!probability distribution}, a mathematical statement about the possible values that the random observations of spectral reflectance can take on, displays the relative frequency with which each possible value of spectral reflectance is observed. 
The \textit{least absolute deviations (LAD)}\index[concepts]{Data!statistical analysis!LAD}, a popular optimization technique, were used to show the main trend of the distribution of spectral reflectance values along the spectra (Tab.\ref{tab:12}). Non-linear logistic analysis\index[concepts]{Data!statistical analysis!non-linear logistic analysis} is applied to specify and estimate a model of spectral distribution, in which the dependent variable (i.e. the value of spectral reflectance for each single observation) is not continuous, but discrete and independent for each case (Tab.\ref{tab:11}). \\
As the observations were made in a repeated way, we also analyzed the data pool using autocorrelation as well: Tab.\ref{tab:13} and Tab.\ref{tab:14} for \textit{P.oceanica} and Tab.\ref{tab:28} for carbonate sand. The \textit{autocorrelation}\index[concepts]{Data!statistical analysis!autocorrelation}, a cross-correlation\index[concepts]{Data!statistical analysis!cross-correlation} of an observed value with itself, made on two different days (14th and 15th October), shows the similarity between observations as a function of the time separation between them. 
As can be seen on Fig.\ref{fig:A.21}, Fig.\ref{fig:A.22} and Fig.\ref{fig:A.29}, it is a useful tool for finding repeating patterns for a single value of observations, such as the presence of a periodic signal which has been buried under noise. The autocorrelation analysis has been performed using both \ac{SPSS} and \ac{Gretl} software. \\
The method of \textit{least squares}\index[concepts]{Data!statistical analysis!least squares, method of} minimizes the sum of the squares of the errors made in solving every single observation made. We tried different approaches of the \textit{least squares} methods: Tab.\ref{tab:15}, Tab.\ref{tab:18}, Tab.\ref{tab:27}.\\
To focus on the relationship between a dependent variable (each single observation of the spectral reflectance) and independent variables (three various depths: 0.5, 1.5 and 2.5 meters) we used the \textit{regression analysis}, which enables modeling and analyzing values of the spectral reflectance at several depths: Tab.\ref{tab:5}, Tab.\ref{tab:6}, Tab.\ref{tab:7}. 
The summary of the values of spectral reflectance properties on various depths is presented on the Tab.\ref{tab:A.7}.
The \textit{k-means cluster analysis}\index[concepts]{Data!statistical analysis!k-means cluster analysis} has been used for partitioning observation sets into k clusters (Tab.\ref{tab:20}, Tab.\ref{tab:21}, Tab.\ref{tab:22}, Tab.\ref{tab:23}) in which each observation belongs to the cluster with the nearest mean.
It enables to highlight the main areas of the location of values of spectral reflectance along the spectra.\\
To estimate spectral density of a spectral reflectance measurements we used periodogram\index[concepts]{Data!statistical analysis!periodogram} function (Tab.\ref{tab:30}), 
which corresponds to the general spectrum of the observations with representation of a variable quantity.  
Thus, the periodogram (Fig. \ref{fig:A.31}) of the series of observation\index[concepts]{Data!statistical analysis!observations, series of} highlights so-called “dead spots” of low power between the frequencies we wish to
exclude (can be seen clearly at the end of the graphs, e.g. Fig.\ref{fig:A.12} ) 
and the frequencies we want to retain (the general profile that lies in the 450-850 nm, Fig.\ref{fig:A.14}, enlarged part of 500-660nm).\\
The visualization of the data plotting was made by means of Gnuplot software, which enables fine
drawing and advanced displaying of large amounts of serial data, ultimate control over graph properties, 
the simplicity of plotting and the ease of scripting (see Fig.\ref{fig:4.13}, \nameref{app}).
The graphs illustrating spectral signatures of various seafloor cover types display the mean values and
areas of quartiles\index[concepts]{Data!statistical analysis!quartiles Q1-Q3} (Q1, Q3, shaded areas on graph, Fig. \ref{fig:4.10}, Fig.\ref{fig:4.14}), which cover most probable data distribution
(spectral reflectance) for each spectral band. In x-axis displayed are the areas of 400 – 950 nm in
spectra, where the measurements were done; the y-axis was adjusted for the better visualization of the
\textit{P.oceanica} spectra: as its spectral values are mostly located in the lower part of the spectra (usually no
more than 0.20 nm, except for borders), we did not extend the y-scale up to 100 percent, in order to
focus on the most necessary area of spectral values (Fig.\ref{fig:4.14}, Fig.\ref{fig:4.4} and \nameref{app}). The mean values are
highlighted using bold red points, as we used these values for plotting the final graph of \textit{P.oceanica} spectra in
Ligaria beach, Agia Pelagia.

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.50]{GNU-13.jpg}
		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 1-100. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
		\label{fig:4.13}
	\end{center}
\end{figure}

A statistical hypothesis\index[concepts]{Data!statistical analysis!hypothesis testing} test has been applied for the making decision and controlling the wealth of the
observational data of the hyperspectral measurements of the water reflectance.
The received results are statistically significant if they are unlikely to have occurred by chance alone,
according to a pre-determined threshold probability, the significance level. Therefore, we applied
critical tests of significance to analyze the measured data according to their significant value.
Answering the first research question, we suggest the following statement.

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.5]{GNU-16.jpg}
		\caption{Multiplot of spectral reflectance of carbonate sand. Series 1-75. Shown mid spread of the statistical quartiles Q1 and Q3 (vertical dashes) and	mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
		\label{fig:4.14}
	\end{center}
\end{figure}

Assuming that the null hypothesis, Hypothesis Ho, is true, which is “the spectral distinguishability of
the seagrass \textit{P.oceanica} from other seafloor types is not changing with varying \textit{in-situ} conditions,
Ho: μ1 =μ2 =μ3=...= μn.”, what is the probability of observing the values of spectral reflectance of
\textit{P.oceanica} at various depths, for the test statistic, that is at least as extreme as the values of spectral
reflectance of \textit{P.oceanica} at these depths, that was actually observed? The alternative Hypothesis Ha
Fig.15. Multiplot display of spectral reflectance of \textit{P.oceanica}. Measurement sets 1-100, 15th
October. Bold red dots show the mean values within each dataset; vertical areas: quartiles Q1-Q2.
claims the opposite statement: “the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly
changing with varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.
Two statistical approaches have been tested for confirmatory data analysis and hypothesis testing.
To analyze average values of spectral reflectance of \textit{P.oceanica} by means of the measured sets we calculated 
mean values for the total set of measurements 1 – 350 (15th October): mean of min and max values, 
mean of average for each waveband, mean for statistical quartiles for each wavelength and mean for median, respectively. 
The adjusted averages evaluate mean values, needed for statistical analysis of the datasets. 
The outcome of this calculations is shown in the Tab.\ref{tab:5} and Tab.\ref{tab:6} (\nameref{app}). 
As a result, we have mean values of spectral reflectance of \textit{P.oceanica} for extreme, average, median and both quartiles,
independent of the individual sets 1-350.

\paragraph{ANOVA-testing}\index[concepts]{Data!statistical analysis!ANOVA}
We applied \ac{ANOVA} (ANalysis Of Variance) testing for the analysis of the probability of the
spectral reflectance of \textit{P.oceanica} to be more or less spectrally distinct from other seafloor cover
types with changing depth. W are checking the statement Hypothesis Ho, is true, which is “the
spectral distinguishability of the seagrass \textit{P.oceanica} from other seafloor types is not changing with
varying \textit{in-situ} conditions, Ho: μ1 =μ2 =μ3=...= μn.” which gives the following outcome.
P(reject H0|H0 is valid)= P(X$>$c|p=)=.05
where c comes for critical values and p=probability.
The result P=.05 is (Tab.\ref{tab:8}), hence, very small, which makes the statement of Hypothesis Ho highly unlikely
(less than 1 in a 10 chance). The one-way \ac{ANOVA} highlighted a significant difference between data
of spectral reflectances of \textit{P.oceanica} at different depths (Tab.\ref{tab:9}), i.e. true is the opposite statement:
Hypothesis Ha=“the spectral discernibility of the seagrass \textit{P.oceanica} is distinctly changing with
varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ$\neq$μ2$\neq$μ3$\neq$ ... $\neq$μn“.

\paragraph{Student-t testing}\index[concepts]{Data!statistical analysis!Student-t testing}
The Student-t test, one of the most commonly used techniques for testing a hypothesis on the basis
of a difference between sample means, is used for the data analysis. In our case the Student t-test demonstrates, if the variation
between two analyzed groups – values of spectral reflectance of seagrass \textit{P.oceanica} and sand - is
significant. Therefore, we use Student t-test to compare two sets of quantitative data of spectral
reflectance of \textit{P.oceanica} and sand, respectively, with samples collected independently of one
another. The Student-t test can be performed knowing just the means, standard deviation, and number
of data points. Therefore, we used the data (\nameref{app}) of means of spectral reflectance of the both
 cover types within data sets, their standard deviation, and the number of data points.
The results show the following outcome.
As a result of the statistical testing, we came to the following conclusion.
The Hypothesis Ha is true, which claims that the spectral discernibility of the seagrass 	\textit{P.oceanica} is
distinctly changing with varying \textit{in-situ} conditions, i.e. increasing depth, Ho: μ1≠μ2≠μ3≠...≠ μn .
Statistical results are illustrated with Error bars made using Excel's vertical box and Whisker Charts
(Box Plots).
The graphs were plotted on the basis of the following statistical data calculated from the sampling
measurements data: Table 2.

\begin{table}
\rowcolors{1}{Khaki}{LightYellow}
\caption{Statistical analysis of the measurements of spectral reflectance of 	extit{P.oceanica} (fragment)}
\centering
  \begin{tabular}{| c | c | c | c | c |}
    \hline
   	 \textbf{Statistical values} & \textbf{318.19} & \textbf{418.3448} & \textbf{518.8361} & \textbf{619.3662} \\ \hline \hline
	    Mean & 0.034521 & 0.049235 & 0.076088 & 0.085287 \\ \hline
	    St Dev & 0.019849 & 0.030338 & 0.042587 & 0.046569 \\ \hline
	   Median & 0.033035 & 0.046197 & 0.07704 & 0.080277 \\ \hline
	   Q1 & 0.019293 & 0.023706 & 0.045606& 0.049468 \\ \hline
	   Q3 & 0.042477 & 0.057934 & 0.092289 & 0.114191 \\ \hline
	   Minimum & 0.007898 & 0.010087 & 0.021373 & 0.022691 \\ \hline
	   Maximum & 0.076835 & 0.110906 & 0.18938 & 0.216231 \\ \hline
	   25th Pct & 0.019293 & 0.023706 & 0.045606 & 0.049468 \\ \hline
	   50th Pct & 0.013743 & 0.022491 & 0.031435 & 0.030808 \\ \hline
	   75th Pct & 0.009441 & -0.01174 & -0.01525 & -0.03391 \\ \hline
	   Min & 0.011395 & 0.013618 & 0.024232 & 0.026777 \\ \hline
	   Max & 0.034358 & 0.052972 & 0.097091 & 0.1 \\ \hline
  \end{tabular}
   \label{tab:4}
\end{table}

where the abbreviations stand for the following values:
Mean: Average of the data to be plotted, AVERAGE\index[concepts]{Data!statistical analysis!average} {data}
St Dev: Standard deviation\index[concepts]{Data!statistical analysis!standard deviation}, \ac{STDEV} {data}
Median: Median of the data, MEDIAN{data}
Calculating interquartile ranges: Q1=First quartile, PERCENTILE({data}*0.25) and Q3=Third
quartile, PERCENTILE({data}*0.75)
Minimum: Minimum value,MIN{data}
Maximum: Maximum value, MAX{data}
25th Pct: Plotting value of first quartile = Q1
50th Pct: Plotting value of median = Median\index[concepts]{Data!statistical analysis!median}-Q1
75th Pct: Plotting value of third quartile = Median-Q3
Min: Lower error bar length=Q1 - Minimum
Max: Upper error bar length=Maximum - Q3
The statistical analysis is displayed on the Graph 1.1 which compares the spread in sets of
measurement data of spectral reflectance of under various environmental conditions (depth).

\subsubsection{Results of the remote sensing application}
The \textit{in situ} spectral reflectance data of \textit{P.oceanica} and sand were used to model these seafloor cover
types as the different sensors would percept them: MODIS\index[concepts]{Remote sensing!multispectral imagery!MODIS}, ASTER\index[concepts]{Remote sensing!multispectral imagery!ASTER}, MERIS\index[concepts]{Remote sensing!multispectral imagery!MERIS}, SeaWiFS\index[concepts]{Remote sensing!multispectral imagery!SeaWiFS} and \ac{CZCS}\index[concepts]{Remote sensing!multispectral imagery!CZCS}.\\
These sensors possess over various percepting characteristics which we briefly illustrated in a small summary table (Tab.\ref{tab:1})
These models simulate spectral views of the chosen sensors, how these sensors will "see" seagrass (Fig.\ref{fig:28}) and sand as
pixels, with accepted default atmosphere and water column effects (given by WASI software). In such
a way we defined an empirical upper limit to the discriminative potential of these sensors. To create
these sensor-specific reflectance spectra, spectral responses of each of these sensors were applied for
the spectral signature of seafloor cover types in full-resolution spectra.

\subsection{Image processing using Erdas Imagine}
The analysis of the imagery of Cretan coasts is based on the images classification and is aimed to
investigate the distribution of the seagrass \textit{P.oceanica} within the research area.
In the image processing part of the research supervised classification has been applied to the aerial
and satellite images. Seagrass meadows and other seafloor cover types were evaluated through a
detailed examination of the imagery. \\
Seagrass beds are clearly visible in color aerial Google Earth\index[concepts]{Remote sensing!aerial imagery!GoogleEarth}
photographs (Fig.\ref{fig:A.36}, Fig.\ref{fig:A.37}), contrasting against slightly-yellow and brownish sand bottom. The seagrass areas are
detected using different bands combinations, masked and studied for the estimation of the changes in
the areas. The classification is based on the spectral properties of the \textit{P.oceanica}, such as brightness,
colour, texture and structure of the seagrass mattes (Fig.\ref{fig:38}a). The raster-based mapping
includes supervised classification with training sites of seagrasses (10-15 set areas) in different bands
for each photograph by classification a series of polygons characteristic of each of the sea floor
bottom types: sandy surface, seagrass bed for each species including \textit{P.oceanica} (meadows), patchy
seagrass bed and algae on rock, rocks, muddy surface, etc (Fig.\ref{fig:38}b).

\begin{figure}
	\centering
	\subfloat [small-scaled (1:54,000)]{\label{fig:38a}\includegraphics[width=0.8\textwidth]{Fig35a.jpg}}\\
	\subfloat [enlarged (1:10,000)]{\label{fig:38b}\includegraphics[width=0.8\textwidth]{Fig35b.jpg}}
	\caption{Seafloor types of the Arina beach: \textit{P.oceanica}, sand, rocks. \\ \href{http://www.google.com/earth/index.html}{Google Earth} images}
	\label{fig:38}
\end{figure}

On the basis of the image processing and classification, applied to the northern coats of Crete, the
marine bottom types between 10 m depth has been mapped, including the limit depths of the of the
seagrass meadows provided on the basis of the available and collected field data. The classification
has been completed using ENVI and Erdas Imagine software.

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.40]{Fig-36.jpg}
		\caption{Results of the image classification in \href{http://www.erdas.com/products/ERDASIMAGINE/ERDASIMAGINE/Details.aspx}{Erdas Imagine}: \\ seagrass distribution in Bali area, Crete. }
		\label{fig:39}
	\end{center}
\end{figure}

\subsection{Seagrass mapping using ArcGIS}

\subsubsection{Data integration}\index[concepts]{Data!integration}
The integrated approach used in this research work has high potential as a means to monitor changes
in seagrass landscape occurring in shallow waters over Crete area.

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.25]{Fig-37.jpg}
		\caption{Google aerial images incorporated into the GIS project: fragment of \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} layout}
		\label{fig:4.17}
	\end{center}
\end{figure}

It encompasses the integration of
high resolution aerial color Google Earth photography, spaceborne satellite imagery, assessment of
spectral signatures using WASI software\index[concepts]{Seawater optical properties!RTM!WASI}, image processing by means of Erdas Imagine (Fig.\ref{fig:39}) and Arc GIS
based mapping. The use of GIS for data incorporation (Fig.\ref{fig:4.17}), storage, analyses, visualizing and mapping
enables to analyze environmental changes within seagrass landscapes based on data from various
sources: aerial and satellite images, geographically referenced maps of Crete island and results of
images classification showing areas of seagrass distribution.
The final mapping has been supported in ArcGIS through the data exporting, conversion and
integration of various data in one GIS-project. Data collected during the fieldwork, imagery
of the seagrass distribution are added into a GIS dataset for the assessment and spatial analysis.

\subsubsection{Upscale mapping of the seagrass landscapes}\index[concepts]{concepts}{Seagrasses!spatial structure}
Although ocean pelagic landscapes have a high degree of spatial variance and less structurally
complex, comparing to the terrestrial ones, the landscape-level phenomena have similar features, and
there are accepted definitions of landscapes elements within the seagrass meadows \cite{Robbins94}\label{Robbins94}. 
In general, the structure of seagrass landscape is simpler than that of the terrestrial
ecosystems in biodiversity and complexity; however, seagrass landscapes show variation in spatial
patterns over different special scales. The complexity of the landscape of seagrass meadow is shown by the
measure of patches in size and shape, expressed in ratio of patch perimeter to area. The fragmentation
of the seagrass landscapes is expressed in contiguity displaying patch aggregation within meadows.\\
The general principles of the hierarchy within the seagrass landscapes are based upon the quantitative
analysis of the spatial patterns, consisted by components and separate elements. Thus, bunches of
individual shoots construct patches, the first hierarchical level. Patches are arranged into discrete
clumps of mattes (at a scale of centimeters to meters) which, in turn, make up beds with l-100m in
diameter. Finally, seagrass beds are arranged into meadows that may extend over kilometer-wide
areas, historically defined as landscapes \cite{Robbins94}\label{Robbins94}.

\begin{figure}[h]
	\centering
	\subfloat [Meadows of the seagrass. Source: \href{http://www.google.com/earth/index.html}{Google Earth}]{\label{fig:41a}\includegraphics[width=0.4\textwidth]{Fig-38a.jpg}}
	\subfloat [Patches (mattes) of the seagrass. Source: \textit{in-situ} videometric measurements]{\label{fig:41b}\includegraphics[width=0.47\textwidth]{Fig-38b.jpg}}
	\caption{Variations in spatial structure of the seagrass landscapes}
	\label{fig:41}
\end{figure}

Besides spatial structure of the seagrass meadows, there are strong and complex patterns of depth zoning, specific to individual
seagrass species, but such detailed classification goes beyond the scope of the current MSc research.
Unique feature of seagrass meadows, most characteristic for shallow areas, is dynamics of their
landscapes. Homogeneous, continuous seagrass meadows can reach in size up to 40 m2 (Fig.\ref{fig:38} a),
however they are often interrupted by gaps and channels of open spaces generated by the complex,
disturbing sedimentation processes, turbulence and turbidity regime of waters and landscape dynamics,
which form gaps and channels within meadows, making them “patchy-looking”, i.e. diversified by
separate mattes (Fig.\ref{fig:38} b). The traditional definition for gap in vegetation cover is “disturbance generated
openings in either floral or faunal cover” \cite{Connell78}\label{Connell78}.

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.25]{Fig-39.jpg}
		\caption{Results of the unsupervised classification of the seafloor cover types and land structure, Agia Pelagia; raster layer read into the \href{http://www.esri.com/software/arcgis/index.html}{ArcGIS} project}
		\label{fig:42}
	\end{center}
\end{figure}

Formation and increase of these
gaps within the mosaic of seagrass meadows is caused by different reasons. The most probable
drivers for the process of gaps formation within seagrass meadows are removal of interior vegetation,
differential growth of seagrass meadows and increased sedimentation. Thus, storms lead to severe
deposition of sediments, burying parts of seagrass meadow, the same effect has movements and
deposition of sediments during and after floods\cite{Bell99}\label{Bell99}.
Increased nutrient sedimentation, especially phosphorus, were explored \cite{Jensen01}\label{Jensen01} as a
potential mechanism for increasing patch dynamics and morphological plasticity within seagrass
meadows. Finally, increasing the degree of fragmentation of the landscapes of 	extit{P.oceanica} meadows
can be caused by the invasion of alien species, such as Cymodocea nodosa, Caulerpa prolifera,
Caulerpa taxifolia. Invaders are in general strong colonizers comparing to native \textit{P.oceanica}: they
occupy much greater habitat space within the regressed meadows of stressed native seagrass
\cite{Montefalcone10}\label{Montefalcone10}. However, on the northern coasts of Crete the only dominating seagrass
species is \textit{P.oceanica}.
Morphological differences in scale of seagrass landscape formations, discussed above, cause need for
the different-scale mapping. Therefore, the investigation of the seagrass meadows at different levels is
performed using underwater videometric measurements, aerial and satellite imagery.

\subsubsection{Results of GIS mapping}
This research part results in series of thematic maps showing seagrass spatial distribution over the Island of
Crete. The research is based on the aerial images classification and analysis using ArcGIS and Erdas Imagine
software. The map of the seagrass locations within the research area of Ligaria beach illustrates the seagrass
distribution at a more detailed scale.

\pagebreak
\section{Discussion and conclusions}\label{sec:5}
The suggested methodology of the analysis of seagrass adaptation towards different environments of
the coastal zones can be applied to other marine coastal areas with growing seagrass.\\
More than 50 \% of the world population lives within one km of the coast which results in continued
anthropogenic pressures in coastal regions. \\Therefore, management of coastal resources and shelf
zone protection in the Mediterranean Sea become increasingly important and require large-scale
monitoring and mapping of the shelf zone as vital instrument for the environmental assessment.
\pagebreak

\section*{Recommendations}\label{sec:6}
\markright{Recommendations}
\addcontentsline{toc}{section}{Recommendations}

To make further studies of \textit{P.oceanica} more effective we would suggest the following recommendations to be considered by the future researchers:
	\begin{description}
		\item[First:] To extend the research area towards the eastern part of the Crete Island, in order to received more regular observations of the 	seagrass locations.
		\item[Second:] To use different sources of imagery and thus, to increase the total collection of scenes covering the research area.
		\item[Third:] To extend the temporal period of the imagery coverage, once the data are available. The current work only includes images covering short temporal period (ca 10-year); further estimation of the dynamics of \textit{P.oceanica} along the coasts of Crete would increase our understanding of the long-scale temporal variations of the seagrass distribution.
		\item[Fourth:] To apply various classifications methods for the available imagery in order to compare the results received by means of various techniques
		\item[Fifth:] To simulate various environmental conditions while modelling optical properties of different seafloor cover types. Not only the depths and 	the chemical content of the seawater should be considered, but also other factors determining the effect of the ecology and health of \textit{P.oceanica}.
		\item[Sixth:] To consider seafloor geomorphology among other factors determining seagrass distribution. If possible, to find out bathymetric data for the research area, and to overlay them with existing images and maps, in order to analyse correlation between spatial distribution of seagrass \textit{P.oceanica} and underwater relief along Cretan coasts. 
		\item[Seventh:] Other RTM software may be tested and the modelling outcomes compared.
		\item[Eighth:] Application of various open source \ac{GIS} (\ac{ILWIS}, \ac{GRASS}) could be very useful for the validation of the cartographic results, assessment of accuracy and comparison of various classification methods. 
		\item[Ninth:] The analysis of the health indicators of the seagrass (such as number of leaves per shoot, biomass estimation within the single shot, etc) was not considered in the current work, as it would go beyond the scope of the MSc thesis. However, ecological investigations could be used for the assessment of the vulnerability of the seagrass meadows in various locations on Crete. 
		\item[Tenth:] A flexible combination of the multi-scale mapping and results of the fieldwork measurements with \ac{GPS}-referenced underwater footage would enable more profound analysis of the coastal environment on Crete. 
	\end{description}

\section*{Keywords}
\markright{Keywords}
\addcontentsline{toc}{section}{Keywords}
\textit{P.oceanica}, seagrass monitoring, spectral reflectance, remote sensing, WASI, Google Earth, Crete
\pagebreak

%%%%%%%%%%%%%%%%%%%  Appendix  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{appendices}
\pagenumbering{roman}
\appendix \label{app}
\appendixpage
\counterwithin{figure}{section}
\counterwithin{table}{section}
%\renewcommand{\thefigure}{A-\arabic{figure}}
%\renewcommand{\thetable}{A-\arabic{table}}

\section{appendices}\label{app}

\subsection{\textit{Appendix-1}. Capturing aerial imagery from the Google Earth: grabbing process}
%\appendixname{\textbf{1: \\Capturing aerial imagery from the Google Earth: grabbing process}}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.6]{Fig-22.jpg}
	\caption{Capturing aerial imagery from the Google Earth: grabbing process}
	\label{fig:A.1}
\end{figure}
\pagebreak

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{Script_Gdal.jpg}
		\caption{Script command of FWTools2.4.7 enabling to reduce the size of the aerial images, from .tif to .ecw format.}
	\end{center}
	\label{fig:A.2}
\end{figure}
\pagebreak

\subsection[\textit{Appendix-2}. Instrumental adjustment and tuning]{\textit{Appendix-2}. Instrumental adjustment and tuning \\ (Trios-RAMSES setup)}
%\appendixname{\textbf{2:\\ Instrumental adjustment and tuning (Trios-RAMSES setup)}}

\begin{table}[H]\footnotesize
	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer and \ac{SAM} during measurement sets. Selected examples (14. X, set 1 and 15.X, set 4).}\label{tab:A.1}
	\centering
	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
	  \begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
		\hline
		    \textbf{Parameters} & \textbf{15 October, set 1} & \textbf{14 October, set 4} \\ \hline \hline
		     Version &1 & 1 \\ \hline
		     IDData & 78B1-2009-10-15-08-56-22-280-272 & 78B1-2009-10-14-15-15-59-342-726 \\ \hline
		     IDDevice & SAM-820C & SAM-8204\\ \hline
		     IDDataType & SPECTRUM & SPECTRUM \\ \hline
		     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
		     DateTime & 2009-10-15 08:56:22 & 10/14/2009, 15:15:59 \\ \hline
		     PositionLatitude & 35.2451=35°14'29.4354” & 35.24752 = 35° 14' 51.072" \\ \hline
		     PositionLongitude & 25.01269=25°0'45.6834" & 25.0098 = 25°0'35.2794" \\ \hline
		     Comment & \textit{P.oceanica} & \textit{P.oceanica} \\ \hline
		     CommentSub1 & 0.5m depth & 2.5m depth \\ \hline
		     CommentSub2 & diffuse & diffuse \\ \hline
		     CommentSub3 & Agia Pelagia & Agia Pelagia site 2, close to rocks \\ \hline
		     IDMethodType & SAM Control & SAM Control \\ \hline
		    MethodName & SAM-820C & SAM-8204 \\ \hline
		    Mission & No Mission & No Mission \\ \hline
		    MissionSub & 1 & 1 \\ \hline
		    RecordType & 0 & 0 \\ \hline
		    CalFactor & 1 & 1 \\ \hline
		   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
		   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
		   IntegrationTime & 512 & 1024 \\ \hline
		    P31 & -1 & -1 \\ \hline
		   P31e & 0 & 0 \\ \hline
		   PathLength & +INF & +INF \\ \hline
		   RAWDynamic & 65535 & 65535 \\ \hline
		   Temperature & +NAN & +NAN \\ \hline
		   Unit1 & 0101 Wavelength nm &  1 1 Wavelength nm \\ \hline
		   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
		   Unit3 & f0-06 Error mW/(m2 nm) & 3-3 Error mW/(m2 nm Sr) \\ \hline
		  Unit4 & f1-00 Status & f1-00 Status \\ \hline
	  \end{tabular}	
\end{table}
\pagebreak

\begin{table}[H]\footnotesize
	\caption{Attributes of the Trios-RAMSES hyperspectral radiometer during measurement sets. Selected examples (15. X, set 1 and 14.X, set 4).}\label{tab:A.2}	\centering
	\rowcolors{1}{Lavender}{GhostWhite}
	\begin{tabular}{| p{3cm} | p{3cm} | p{5cm} |}
	\hline
	    \textbf{Parameters} & \textbf{14 October, set 1} & \textbf{15 October, set 4} \\ \hline \hline
	     Version &1 & 1 \\ \hline
	     IDData & 78B1-2009-10-14-15-32-02-078-323 & 78B1-2009-10-15-10-36-02-217-136 \\ \hline
	     IDDevice & SAM-820C & SAM-8204\\ \hline
	     IDDataType & SPECTRUM & SPECTRUM \\ \hline
	     IDDataTypeSub1 & CALIBRATED & CALIBRATED \\ \hline
	      DateTime & 2009-10-14 15:32:02 & 2009-10-15 10:36:02 \\ \hline
	     PositionLatitude & 35.24752= 35°14'51.072" & 35.245071=35°14'42.2556" \\ \hline
	     PositionLongitude & 25.0098= 25°0'35.2794" & 25.012661=25°0'45.5796" \\ \hline
	     Comment & \textit{P.oceanica} & \textit{P.oceanica} \\ \hline
	     CommentSub1 & 3.5m depth & 1.5m depth \\ \hline
	     CommentSub2 & diffuse & diffuse \\ \hline
	     CommentSub3 & Agia Pelagia & Agia Pelagia site 1, close to rocks \\ \hline
	     IDMethodType & SAM Control & SAM Control \\ \hline
	    MethodName & SAM-820C & SAM-8204 \\ \hline
	    Mission & No Mission & No Mission \\ \hline
	    MissionSub & 1 & 1 \\ \hline
	    RecordType & 0 & 0 \\ \hline
	    CalFactor & 1 & 1 \\ \hline
	   IDDataBack & DLAB-2008-02-06-14-13-18-865-675 & DLAB-2008-01-25-20-42-29-607-062 \\ \hline
	   IDDataCal & DLAB-2008-02-06-14-23-00-187-767 & DLAB-2008-01-28-08-40-24-220-395 \\ \hline
	   IntegrationTime & 64 & 32 \\ \hline
	    P31 & -1 & -1 \\ \hline
	   P31e & 0 & 0 \\ \hline
	   PathLength & +INF & +INF \\ \hline
	   RAWDynamic & 65535 & 65535 \\ \hline
	   Temperature & +NAN & +NAN \\ \hline
	   Unit1 & 01-01 Wavelength nm &  01-01 Wavelength nm \\ \hline
	   Unit2 & 03-06 Intensity mW/(m2nm) & 03 03 Intensity mW/(m2 nm Sr) \\ \hline
	   Unit3 & f0-06 Error mW/(m2 nm) & f0-03 Error mW/(m2 nm Sr) \\ \hline
	  Unit4 & f1-00 Status & f1-00 Status \\ \hline
	\end{tabular}
\end{table}
\pagebreak

\subsection[\textit{Appendix-3}. Types of seagrass structural patterns]{\textit{Appendix-3}. Types of seagrass structural patterns, Ligaria beach, Crete}
%\appendixname{\textbf{3: \\Types of seagrass structural patterns, Ligaria beach, Crete}}

\begin{figure}[h]
	\centering
	\subfloat [Fragment of meadow]{\includegraphics[width=0.3\textwidth]{Fragment_meadow.jpg}}
	\hspace{1mm}
	\subfloat [Aggregated seagrass patch]{\includegraphics[width=0.3\textwidth]{aggreg_patch.jpg}}
	\hspace{1mm}
	\subfloat [Isolated patch of seagrass]{\includegraphics[width=0.3\textwidth]{isolated_patch.jpg}}
	\caption{Types of seagrass structural patterns, Ligaria beach, Crete}
	\label{fig:A.3}
\end{figure}

\subsection{\textit{Appendix-4}. Sampling design of the depths measurements. Ligaria beach, Crete}
%\appendixname{\textbf{4: \\Sampling design of the depths measurements. Ligaria beach, Crete}}

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.35]{Sampling_design.jpg}
		\caption{Scheme of the depths measurements; fieldwork on Ligaria beach}\label{fig:A.4}
	\end{center}	
\end{figure}
\pagebreak

\subsection{\textit{Appendix-5}. Results of videographic measurements: seafloor types on Crete island}
%\appendixname{\textbf{5: \\Results of videographic measurements: seafloor types on Crete island}}

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.45]{Fig-3-14.jpg}
		\caption{Ligaria beach, Crete: seafloor types}\label{fig:A.5}
	\end{center}
\end{figure}

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.25]{Fig-3-15.jpg}
		\caption{Various seafloor types; Ligaria beach, Crete}\label{fig:A.6}
	\end{center}
\end{figure}
\pagebreak

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.35]{Fig-3-16.jpg}
		\caption{Measurement underwater equipment}\label{fig:A.7}		
	\end{center}
\end{figure}
\pagebreak

\subsection{\textit{Appendix-6}. Data pre-processing}
%\appendixname{\textbf{6: \\Data pre-processing}}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.55]{Fig-24.jpg}
	\caption{Script written on Python, for the interpolation raw data of the Trios-RAMSES measurements}
	\label{fig:A.8}
\end{figure}
\pagebreak

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{Fig-28.jpg}
	\caption{Fragment of \textit{Bottom.R} file: values of spectral measurements of the
		seagrasses (various species), sand, silt and artificial spectrum of constant albedo}
	\label{fig:A.9}
\end{figure}
\pagebreak

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.35]{Interpolation.jpg}
		\caption{Interpolation of the spectral measurements by means of Open Office}
		\label{fig:A.10}
	\end{center}
\end{figure}
\pagebreak

\subsection{\textit{Appendix-7}. Data processing and statistical evaluation: Results of the measurements of the seafloor optical properties.}
%\appendixname{\textbf{7:\\ Data processing and statistical evaluation: \\ Results of the measurements of the seafloor optical properties.}}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-21.jpg}
	\caption{Multiplot graph showing spectral reflectance of the seawater without sediments, measured in aquarium tank, Agia Pelagia district, Crete. \href{http://www.gnuplot.info/}{Gnuplot} \\Two complimentary graphs below show the results of the statistical analysis}
	\label{fig:27}
\end{figure}
\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.20]{App-1.jpg}
		\caption{Values of spectral reflectances of seagrass \textit{P.oceanica} (fragment)}
		\label{fig:A.11}
	\end{center}
\end{figure}
\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-401-420.jpg}
		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 401-420. \href{http://www.gnuplot.info/}{Gnuplot}}
		\label{fig:A.12}
	\end{center}
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-401-420-candles.jpg}
		\caption{Statistical analysis of the measurement data: spectral reflectance of the \textit{P.oceanica}. 
		Visualization of the interquartile ranging. Example of data set 401-420. Shown mid-spread of  statistical quartiles Q1 and Q3, min and max values within the range. \href{http://www.gnuplot.info/}{Gnuplot}}
		\label{fig:A.13}
	\end{center}
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-enlarged-M-401-420.jpg}
		\caption{Enlarged fragment of the statistical analysis of the measurement data. Example of data set 401-420. 
		Visualization of the interquartile ranging and plotted together with measurement data. \href{http://www.gnuplot.info/}{Gnuplot}}
		\label{fig:A.14}
	\end{center}
\end{figure}
\pagebreak
%%%

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Model summary of the \textit{regression} analysis: curve estimation and ANOVA table tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}.\ac{SPSS}}
	\begin{center}
		\begin{tabular}{|c|c|c|c|}
			\hline\hline
			\textbf{R} & \textbf{R square} & \textbf{Adjusted R square} & \textbf{Std. Error of the Estimate} \\ \hline\hline
			.277 1 & .077 & .068126692 & .462 \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:5}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{\textit{ANOVA} table: exponential curve estimation in the regression analysis, tested for single observations within one measurement set: spectral reflectance of \textit{P.oceanica}. \ac{SPSS}}
	\begin{center}
		\begin{tabular}{|c|c|c|c| c|c|}
			\hline\hline
			& \textbf{Sum of squares} & \textbf{df} & \textbf{Mean Square} & F & Sig.\\ \hline\hline
			Regression & 1.810 & 1 & 1.810 & 8.477 & .004\\ \hline
			Residual & 21.781 & 102 & .214 & & \\ \hline
			Total & 23.592 & 103 & & & \\ \hline
		\end{tabular}
	\end{center}
	The independent variable is wavelength. 
	\label{tab:6}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Coefficients of the \textit{regression analysis} (exponential curve estimation) of the spectral reflectance of \textit{P.oceanica}. \href{http://www.spss.com/}{SPSS}}
	\begin{center}
		\begin{tabular}{| c | c | p{2cm}| c | c | c |}
			\hline\hline
			& \multicolumn{2}{|c|}{Unstandardized Coefficients}\\
			\cline{2-3}
			& B & Std.Error & Stand. Coef. Beta & t & Sig.\\ \hline\hline
			wavelength & -.001 & .000 & -.277 & -2.912 & .004 \\ \hline
			constant & .464 & .121 & & 3.821 & .000 \\ \hline
		\end{tabular}
	\end{center}
	The underlying process assumed is independence (white noise).\\
	 Based on the asymptotic chi-square approximation.
	\label{tab:7}
\end{table}
\pagebreak

\begin{figure}[h]
	\begin{center}
	\includegraphics[scale=0.20]{GNU-Stat-M-401-420.jpg}
	\caption{Fragment of the statistical analysis of the \textit{P.oceanica} reflectance. Example of data set 401-420. 
		Visualization of the measured data together with statistical values: interquartile ranging, medians, means. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:A.15}
	\end{center}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.25]{GNU-Radiance_Bezier.jpg}
	\caption{Radiance of the seawater with sediments. Measured in aquarium tank. Bezier Interpolation. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:A.16}
\end{figure}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Results of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters. \ac{SPSS}}\label{tab:8}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|c|}
		\hline\hline
		\textbf{Source of Variation} & \textbf{SS} & \textbf{df} & \textbf{MS} & \textbf{F} & \textbf{P value} & \textbf{F crit} \\ \hline
		Between Groups & 373841.7048 & 2 & 186920.8524 & 407.85359 & 1.11677 & 3.01153 \\ \hline
		Within Groups & 261233.1668 & 570 & 458.3038014 & & & \\ \hline
		Total &  635074.8716 & 572 & & & & \\ \hline
	\end{tabular}
	\end{center} 
	P more than .05, which means that there is a significant difference in radiance of \textit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).

\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Summary of the \textit{ANOVA one-way} analysis: results of the \textit{single factor} (depth) testing of the radiance of \textit{P.oceanica} at various depths: 0.5, 1.5 and 2.5 meters. \href{http://www.spss.com/}{SPSS}}\label{tab:A.7}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|}
		\hline\hline
		\textbf{Groups} & \textbf{Count} & \textbf{Sum} & \textbf{Average} & \textbf{Variance} \\ \hline\hline
		Column 1 & 191 &	760.126692 & 3.9797209 & 18.6948046 \\ \hline
		Column 2 & 191 &	4527.214218 & 23.7026922  & 251.4863978 \\ \hline
		Column 3 & 191 &	12465.19972 & 65.2628257 & 1104.7302015 \\ \hline
	\end{tabular}
	\end{center}
	P$>$0.05, which means that there is a significant difference in radiance of \textit{P.oceanica} at three different depth (0.5, 1.5 and 2.5).

\end{table}
\pagebreak

\rowcolors{1}{GhostWhite}{LightCyan}
\begin{longtable}{|c|c|c|c|c|c|c|}
	\caption{Results of the statistical analysis of spectral reflectance of \textit{P.oceanica}, sets 1-350). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 15th October}\label{tab:A.8} \\
	  \multicolumn{1}{|l|}{\textbf{wl}} &
	   \multicolumn{1}{l|}{\textbf{mean}} & 
	   \multicolumn{1}{l|}{\textbf{Q1}} & 
	   \multicolumn{1}{l|}{\textbf{min}} & 
	   \multicolumn{1}{l|}{\textbf{Q3}} & 
	   \multicolumn{1}{l|}{\textbf{max}} & 
	   \multicolumn{1}{l|}{\textbf{median}} \\ \hline
	\endfirsthead

	\multicolumn{7}{c}{Table A.8: Spectral reflectance of \textit{P.oceanica} \textit{(continued from previous page)}}\\
	\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
	\endhead

	\textbf{318.23354} & 0.030467 & 0.038907 & 0.029764 & 0.069762 & 0.116100 & 0.048984 \\ \hline
	\textbf{321.57666} & 0.028129 & 0.037310 & 0.027965 & 0.066572 & 0.110918 & 0.046450 \\ \hline
	\textbf{324.92025} & 0.030163 & 0.038570 & 0.029502 & 0.068216 & 0.114120 & 0.047710 \\ \hline
	\textbf{328.26428} & 0.030457 & 0.038936 & 0.029789 & 0.069098 & 0.116080 & 0.048574 \\ \hline
	\textbf{331.60876} & 0.031407 & 0.040192 & 0.030731 & 0.071247 & 0.119681 & 0.050299 \\ \hline
	\textbf{334.95367} & 0.032480 & 0.041545 & 0.031781 & 0.073494 & 0.123651 & 0.051941 \\ \hline
	\textbf{338.29900} & 0.033560 & 0.043030 & 0.032847 & 0.076252 & 0.128290 & 0.054334 \\ \hline
	\textbf{341.64474} & 0.034067 & 0.043938 & 0.033356 & 0.077587 & 0.131152 & 0.055506 \\ \hline
	\textbf{344.99089} & 0.035346 & 0.045272 & 0.034616 & 0.080212 & 0.135395 & 0.057653 \\ \hline
	\textbf{348.33743} & 0.036199 & 0.046719 & 0.035465 & 0.082716 & 0.139829 & 0.059655 \\ \hline
	\textbf{351.68436} & 0.037336 & 0.048027 & 0.036587 & 0.085062 & 0.144101 & 0.061509 \\ \hline
	\textbf{355.03166} & 0.038611 & 0.049598 & 0.037852 & 0.087963 & 0.149193 & 0.063835 \\ \hline
	\textbf{358.37932} & 0.039897 & 0.051320 & 0.039131 & 0.091143 & 0.154409 & 0.066354 \\ \hline
	\textbf{361.72735} & 0.040712 & 0.052408 & 0.039951 & 0.093252 & 0.157952 & 0.068034 \\ \hline
	\textbf{365.07572} & 0.041541 & 0.053517 & 0.040782 & 0.095491 & 0.161650 & 0.069682 \\ \hline
	\textbf{368.42442} & 0.042378 & 0.054625 & 0.041622 & 0.097634 & 0.164944 & 0.071524 \\ \hline
	\textbf{371.77346} & 0.043682 & 0.056205 & 0.042926 & 0.100782 & 0.169458 & 0.074006 \\ \hline
	\textbf{375.12281} & 0.044525 & 0.057191 & 0.043779 & 0.102916 & 0.172533 & 0.075888 \\ \hline
	\textbf{378.47247} & 0.045786 & 0.058849 & 0.045039 & 0.105993 & 0.177150 & 0.078339 \\ \hline
	\textbf{381.82243} & 0.047361 & 0.060698 & 0.046615 & 0.109681 & 0.182421 & 0.081219 \\ \hline
	\textbf{385.17269} & 0.048457 & 0.062067 & 0.047717 & 0.112447 & 0.186097 & 0.083586 \\ \hline
	\textbf{388.52322} & 0.049503 & 0.063533 & 0.048767 & 0.115361 & 0.190440 & 0.086143 \\ \hline
	\textbf{391.87402} & 0.051271 & 0.065536 & 0.050529 & 0.119310 & 0.196618 & 0.089165 \\ \hline
	\textbf{395.22508} & 0.052033 & 0.066436 & 0.051298 & 0.121231 & 0.199547 & 0.090884 \\ \hline
	\textbf{398.57640} & 0.052346 & 0.066859 & 0.051625 & 0.122328 & 0.201038 & 0.091905 \\ \hline
	\textbf{401.92796} & 0.053246 & 0.067963 & 0.052517 & 0.124750 & 0.204926 & 0.093940 \\ \hline
	\textbf{405.27975} & 0.054472 & 0.069505 & 0.053749 & 0.127847 & 0.209718 & 0.096463 \\ \hline
	\textbf{408.63176} & 0.055492 & 0.070692 & 0.054773 & 0.130382 & 0.213406 & 0.098620 \\ \hline
	\textbf{411.98399} & 0.056331 & 0.071703 & 0.055621 & 0.132628 & 0.216692 & 0.100545 \\ \hline
	\textbf{415.33642} & 0.057682 & 0.073235 & 0.056976 & 0.135758 & 0.221309 & 0.103354 \\ \hline
	\textbf{418.68905} & 0.058942 & 0.074842 & 0.058240 & 0.138984 & 0.226028 & 0.106207 \\ \hline
	\textbf{422.04186} & 0.060504 & 0.076815 & 0.059805 & 0.142833 & 0.231707 & 0.109487 \\ \hline
	\textbf{425.39485} & 0.062547 & 0.079376 & 0.061847 & 0.147773 & 0.238892 & 0.113584 \\ \hline
	\textbf{428.74801} & 0.064817 & 0.082141 & 0.064115 & 0.153137 & 0.246681 & 0.118126 \\ \hline
	\textbf{432.10132} & 0.066664 & 0.084493 & 0.065967 & 0.157743 & 0.253230 & 0.122093 \\ \hline
	\textbf{435.45478} & 0.068616 & 0.086926 & 0.067925 & 0.162414 & 0.259765 & 0.126092 \\ \hline
	\textbf{438.80838} & 0.070690 & 0.089761 & 0.069999 & 0.167825 & 0.267712 & 0.130580 \\ \hline
	\textbf{442.16211} & 0.072701 & 0.092360 & 0.072010 & 0.172808 & 0.275027 & 0.134721 \\ \hline
	\textbf{445.51596} & 0.074550 & 0.094801 & 0.073855 & 0.177420 & 0.282203 & 0.138517 \\ \hline
	\textbf{448.86992} & 0.076196 & 0.096995 & 0.075496 & 0.181512 & 0.288621 & 0.141853 \\ \hline
	\textbf{452.22397} & 0.077581 & 0.098795 & 0.076876 & 0.184975 & 0.294309 & 0.144634 \\ \hline
	\textbf{455.57812} & 0.079009 & 0.100624 & 0.078296 & 0.188516 & 0.300251 & 0.147486 \\ \hline
	\textbf{458.93235} & 0.080107 & 0.102089 & 0.079392 & 0.191424 & 0.304933 & 0.149762 \\ \hline
	\textbf{462.28665} & 0.081264 & 0.103609 & 0.080545 & 0.194453 & 0.309937 & 0.152238 \\ \hline
	\textbf{465.64102} & 0.082366 & 0.105087 & 0.081658 & 0.197331 & 0.314862 & 0.154601 \\ \hline
	\textbf{468.99543} & 0.083316 & 0.106273 & 0.082589 & 0.199757 & 0.319033 & 0.156521 \\ \hline
	\textbf{472.34990} & 0.084192 & 0.107427 & 0.083464 & 0.202075 & 0.323294 & 0.158350 \\ \hline
	\textbf{475.70439} & 0.085122 & 0.108714 & 0.084392 & 0.204674 & 0.327802 & 0.160310 \\ \hline
	\textbf{479.05891} & 0.086278 & 0.110348 & 0.085546 & 0.207940 & 0.332934 & 0.162881 \\ \hline
	\textbf{482.41344} & 0.088011 & 0.112712 & 0.087275 & 0.212533 & 0.340090 & 0.166595 \\ \hline
	\textbf{485.76798} & 0.090064 & 0.115295 & 0.089321 & 0.217585 & 0.348041 & 0.170644 \\ \hline
	\textbf{489.12252} & 0.091822 & 0.117678 & 0.091073 & 0.222210 & 0.355070 & 0.174440 \\ \hline
	\textbf{492.47704} & 0.093593 & 0.120058 & 0.092835 & 0.226915 & 0.361882 & 0.178261 \\ \hline
	\textbf{495.83154} & 0.095851 & 0.123152 & 0.095072 & 0.232757 & 0.370879 & 0.182979 \\ \hline
	\textbf{499.18600} & 0.098498 & 0.126592 & 0.097689 & 0.239161 & 0.380603 & 0.188192 \\ \hline
	\textbf{502.54043} & 0.100497 & 0.129243 & 0.099648 & 0.243918 & 0.387852 & 0.192036 \\ \hline
	\textbf{505.89480} & 0.101834 & 0.131069 & 0.100939 & 0.246962 & 0.392456 & 0.194588 \\ \hline
	\textbf{509.24911} & 0.103094 & 0.133022 & 0.102139 & 0.249966 & 0.397001 & 0.196947 \\ \hline
	\textbf{512.60336} & 0.104998 & 0.136069 & 0.103967 & 0.254523 & 0.404017 & 0.200556 \\ \hline
	\textbf{515.95752} & 0.107481 & 0.139807 & 0.106364 & 0.259986 & 0.412790 & 0.205238 \\ \hline
	\textbf{519.31159} & 0.110222 & 0.143976 & 0.109022 & 0.265948 & 0.421370 & 0.210237 \\ \hline
	\textbf{522.66557} & 0.112891 & 0.148090 & 0.111650 & 0.272217 & 0.430485 & 0.215339 \\ \hline
	\textbf{526.01943} & 0.115799 & 0.152266 & 0.114534 & 0.278942 & 0.440382 & 0.220946 \\ \hline
	\textbf{529.37318} & 0.118594 & 0.156156 & 0.117302 & 0.285384 & 0.450045 & 0.226407 \\ \hline
	\textbf{532.72680} & 0.121111 & 0.159690 & 0.119794 & 0.291491 & 0.458964 & 0.231403 \\ \hline
	\textbf{536.08029} & 0.123519 & 0.163063 & 0.122169 & 0.297338 & 0.467779 & 0.236144 \\ \hline
	\textbf{539.43363} & 0.125694 & 0.166156 & 0.124310 & 0.302678 & 0.475842 & 0.240492 \\ \hline
	\textbf{542.78681} & 0.127313 & 0.168524 & 0.125897 & 0.306828 & 0.482311 & 0.243819 \\ \hline
	\textbf{546.13983} & 0.128393 & 0.170255 & 0.126944 & 0.309747 & 0.487098 & 0.246131 \\ \hline
	\textbf{549.49267} & 0.129369 & 0.171844 & 0.127887 & 0.312256 & 0.491208 & 0.248155 \\ \hline
	\textbf{552.84533} & 0.130192 & 0.173167 & 0.128678 & 0.314539 & 0.495323 & 0.249936 \\ \hline
	\textbf{556.19780} & 0.131149 & 0.174662 & 0.129597 & 0.317126 & 0.499790 & 0.251984 \\ \hline
	\textbf{559.55006} & 0.131844 & 0.175865 & 0.130257 & 0.319385 & 0.503907 & 0.253726 \\ \hline
	\textbf{562.90211} & 0.132098 & 0.176510 & 0.130479 & 0.320731 & 0.506658 & 0.254511 \\ \hline
	\textbf{566.25394} & 0.131944 & 0.176786 & 0.130273 & 0.321450 & 0.508802 & 0.254761 \\ \hline
	\textbf{569.60554} & 0.131139 & 0.176171 & 0.129456 & 0.320617 & 0.508657 & 0.253650 \\ \hline
	\textbf{572.95690} & 0.129337 & 0.174497 & 0.127627 & 0.317523 & 0.505482 & 0.250781 \\ \hline
	\textbf{576.30800} & 0.126457 & 0.171261 & 0.124730 & 0.311678 & 0.498350 & 0.245834 \\ \hline
	\textbf{579.65885} & 0.121915 & 0.166084 & 0.120186 & 0.302078 & 0.485661 & 0.237873 \\ \hline
	\textbf{583.00943} & 0.116464 & 0.159695 & 0.114741 & 0.289985 & 0.469210 & 0.227889 \\ \hline
	\textbf{586.35972} & 0.110124 & 0.152038 & 0.108417 & 0.275585 & 0.449536 & 0.216026 \\ \hline
	\textbf{589.70973} & 0.102719 & 0.142758 & 0.101038 & 0.258779 & 0.425248 & 0.201868 \\ \hline
	\textbf{593.05945} & 0.094713 & 0.132871 & 0.093063 & 0.240383 & 0.399506 & 0.186687 \\ \hline
	\textbf{596.40885} & 0.086028 & 0.121987 & 0.084416 & 0.220594 & 0.372134 & 0.170438 \\ \hline
	\textbf{599.75794} & 0.077236 & 0.111113 & 0.075669 & 0.200557 & 0.343876 & 0.154016 \\ \hline
	\textbf{603.10670} & 0.070071 & 0.102079 & 0.068539 & 0.183962 & 0.320865 & 0.140543 \\ \hline
	\textbf{606.45513} & 0.064972 & 0.095380 & 0.063465 & 0.172189 & 0.304973 & 0.131080 \\ \hline
	\textbf{609.80321} & 0.061747 & 0.091191 & 0.060257 & 0.164863 & 0.295132 & 0.125235 \\ \hline
	\textbf{613.15094} & 0.059707 & 0.088656 & 0.058235 & 0.160355 & 0.289194 & 0.121614 \\ \hline
	\textbf{616.49830} & 0.058283 & 0.086915 & 0.056821 & 0.157314 & 0.285336 & 0.119161 \\ \hline
	\textbf{619.84529} & 0.057038 & 0.085386 & 0.055588 & 0.154781 & 0.281737 & 0.117003 \\ \hline
	\textbf{623.19189} & 0.055996 & 0.084129 & 0.054556 & 0.152697 & 0.278854 & 0.115243 \\ \hline
	\textbf{626.53811} & 0.055200 & 0.083282 & 0.053763 & 0.151068 & 0.277016 & 0.113879 \\ \hline
	\textbf{629.88392} & 0.054368 & 0.082234 & 0.052931 & 0.149414 & 0.275056 & 0.112433 \\ \hline
	\textbf{633.22932} & 0.053299 & 0.080860 & 0.051869 & 0.147176 & 0.271850 & 0.110512 \\ \hline
	\textbf{636.57430} & 0.052033 & 0.079147 & 0.050611 & 0.144415 & 0.268135 & 0.108222 \\ \hline
	\textbf{639.91885} & 0.050841 & 0.077438 & 0.049432 & 0.141810 & 0.265181 & 0.106188 \\ \hline
	\textbf{643.26296} & 0.049737 & 0.075977 & 0.048336 & 0.139466 & 0.263239 & 0.104485 \\ \hline
	\textbf{646.60662} & 0.048467 & 0.074330 & 0.047076 & 0.136488 & 0.260596 & 0.102251 \\ \hline
	\textbf{649.94982} & 0.046762 & 0.072058 & 0.045397 & 0.132376 & 0.255761 & 0.099096 \\ \hline
	\textbf{653.29255} & 0.044482 & 0.068866 & 0.043159 & 0.126637 & 0.248247 & 0.094730 \\ \hline
	\textbf{656.63481} & 0.041366 & 0.064447 & 0.040108 & 0.118961 & 0.236639 & 0.088686 \\ \hline
	\textbf{659.97657} & 0.038118 & 0.059971 & 0.036997 & 0.110774 & 0.223366 & 0.082411 \\ \hline
	\textbf{663.31784} & 0.035847 & 0.056353 & 0.034750 & 0.104541 & 0.212843 & 0.077616 \\ \hline
	\textbf{666.65861} & 0.034385 & 0.054006 & 0.033356 & 0.100476 & 0.205602 & 0.074622 \\ \hline
	\textbf{669.99886} & 0.033576 & 0.052732 & 0.032604 & 0.098106 & 0.201320 & 0.072881 \\ \hline
	\textbf{673.33858} & 0.033122 & 0.052033 & 0.032194 & 0.096638 & 0.198744 & 0.072021 \\ \hline
	\textbf{676.67777} & 0.032897 & 0.051894 & 0.032003 & 0.095851 & 0.197694 & 0.071600 \\ \hline
	\textbf{680.01641} & 0.032892 & 0.052265 & 0.032012 & 0.095789 & 0.198389 & 0.071686 \\ \hline
	\textbf{683.35451} & 0.033331 & 0.053301 & 0.032432 & 0.097053 & 0.201578 & 0.072478 \\ \hline
	\textbf{686.69203} & 0.033823 & 0.054940 & 0.032870 & 0.098577 & 0.206179 & 0.073481 \\ \hline
	\textbf{690.02899} & 0.034458 & 0.056468 & 0.033426 & 0.099736 & 0.209895 & 0.074120 \\ \hline
	\textbf{693.36536} & 0.034436 & 0.056844 & 0.033329 & 0.100112 & 0.211708 & 0.074967 \\ \hline
	\textbf{696.70114} & 0.034695 & 0.057874 & 0.033554 & 0.100594 & 0.213198 & 0.075367 \\ \hline
	\textbf{700.03633} & 0.034756 & 0.058426 & 0.033643 & 0.101504 & 0.215413 & 0.075789 \\ \hline
	\textbf{703.37090} & 0.034209 & 0.058339 & 0.033206 & 0.101601 & 0.215696 & 0.075467 \\ \hline
	\textbf{706.70485} & 0.033056 & 0.057120 & 0.031999 & 0.100262 & 0.213853 & 0.073792 \\ \hline
	\textbf{710.03817} & 0.031680 & 0.055368 & 0.030599 & 0.097286 & 0.209716 & 0.071411 \\ \hline
	\textbf{713.37085} & 0.030077 & 0.053314 & 0.028958 & 0.093925 & 0.206402 & 0.068982 \\ \hline
	\textbf{716.70288} & 0.028935 & 0.050808 & 0.027789 & 0.089359 & 0.202502 & 0.066018 \\ \hline
	\textbf{720.03426} & 0.026287 & 0.047084 & 0.025162 & 0.082940 & 0.193535 & 0.061184 \\ \hline
	\textbf{723.36497} & 0.023116 & 0.041962 & 0.022192 & 0.075244 & 0.182863 & 0.055350 \\ \hline
	\textbf{726.69500} & 0.019755 & 0.037276 & 0.019223 & 0.067583 & 0.171617 & 0.049522 \\ \hline
	\textbf{730.02435} & 0.016355 & 0.032668 & 0.016168 & 0.059609 & 0.159449 & 0.042677 \\ \hline
	\textbf{733.35299} & 0.013247 & 0.028377 & 0.013325 & 0.052370 & 0.147592 & 0.036934 \\ \hline
	\textbf{736.68094} & 0.010923 & 0.024633 & 0.011200 & 0.046727 & 0.138634 & 0.032350 \\ \hline
	\textbf{740.00817} & 0.009740 & 0.022355 & 0.010145 & 0.042861 & 0.132244 & 0.029414 \\ \hline
	\textbf{743.33467} & 0.008848 & 0.020756 & 0.009299 & 0.040084 & 0.127367 & 0.027450 \\ \hline
	\textbf{746.66045} & 0.008789 & 0.019615 & 0.009272 & 0.038166 & 0.123288 & 0.025956 \\ \hline
	\textbf{749.98548} & 0.007817 & 0.018631 & 0.008305 & 0.036577 & 0.119290 & 0.024701 \\ \hline
	\textbf{753.30975} & 0.007533 & 0.017951 & 0.008013 & 0.035542 & 0.116242 & 0.023967 \\ \hline
	\textbf{756.63327} & 0.007777 & 0.018797 & 0.008273 & 0.036601 & 0.119191 & 0.024621 \\ \hline
	\textbf{759.95601} & 0.008651 & 0.020234 & 0.009191 & 0.039371 & 0.128165 & 0.026558 \\ \hline
	\textbf{763.27797} & 0.008355 & 0.020033 & 0.008883 & 0.038779 & 0.126589 & 0.026439 \\ \hline
	\textbf{766.59914} & 0.007232 & 0.017224 & 0.007708 & 0.033862 & 0.113056 & 0.022658 \\ \hline
	\textbf{769.91951} & 0.005575 & 0.015235 & 0.005970 & 0.029884 & 0.103584 & 0.020196 \\ \hline
	\textbf{773.23907} & 0.006016 & 0.014264 & 0.006407 & 0.028277 & 0.100504 & 0.019270 \\ \hline
	\textbf{776.55781} & 0.005801 & 0.013966 & 0.006189 & 0.027876 & 0.099265 & 0.019096 \\ \hline
	\textbf{779.87573} & 0.006019 & 0.014157 & 0.006397 & 0.028320 & 0.100314 & 0.019322 \\ \hline
	\textbf{783.19280} & 0.006136 & 0.014643 & 0.006524 & 0.028946 & 0.101978 & 0.019978 \\ \hline
	\textbf{786.50903} & 0.006559 & 0.015475 & 0.006967 & 0.030261 & 0.104960 & 0.021034 \\ \hline
	\textbf{789.82440} & 0.007038 & 0.016900 & 0.007440 & 0.031880 & 0.107764 & 0.022351 \\ \hline
	\textbf{793.13891} & 0.007598 & 0.017699 & 0.008021 & 0.033662 & 0.111651 & 0.023996 \\ \hline
	\textbf{796.45254} & 0.008137 & 0.018912 & 0.008556 & 0.035621 & 0.114760 & 0.025299 \\ \hline
	\textbf{799.76528} & 0.008607 & 0.020030 & 0.009050 & 0.037323 & 0.118123 & 0.026535 \\ \hline
	\textbf{803.07713} & 0.009078 & 0.020935 & 0.009544 & 0.038749 & 0.120699 & 0.027598 \\ \hline
	\textbf{806.38808} & 0.009299 & 0.021600 & 0.009772 & 0.039865 & 0.122278 & 0.028669 \\ \hline
	\textbf{809.69811} & 0.009526 & 0.022082 & 0.009994 & 0.040470 & 0.124400 & 0.029071 \\ \hline
	\textbf{813.00722} & 0.009575 & 0.022239 & 0.010086 & 0.040394 & 0.124681 & 0.029112 \\ \hline
	\textbf{816.31539} & 0.010059 & 0.021517 & 0.010572 & 0.039929 & 0.124390 & 0.028547 \\ \hline
	\textbf{819.62263} & 0.008777 & 0.020295 & 0.009288 & 0.037598 & 0.118055 & 0.026799 \\ \hline
	\textbf{822.92891} & 0.007647 & 0.017661 & 0.008133 & 0.034147 & 0.112749 & 0.024239 \\ \hline
	\textbf{826.23423} & 0.006897 & 0.015734 & 0.007343 & 0.030561 & 0.104462 & 0.021328 \\ \hline
	\textbf{829.53858} & 0.005932 & 0.013286 & 0.006340 & 0.026347 & 0.095980 & 0.018276 \\ \hline
	\textbf{832.84195} & 0.004547 & 0.010824 & 0.005261 & 0.022895 & 0.087424 & 0.015291 \\ \hline
	\textbf{836.14433} & 0.003285 & 0.009222 & 0.003567 & 0.019407 & 0.080997 & 0.013090 \\ \hline
	\textbf{839.44571} & 0.002562 & 0.008010 & 0.002864 & 0.017472 & 0.076515 & 0.011432 \\ \hline
	\textbf{842.74608} & 0.002580 & 0.007522 & 0.002094 & 0.016318 & 0.073901 & 0.010511 \\ \hline
	\textbf{846.04544} & 0.003121 & 0.007283 & 0.003442 & 0.015689 & 0.080570 & 0.010195 \\ \hline
	\textbf{849.34376} & 0.002719 & 0.006935 & 0.002961 & 0.016016 & 0.076860 & 0.009926 \\ \hline
	\textbf{852.64105} & 0.002742 & 0.006733 & 0.002994 & 0.015949 & 0.070643 & 0.009824 \\ \hline
	\textbf{855.93730} & 0.002840 & 0.007120 & 0.003106 & 0.015919 & 0.074262 & 0.010103 \\ \hline
	\textbf{859.23249} & 0.002653 & 0.006921 & 0.002897 & 0.016121 & 0.070377 & 0.009889 \\ \hline
	\textbf{862.52661} & 0.002825 & 0.006846 & 0.003153 & 0.016677 & 0.078344 & 0.009943 \\ \hline
	\textbf{865.81966} & 0.002508 & 0.007308 & 0.002710 & 0.016406 & 0.073265 & 0.010239 \\ \hline
	\textbf{869.11162} & 0.002322 & 0.007310 & 0.002532 & 0.017118 & 0.081434 & 0.010382 \\ \hline
	\textbf{872.40249} & 0.001873 & 0.007289 & 0.002542 & 0.017948 & 0.086487 & 0.010688 \\ \hline
	\textbf{875.69226} & 0.001769 & 0.007640 & 0.001769 & 0.017981 & 0.074717 & 0.010912 \\ \hline
	\textbf{878.98091} & 0.002327 & 0.008080 & 0.002821 & 0.018625 & 0.082462 & 0.011617 \\ \hline
	\textbf{882.26844} & 0.001661 & 0.008815 & 0.002284 & 0.019452 & 0.079564 & 0.011996 \\ \hline
	\textbf{885.55484} & 0.002940 & 0.008588 & 0.003400 & 0.021344 & 0.080998 & 0.012489 \\ \hline
	\textbf{888.84010} & 0.002827 & 0.009126 & 0.003467 & 0.022626 & 0.089420 & 0.013547 \\ \hline
	\textbf{892.12421} & 0.002523 & 0.010281 & 0.002997 & 0.023568 & 0.129197 & 0.015324 \\ \hline
	\textbf{895.40716} & 0.001578 & 0.010879 & 0.001578 & 0.025648 & 0.091357 & 0.016328 \\ \hline
	\textbf{898.68893} & 0.004077 & 0.013098 & 0.004709 & 0.031217 & 0.101146 & 0.019767 \\ \hline
	\textbf{901.96953} & 0.004560 & 0.014295 & 0.005228 & 0.032824 & 0.141000 & 0.020699 \\ \hline
	\textbf{905.24894} & 0.004167 & 0.014798 & 0.004249 & 0.034899 & 0.135645 & 0.022265 \\ \hline
	\textbf{908.52716} & 0.003891 & 0.015616 & 0.002986 & 0.037403 & 0.150828 & 0.024065 \\ \hline
	\textbf{911.80416} & 0.003641 & 0.018706 & 0.004306 & 0.043272 & 0.115375 & 0.027328 \\ \hline
	\textbf{915.07995} & 0.003120 & 0.018093 & 0.003995 & 0.045688 & 0.135804 & 0.027234 \\ \hline
	\textbf{918.35451} & 0.006215 & 0.020162 & 0.007229 & 0.045213 & 0.199711 & 0.029950 \\ \hline
	\textbf{921.62783} & 0.004669 & 0.021630 & 0.005940 & 0.047785 & 0.245588 & 0.032131 \\ \hline
	\textbf{924.89991} & 0.003051 & 0.019678 & 0.003957 & 0.051741 & 0.127093 & 0.032278 \\ \hline
	\textbf{928.17074} & 0.002955 & 0.021413 & 0.004277 & 0.053042 & 0.266342 & 0.035277 \\ \hline
	\textbf{931.44030} & 0.002679 & 0.023349 & 0.004029 & 0.057574 & 0.225003 & 0.036074 \\ \hline
	\textbf{934.70859} & 0.003446 & 0.023229 & 0.004285 & 0.059098 & 0.549439 & 0.037722 \\ \hline
	\textbf{937.97559} & 0.005783 & 0.026132 & 0.007307 & 0.065475 & 0.195072 & 0.040848 \\ \hline
	\textbf{941.24130} & 0.005209 & 0.024892 & 0.006147 & 0.058814 & 0.281685 & 0.039602 \\ \hline
	\textbf{944.50571} & 0.003354 & 0.025994 & 0.004299 & 0.064521 & 0.232462 & 0.041118 \\ \hline
	\textbf{947.76881} & 0.003156 & 0.023476 & 0.003813 & 0.059536 & 0.246006 & 0.037035 \\ \hline
	\textbf{951.03058} & 0.002296 & 0.026581 & 0.002380 & 0.067924 & 0.265845 & 0.041463 \\ \hline
\end{longtable}
\pagebreak

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.30]{GNU-Irr-Splines.jpg}
		\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Splines interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}­}
		\label{fig:A.17}
	\end{center}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.30]{GNU-Irr.jpg}
	\caption{Irradiance of the seawater, measured in aquarium tank. Smooth Bezier interpolation. Visualization in \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:A.18}
\end{figure}
\pagebreak
\newpage
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-gamma-rad.jpg}
		\caption{Frequency normality test against gamma distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}}
		\label{fig:A.19}
	\end{center}
\end{figure}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{G-normal-dist.jpg}
	\caption{Frequency normality test against normal distribution: radiance of the seawater, measured in aquarium tank. Visualization in \href{http://gretl.sourceforge.net/}{Gretl}­}
	\label{fig:A.20}
\end{figure}

%%% Gretle start
\begin{table}
	\caption{Nonlinear model: results of the \textit{logistic analysis} of the seawater radiance with sediments (15.X.), Series:V16. \href{http://gretl.sourceforge.net/}{Gretl}}
	\begin{center}
	Model 2: Logistic, using observations 1:01--8:23 ($T$ = 191)\\
	Dependent variable: v16\\
	$\hat{y} = 1 / (1 + e^{-X\hat{\beta}})$

	\vspace{1em}
	\begin{tabular}{|c|c|c|c|c|}
		  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
		const &  1.58324 &     0.407258 &       3.8876 &         0.0001 \\
		v1 &   $-$0.00512630 &     0.000615503 &       $-$8.3286 &         0.0000 \\
	\end{tabular}

	\vspace{1em}Statistics based on the transformed data:

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Sum squared resid &  462.8481 & S.E. of regression &  1.564906 \\
		$R^2$ &  0.268481 & Adjusted $R^2$ &  0.264610 \\
		$F(1, 189)$ &  69.36634 & P-value($F$) &  1.63\textrm{e--14} \\
		Log-likelihood & $-$355.5467 & Akaike criterion &  715.0935 \\
		Schwarz criterion &  721.5980 & Hannan--Quinn &  717.7281 \\
		$\hat{\rho}$ &  0.980662 & Durbin--Watson &  0.020351 \\
	\end{tabular}

	\vspace{1em}Statistics based on the original data:

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Mean dependent var &  0.262542 & S.D. dependent var &  0.205031 \\
		Sum squared resid &  10.80565 & S.E. of regression &  0.239108 \\
	\end{tabular}
	\end{center}
	\label{tab:11}
\end{table}
%%% end Gretl

%%%
\begin{table}
	\rowcolors{1}{Lavender}{LightYellow}
	\caption{Robust estimation of the seawater radiance (measured in aquarium tank, Heraklion): results of the \textit{least absolute deviation} (LAD), Series:V16. \href{http://gretl.sourceforge.net/}{Gretl}}
	\begin{center}
		Model 6: LAD, using observations 1:01--8:23 ($T$ = 191)\\
		Dependent variable: v16\\
	\vspace{1em}

	\begin{tabular}{|c c c c c|}
		  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
		const &   0.738911 &     0.0862011 &       8.5719 &         0.0000 \\
		v1 &   $-$0.000775186 &     0.000114010 &       $-$6.7993 &         0.0000 \\
	\end{tabular}

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Median depend. var &  0.227857 & S.D. dependent var &  0.205031 \\
		Sum absolute resid &  30.22342 & Sum squared resid &  8.059632 \\
		Log-likelihood &  28.74727 & Akaike criterion & $-$53.49454 \\
		Schwarz criterion & $-$46.99000 & Hannan--Quinn & $-$50.85990 \\
	\end{tabular}
	\label{tab:12}
	\end{center}
\end{table}
%%%
\pagebreak
\newpage

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline\hline
		& & & \multicolumn{3}{c|}{Box-Ljung Statistic}\\
		\cline{4-6}
			Lag & Autocorrelation & Std error(a) & Value & df & Sig (b)\\ \hline\hline
			1 & .844 & .071 & 141.324 & 1 & .000 \\ \hline
			2 & .816 & .071 & 273.208 & 2 & .000 \\ \hline
			3 & .799 & .071 & 400.323 & 3 & .000 \\ \hline
			4 & .780 & .071 & 522.217 & 4 & .000 \\ \hline
			5 & .723 & .070 & 627.372 & 5 & .000 \\ \hline
			6 & .700 & .070 & 726.581 & 6 & .000 \\ \hline
			7 & .745 & .070 & 839.623 & 7 & .000 \\ \hline
			8 & .624 & .070 & 919.336 & 8 & .000 \\ \hline
			9 & .584 & .070 & 989.540 & 9 & .000 \\ \hline
			10 & .562 & .069 & 1054.920 & 10 & .000 \\ \hline
			11 & .522 & .069 & 1111.681 & 11 & .000 \\ \hline
			12& .464 & .069 & 1156.693 & 12 & .000 \\ \hline
			13 & .475 & .069 & 1204.314 & 13 & .000 \\ \hline
			14 & .463 & .069 & 1249.726 & 14 & .000 \\ \hline
			15 & .401 & .069 & 1283.916 & 15 & .000 \\ \hline
			16 & .389 & .068 & 1316.427 & 16 & .000 \\ \hline
	\end{tabular}
	\end{center}
	The underlying process assumed is independence (white noise).\\
	Based on the asymptotic chi-square approximation.
	\label{tab:13}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.45]{Autocorr.jpg}
	\caption{\textit{Autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
	\label{fig:A.21}
\end{figure}
\pagebreak
\newpage

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Results of the partial \textit{autocorrelation analysis} of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.), Series:V3. \href{http://www.spss.com/}{SPSS}}\label{tab:14}
	\begin{center}
	\begin{tabular}{|c|c|c|}
		\hline\hline
			Lag & Partial Autocorrelation & Std error\\ \hline\hline
			1 & .844 & .072 \\ \hline
			1 &.358 & .072 \\ \hline
			3 & .217 &.072 \\ \hline
			4 & .125 &.072 \\ \hline
			5 & -.093 & .072 \\ \hline
			6 & .003 & .072 \\ \hline
			7 & .310 & .072 \\ \hline
			8 & -.403 & .072 \\ \hline
			9 & -.117 & .072 \\ \hline
			10 & .033 & .072 \\ \hline
			11 & -.104 & .072 \\ \hline
			12 & .023 &.072 \\ \hline
			13 & .246 &.072 \\ \hline
			14 & -.164 & .072 \\ \hline
			15 & .084 & .072 \\ \hline
			16 & .152 & .072 \\ \hline
	\end{tabular}
	\end{center}
\end{table}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.45]{Partial_Corr.jpg}
	\caption{Partial correlation analysis of the measurement set 1-16 of the spectral reflectance of \textit{P.oceanica} (15.X.). Visualization in \href{http://www.spss.com/}{SPSS}}
	\label{fig:A.22}
\end{figure}
\pagebreak
\newpage

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-09.jpg}
		\caption{Spectral reflectance of the seawater with sediments. Bezier interpolation. Mean value is shown by the vertical impulses linestyle. \href{http://www.gnuplot.info/}{Gnuplot}­}
		\label{fig:A.23}
	\end{center}
\end{figure}

%%% Gretl start
\begin{table}[H]
	\rowcolors{1}{Bisque}{Ivory}
	\begin{center}
		\caption{\textit{Ordinary Least Squares}: results of the \textit{OLS} analysis of the measurement set 326-350 of the spectral reflectance of \textit{P.oceanica} (15.X.). \href{http://gretl.sourceforge.net/}{Gretl}}
		Model 2: OLS, using observations 326--350\\
		Dependent variable: v20 \\

		\vspace{1em}
		\begin{tabular}{c c c c c}
		  & {Coefficient} &  {Std.\ Error} &   {$t$-ratio} &    {p-value} \\[1ex]
		const &  0.414957 &    0.0299342 &      13.8623 &        0.0000 \\
		v1 &   $-$0.000441803 &     4.52185\textrm{e--005} &       $-$9.7704 &         0.0000 \\ \hline\hline
	\end{tabular}

	\vspace{2em}
	\begin{tabular}{lrlr}
		Mean dependent var &  0.134005 & S.D. dependent var &  0.140653 \\
		Sum squared resid &  2.497416 & S.E. of regression &  0.114951 \\
		$R^2$ &  0.335585 & Adjusted $R^2$ &  0.332070 \\
		$F(1, 189)$ &  95.46076 & P-value($F$) &  1.65\textrm{e--18} \\
		Log-likelihood &  143.1678 & Akaike criterion & $-$282.3357 \\
		Schwarz criterion & $-$275.8311 & Hannan--Quinn & $-$279.7010 \\
	\end{tabular}
	\label{tab:15}
	\end{center}
\end{table}
\pagebreak
%%% Gretl end

\begin{figure}[h]
	\begin{center}
		\includegraphics[scale=0.22]{GNU-10.jpg}
		\caption{Spectral reflectance of the seawater without sediments. Interpolation graph. \href{http://www.gnuplot.info/}{Gnuplot} display­}
		\label{fig:A.24}
	\end{center}
\end{figure}

%%% Gretl start
\begin{table}[H]
	\rowcolors{1}{Thistle}{MintCream}
	\begin{center}
		\caption{Summary statistics of measurements set: 301-325. \\ Example of selected variables 12-22. \href{http://gretl.sourceforge.net/}{Gretl}­}\label{tab:16}
		Summary Statistics, using the observations 301--325\\

	\vspace{2em}

	\begin{tabular}{|c|c|c|c|c|}
		Variable & {Mean}& {Median}  & {Minimum}    & {Maximum} \\[1ex]
		v12 & 0.221475 & 0.139032 & 0.00101700 & 0.780730\\
		v13 & 0.217074 & 0.139285 & 0.00108700 & 0.754430\\
		v14 & 0.167442 & 0.110878 & 0.000348000 & 0.575242\\
		v15 & 0.155908 & 0.111737 & 0.0125650 & 0.483602\\
		v16 & 0.158329 & 0.111647 & 0.0101590 & 0.498096\\
		v17 & 0.167575 & 0.111349 & 0.000836000 & 0.579004\\
		v18 & 0.125868 & 0.0863670 & 0.000662000 & 0.417701\\
		v19 & 0.179261 & 0.117898 & 0.000935000 & 0.612021\\
		v20 & 0.141302 & 0.0932400 & 0.000714000 & 0.478343\\
		v21 & 0.228822 & 0.143105 & 0.00149500 & 0.800529\\
		v22 & 0.174892 & 0.114994 & 0.000343000 & 0.592146\\ \hline\hline
	\end{tabular}
	\begin{tabular}{|c|c|c|c|c|}
		\rowcolors{1}{Lavender}{MintCream}
		Variable &  {Std.\ Dev.}  & {Coef.Var.}   & {Skewness}    & {Ex.\ kurtosis} \\[1ex]
		v12 & 0.246134 & 1.11134 & 0.930598 & $-$0.393370\\
		v13 & 0.239324 & 1.10250 & 0.910879 & $-$0.440017\\
		v14 & 0.181797 & 1.08573 & 0.884089 & $-$0.466508\\
		v15 & 0.151811 & 0.973718 & 0.796534 & $-$0.646866\\
		v16 & 0.155834 & 0.984241 & 0.832126 & $-$0.587890\\
		v17 & 0.182383 & 1.08837 & 0.901570 & $-$0.435120\\
		v18 & 0.134253 & 1.06662 & 0.820573 & $-$0.603688\\
		v19 & 0.195217 & 1.08901 & 0.873615 & $-$0.515800\\
		v20 & 0.152787 & 1.08128 & 0.863700 & $-$0.525357\\
		v21 & 0.252696 & 1.10434 & 0.924172 & $-$0.408867\\
		v22 & 0.189797 & 1.08523 & 0.858809 & $-$0.546215\\
	\end{tabular}
	\end{center}

\end{table}
%%% Gretle end
\pagebreak

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Summary of the table with values of the spectral reflectance of \textit{P.oceanica} measured on three different depths, R}
	\begin{center}
	\begin{tabular}{|c|c|c|c|}
		\hline\hline
		wl & X0.5m & X1.5m & X2.5m\\ \hline\hline
		Min.   :318.2   &  Min.   :0.00000     &   Min.   : 0.7792     &   Min.   :0.005096  \\ \hline
		 1st Qu.:476.9   &  1st Qu.:0.03398    &  1st Qu.:14.4084    &  1st Qu.:0.091878  \\ \hline
		 Median :636.1   &   Median :0.10955    &  Median :22.4621   &  Median :0.269328 \\ \hline 
		 Mean   :635.6   &   Mean   :0.11392   &   Mean   :19.6899   &  Mean   :0.301121  \\ \hline
		 3rd Qu.:794.5   &   3rd Qu.:0.17198   &   3rd Qu.:25.8785    &  3rd Qu.:0.431055 \\ \hline  
		 Max.   :950.9     &  Max.   :0.32937    &  Max.   :30.9573   &   Max.   :0.981314  \\ \hline
	\end{tabular}
	\end{center}
		From: DepthData $<$- read.table(file="Three-Depths.dat", sep="", header=T); summary(DepthData)
	\label{tab:17}
\end{table}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-11.jpg}
		\caption{Spectral reflectance of \textit{P.oceanica}. Measurement series 25-50. \\ \href{http://www.gnuplot.info/}{Gnuplot} display­}
		\label{fig:A.25}
	\end{center}
\end{figure}
\pagebreak

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{GNU-12.jpg}
		\caption{Remote sensing reflectance of \textit{P.oceanica}. Series 1-25. \\Shown midspread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range.  \href{http://www.gnuplot.info/}{Gnuplot}­}
		\label{fig:A.26}
	\end{center}
\end{figure}

%%% Gretl start
\begin{table}[htbp]
	\caption{\textit{Two-Stage Least Squares}: results of the \textit{TSLS} analysis of the measurement set 1-191 of the spectral reflectance of \textit{P.oceanica} (15.X.). \href{http://gretl.sourceforge.net/}{Gretl}}
	\begin{center}
		Model 1: TSLS, using observations 1--191\\
		Dependent variable: const\\
		Instrumented: v17 \\
		Instruments: v12 \\

	\vspace{1em}

	\begin{tabular}{|c c c c c|}
		  &  {Coefficient} &  {Std.\ Error} &  {$z$} &    {p-value} \\[1ex]
		v17 &  6.88502 & 0.527503 & 13.0521 & 0.0000 \\
	\end{tabular}

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Mean dependent var &  1.000000 & S.D. dependent var &  0.000000 \\
		Sum squared resid &  97.51535 & S.E. of regression &  0.716407 \\
		Log-likelihood & $-$237.5799 & Akaike criterion &  477.1598 \\
		Schwarz criterion &  480.4121 & Hannan--Quinn &  478.4771 \\
	\end{tabular}

	\vspace{1em}
	\begin{raggedright}
		Hausman test --\\
		\quad Null hypothesis: OLS estimates are consistent\\
		\quad Asymptotic test statistic: $\chi^2(1)$ = 13.2834\\
		\quad with p-value = 0.000267768\\
	\vspace{1ex}
		Weak instrument test -- \\
		\quad First-stage $F(1, 190)$ = 16670.2 \\
	\end{raggedright}

	\end{center}
	\label{tab:18}
\end{table}
%%% Gretl end

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{GNU-14.jpg}
		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 126-150. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
		\label{fig:A.27}
	\end{center}
\end{figure}

%%% Gretl start
\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Basic mathematical formulae used for statistical analysis of the measurement set of the spectral reflectance.}
	\label{tab:13}
	\begin{center}
	\begin{tabular}{|c|c|}
		\hline\hline
		Basic Statistics & Math Formulae \\ \hline\hline
		\hline

		\vspace{1ex}

		Mean & $\Sigma(x_i/n=\bar x)$ \\ \hline
		Variance & $1 \over n-1$ $\Sigma(x_i-\bar x)=s^2_x$  \\ \hline
		Standard Deviation & $s=s^2$ \\ \hline
		Coefficient of Variation & $s/\bar x$ \\ \hline
		Skewness & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^3$  \\ \hline
		Excess Kurtosis & $1 \over n-1$ $\Sigma(x_i-\bar x)/s^4-3$  \\ \hline
	\end{tabular}
	\end{center}
	\label{tab:19}
\end{table}
%%% Gretl end
\pagebreak

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Initial Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
	\label{tab:13}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|}
		\hline\hline
		& \multicolumn{2}{|c|}{Cluster} &  & \multicolumn{2}{|c|}{Cluster}\\
		 & 1 & 2 &  & 1 & 2 \\ \hline\hline
		V2 &	.353010 &	.000644 &	V16 &	.240091 &	.000374 \\ \hline
		V3 &	 .271519 &	.000394 &	V17 &	.270306 &	.000435 \\ \hline
		V4 &	.273594 &	.000644 &	V18 &	.196188 &	.000211 \\ \hline
		V5 &	.406525 &	.000527 &	V19 &	.301486 &	.000370 \\ \hline
		V6 &	.344973 &	.000509 &	V20 &	.178575 &	.000328 \\ \hline
		V7 &	.395117 &	.000646 &	V21 &	.231708 &	.000440 \\ \hline
		V8 &	.254642 &	.000352 &	V22 &	.186493 &	.000365 \\ \hline
		V9 &	.288881 &	.000267 &	V23 &	.261761 &	.000418 \\ \hline
		V10 &	.191850 &	.000255 &	V24 &	.282032 &	.000264 \\ \hline
		V11 &	.249462 &	.000323 &	V25 &	.190709 &	.000375 \\ \hline
		V12 &	.209622 &	.000414 &	V26 &	.177432 &	.000206 \\ \hline
		V13 &	.218312 &	.000206 &	V27 &	.195104 &	.000268 \\ \hline
		V14 &	.177432 &	.000268 &	V28 &	.256630 &	.000386 \\ \hline
		V15 &	.184836 &	.000223 &	V29 &	.283744 &	.000436 \\ \hline
		 &	 &	 &		V30 &	.406525 &	.000646 \\ \hline
	\end{tabular}
	\end{center}
	\label{tab:20}
\end{table}

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Iteration History(a): results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
	\begin{center}
	\begin{tabular}{|c|c|c|}
		\hline\hline
		& \multicolumn{2}{|c|}{Change in Cluster} \\
		 Iteration & 1 & 2 \\ \hline\hline
		1 &	.326 &	.206 \\ \hline
		2 &	.024 & 	.010 \\ \hline
		3 &	.008 &	.003 \\ \hline
		4 &	.000 &	.000 \\ \hline
	\end{tabular}
	\end{center}
	Convergence achieved due to no or small change in cluster centers. The maximum absolute coordinate change for any center is .000. The current iteration is 4. The minimum distance between initial centers is 1.435.
\label{tab:21}
\end{table}

\begin{table}[H]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Number of Cases in each Cluster: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
		\begin{center}
		\begin{tabular}{|c|c|}
		\hline\hline
			Cluster 1 & 55.000 \\ \hline
			Cluster 2 & 136.000 \\ \hline
			Valid &	191.000 \\ \hline
			Missing & .000 \\ \hline
		\end{tabular}
		\end{center}
	\label{tab:22}
\end{table}
\pagebreak

\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Final Cluster Centers: results of the \textit{K-means} analysis of the measurement set 201-236 of the spectral reflectance of \textit{P.oceanica} (15.X.).  \href{http://www.spss.com/}{SPSS}}
	\begin{center}
	\begin{tabular}{|c|c|c|c|c|c|}
	\hline\hline
		& \multicolumn{2}{|c|}{Cluster} &  & \multicolumn{2}{|c|}{Cluster}\\
		 & 1 & 2 &  & 1 & 2 \\ \hline\hline
		V2  &	0.265141 &	0.046966 &	V16 &	0.181395 &	0.033191 \\ \hline
		V3 &	0.202643 &	0.036088 &	V17 &	0.202698 &	0.036614 \\ \hline
		V4 &	0.20699 &	0.037789 &	V18 &	0.149364 &	0.028217 \\ \hline
		V5 &	0.298975 &	0.052158 &	V19 &	0.225452 &	0.040193 \\ \hline
		V6 &	0.25915 &	0.045968 &	V20 &	0.135163 &	0.025506 \\ \hline
		V7 &	0.293508 &	0.049314 &	V21 &	0.175176 &	0.031866 \\ \hline
		V8 &	0.193292 &	0.036402 &	V22 &	0.142295 &	0.027148 \\ \hline
		V9 &	0.215848 &	0.038738 &	V23 &	0.19514 &	0.034553 \\ \hline
		V10 &	0.147798 &	0.027489 &	V24 &	0.214946 &	0.038321 \\ \hline
		V11 &	0.188955 &	0.034031 &	V25 &	0.14585 &	0.027665 \\ \hline
		V12 &	0.157947 &	0.029319 &	V26 &	0.134787 &	0.024912 \\ \hline
		V13 &	0.165877 &	0.029961 &	V27 &	0.149217 &	0.028326 \\ \hline
		V14 &	0.137149 &	0.02626 &	V28 &	0.193388 &	0.034999 \\ \hline
		V15 &	0.140563 &	0.026211 &	V29 &	0.214436 &	0.038433 \\ \hline
		 &	 &	 &	V30	 & 0.299183	 & 0.053451 \\ \hline
	\end{tabular}
	\end{center}
	\label{tab:23}
\end{table}

%%%Gretl begin
\begin{table}[H]
	\caption{Results of the \textit{Principal Components Analysis}: measurement variables 1-7 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica} (15.X.). \href{http://gretl.sourceforge.net/}{Gretl}}
	\begin{center}
	Eigenanalysis of the Correlation Matrix\\
	\vspace{1em}

		\begin{tabular}{|c|c|c|c|}
			{Component } &  {Eigenvalue } & {Proportion} & {Cumulative} \\
			1 &   6.4848    &   0.9264    &   0.9264 \\
			2 & 0.4949 &      0.0707  &     0.9971 \\
			3 &  0.0140    &   0.0020    &   0.9991 \\
			4 &    0.0052  &     0.0007    &   0.9998 \\
			5 &   0.0008   &    0.0001     &  1.0000 \\
			6 &  0.0002    &   0.0000      & 1.0000 \\
			7 &   0.0001   &    0.0000     &  1.0000 \\
		\end{tabular}

	Eigenvectors (component loadings)

	\vspace{1ex}

		\begin{tabular}{|c|c|c|c|c|c|c|c|}
			v1        &	     -0.308  &	  -0.880  &	   0.334   &	 -0.126  &	   0.033  &	  -0.042   &	  0.006\\ \hline
			v2      &	      0.383  &	  -0.295  &	  -0.512   &	 -0.230  &	  -0.365  &	  -0.370   &	 -0.425\\ \hline
			v3      &	      0.383  &	  -0.300  &	  -0.465   &	 -0.010  &	   0.562  &	   0.307   &	  0.370\\ \hline
			v4      &	      0.390  &	  -0.145  &	   0.203   &	  0.714  &	  -0.129  &	  -0.415   &	  0.293\\ \hline
			v5      &	      0.392  &	  -0.091  &	   0.300   &	  0.262  &	   0.020  &	   0.568   &	 -0.597\\ \hline
			v6      &	      0.392  &	  -0.011  &	   0.259   &	 -0.410  &	  -0.543  &	   0.299   &	  0.476\\ \hline
			v7      &	      0.390  &	   0.138  &	   0.460   &	 -0.429  &	   0.488  &	  -0.427   &	 -0.113\\ \hline
		\end{tabular}
	\end{center}
	\label{tab:24}
\end{table}
%%%Gretl end
\pagebreak

%%%Gretl begin
\begin{table}[H]
	\begin{center}
		\rowcolors{1}{AntiqueWhite}{LightGoldenrodYellow}
		\caption{Results of the \textit{tobit analysis} with censored dependent variable (2) from the selected dataset Series 151-175 of the s.reflectance of \textit{P.oceanica}. \href{http://gretl.sourceforge.net/}{Gretl}}
			Model 2: Tobit, using observations 1:01--8:23 ($T$ = 191)\\
			Dependent variable: v12\\
		\vspace{1em}
	\begin{tabular}{|c c c c c|}
		  &  {Coefficient} &    {Std.\ Error} & {$z$} & {p-value} \\[1ex]
		const &   $-$0.00413795 &     0.0502867 &     $-$0.0823 &         0.9344 \\
		v2 &   0.890117 &     0.957730 &       0.9294 &         0.3527 \\
	\end{tabular}

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Mean dependent var &  0.035372 & S.D. dependent var &  0.053185 \\
		Censored obs &         1 & sigma &  0.050920 \\
		Log-likelihood &  295.1357 & Akaike criterion & $-$584.2715 \\
		Schwarz criterion & $-$574.5147 & Hannan--Quinn & $-$580.3195 \\
		$\hat{\rho}$ & $-$0.032865 & Durbin--Watson &  2.065699 \\
	\end{tabular}

	\vspace{1em}
	\begin{raggedright}
		Test for normality of residual --\\
		\quad Null hypothesis: error is normally distributed\\
		\quad Test statistic: $\chi^2(2)$ = 190.404\\
		\quad with p-value = 4.51226e-042\\
	\vspace{1ex}
	\end{raggedright}
	\label{tab:25}
	\end{center}
\end{table}
%%%Gretl end

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.38]{GNU-15.jpg}
		\caption{Multiplot of spectral reflectance of \textit{P.oceanica}. Series 151-200. Shown mid-spread of the statistical quartiles Q1 and Q3 (vertical dashes) and mean value within the range (red bold dots). \href{http://www.gnuplot.info/}{Gnuplot} visualization­}
		\label{fig:A.28}
	\end{center}
\end{figure}
\pagebreak

%%% one more Gretl start
\begin{table}[H]
	\rowcolors{1}{BlanchedAlmond}{MintCream}
	\begin{center}
	\caption{Results of the \textit{Prais-Winsten estimation} applied towards variables 15 from the selected dataset 151-175 of the spectral reflectance of \textit{P.oceanica}. \href{http://gretl.sourceforge.net/}{Gretl}}
	\label{tab:26}
		Model 3: Prais--Winsten, using observations 1:01--8:23 ($T$ = 191)\\
		Dependent variable: v12\\
	$\hat{\rho}$ = -0.0343656

	\vspace{1em}

	\begin{tabular}{|c c c c c|}
		  & {Coefficient} &  {Std.\ Error} & {$t$-ratio} &   {p-value} \\[1ex]
		const &   $-$0.00498439 &     0.00977709 &       $-$0.5098 &         0.6108 \\
		v2 &   0.912742 &     0.205904 &      4.4328 &         0.0000 \\
	\end{tabular}

	\vspace{1em}Statistics based on the rho-differenced data:

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Mean dependent var &  0.035372 & S.D. dependent var &  0.053185 \\
		Sum squared resid &  0.490951 & S.E. of regression &  0.050967 \\
		$R^2$ &  0.086511 & Adjusted $R^2$ &  0.081678 \\
		$F(1, 189)$ &  19.67163 & P-value($F$) &  0.000016 \\
		$\hat{\rho}$ & $-$0.001888 & Durbin--Watson &  2.003741 \\
	\end{tabular}
	\end{center}
\end{table}
%%%one more Gretl end

%%%Gretl begin
\begin{table}[H]
	\rowcolors{1}{Gainsboro}{MintCream}
	\begin{center}
	\caption{Results of the \textit{Weighted Least Squares Analysis} applied towards variables 12-20 from the selected dataset 301-325 of the spectral reflectance of \textit{P.oceanica}. \href{http://gretl.sourceforge.net/}{Gretl}}
		Model 1: WLS, using observations 1--95\\
		Dependent variable: v20\\
		Variable used as weight: v12

	\vspace{1em}
	\begin{tabular}{|c|c|c|c|c|}
		  & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
		const &   0.239525 &     0.0771338 & 3.1053 &         0.0025 \\
		v1 &   0.000120809 &     0.000147131 &       0.8211 &         0.4137 \\
	\end{tabular}

	\vspace{1em}Statistics based on the weighted data:

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Sum squared resid &  0.290924 & S.E. of regression &  0.055930 \\
		$R^2$ &  0.007197 & Adjusted $R^2$ & -0.003478 \\
		$F(1, 93)$ &  0.674200 & P-value($F$) &  0.413691 \\
		Log-likelihood &  140.1579 & Akaike criterion & $-$276.3158 \\
		Schwarz criterion & $-$271.2081 & Hannan--Quinn & $-$274.2519 \\
	\end{tabular}

	\vspace{1em}Statistics based on the original data:

	\vspace{1ex}
	\begin{tabular}{lrlr}
		Mean dependent var &  0.136154 & S.D. dependent var &  0.149193 \\
		Sum squared resid &  5.689669 & S.E. of regression &  0.247344 \\
	\end{tabular}
	\label{tab:27}
	\end{center}
\end{table}
%%%Gretl end
\pagebreak

\begin{figure}[H]
	\begin{center}
	\includegraphics[scale=0.4]{G-autocorr-sand51-75.jpg}
	\caption{Graph of the autocorrelation analysis of the measurements of carbonate sand. Series 51-75, variable 27. \href{http://gretl.sourceforge.net/}{Gretl} visualization­}
	\label{fig:A.29}
	\end{center}
\end{figure}

%%% Gretl start
\begin{table}[htbp]
	\rowcolors{1}{Ivory}{GhostWhite}
	\caption{Results of the \textit{autocorrelation analysis} of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
	\begin{center}
	\begin{tabular}{|c c c c c c c||}
		\hline\hline
		LAG & ACF & & PACF & & Q-stat & [p-value] \\ \\ \hline\hline
		1  & 0.9851 & ***  & 0.9851& ***   & 104.8293 & [0.000]\\ \hline
		    2  & 0.9657  &***  &-0.1598  &   ***  &  206.5423  &[0.000]\\ \hline
		    3  & 0.9420  &***  &-0.1337  &  ***   &  304.2787  &[0.000]\\ \hline
		    4  & 0.9144  &***  &-0.1107  &  ***   &  397.2949  &[0.000]\\ \hline
		    5  & 0.8835  &***  &-0.0905  &  ***   &  484.9833  &[0.000]\\ \hline
		    6  & 0.8495 & ***  &-0.0732  &  ***   &  566.8791  &[0.000]\\ \hline
		    7  & 0.8131 & ***  &-0.0565  &  ***   &  642.6707  &[0.000]\\ \hline
		    8  & 0.7748 & ***  &-0.0423  &  ***   &  712.1933  &[0.000]\\ \hline
		    9  & 0.7350 & ***  &-0.0317  &  ***   &  775.4114  &[0.000]\\ \hline
		   10 &  0.6941&  *** & -0.0280 &  ***   &   832.3807 & [0.000]\\ \hline
		   11 &  0.6524 & *** & -0.0221 &  ***   &   883.2508 & [0.000]\\ \hline
		   12 &  0.6103 & *** & -0.0197 &  ***   &   928.2405 & [0.000]\\ \hline
		   13 &  0.5679 & *** & -0.0170 &  ***   &   967.6289 & [0.000]\\ \hline
		   14 &  0.5256 & *** & -0.0162 &  ***   &  1001.7381&  [0.000]\\ \hline
		   15 &  0.4835 & *** & -0.0154 &  ***   &  1030.9237 & [0.000]\\ \hline
		   16 & 0.4418 & *** & -0.0172  &  ***  &  1055.5584 & [0.000]\\ \hline
		   17  & 0.4v004 & *** & -0.0196 &***     &  1076.0231 & [0.000]\\ \hline
		   18 & 0.35v94 & *** & -0.0216  & ***   &  1092.7041 & [0.000]\\ \hline
		   19 &  0.3189 & *** & -0.0224   & ***  &  1105.9911 & [0.000]\\ \hline
		   20 &  0.2788 & *** & -0.0269   & ***  &  1116.2679 & [0.000]\\ \hline
	\end{tabular}
	\end{center}
	\label{tab:28}
\end{table}
%%% Gretl end
\pagebreak

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-QQplot-sand.jpg}
		\caption{Normal Q-Q plot: estimated versus observed values of the measurement of carbonate sand, variable 27. Series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
	\label{fig:A.30}
	\end{center}
\end{figure}

%%% Gretl start
\begin{table}[H]
	\caption{Results of the \textit{Quantile estimates}: measurements 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
	\begin{center}
		Model 2: Quantile estimates, using observations 1:01--5:09 ($T$ = 105)\\
		Dependent variable: v27\\
		$\tau$ = 0.5\\
		Asymptotic standard errors assuming IID errors\\

\vspace{1em}

\begin{tabular}{c c c c c}
	***   & {Coefficient} & {Std.\ Error} & {$t$-ratio} & {p-value} \\[1ex]
	const &  0.837827 &    0.0386417 &      21.6820 &         0.0000 \\
	v1 &   $-$0.00111120 &     6.60949\textrm{e--005} &       $-$16.8122 &         0.0000 \\ \hline \hline
\end{tabular}

\vspace{1ex}
	\begin{tabular}{lrlr}
		Median depend. var &  0.191024 & S.D. dependent var &  0.135439 \\
		Sum absolute resid &  8.625039 & Sum squared resid &  1.370360 \\
		Log-likelihood &  84.64508 & Akaike criterion & $-$165.2902 \\
		Schwarz criterion & $-$159.9822 & Hannan--Quinn & $-$163.1393 \\ \hline \hline
	\end{tabular}
	\end{center}
\label{tab:29}
\end{table}
%%% Gretl end
\pagebreak

%%%Gretl begin
\begin{table}[H]
	\begin{small}
	\rowcolors{1}{Ivory}{LightGoldenrodYellow}
	\caption{Results of the \textit{periodogram} for v30 of the measurement set 51-75 of the spectral reflectance of \textit{carbonate sand}, \href{http://gretl.sourceforge.net/}{Gretl}}
	\label{tab:30}
	\begin{center}
	\begin{tabular}{|p{15mm} p{15mm} p{15mm} p{15mm} || p{15mm} p{15mm}p{15mm} p{15mm}|}
		omega & scaled frequency & periods & log spectral density & omega & scaled frequency & periods & log spectral density \\ \\ \hline\hline
		0.05984      & 1    &      105.00  &      -1.2466  &  1.61568   &   27    &        3.89     &    -11.111\\ 
		 0.11968     &  2   &        52.50   &      -5.8761 &   1.67552   &   28    &        3.75     &    -11.170\\ 
		 0.17952     &  3   &        35.00   &      -4.0657 &  1.73536   &   29    &        3.62     &    -11.208\\ 
		 0.23936     &  4   &        26.25   &      -7.1031 &  1.79520   &   30    &        3.50     &    -11.461\\ 
		 0.29920     &  5   &        21.00   &      -6.1742 &  1.85504   &   31    &        3.39     &    -11.380\\ 
		 0.35904     &  6   &        17.50   &      -7.2772  & 1.91488    & 32     &       3.28      &   -11.686\\ 
		 0.41888     &  7     &      15.00   &      -8.7802 &  1.97472    &  33    &        3.18     &    -11.624\\ 
		 0.47872     &  8     &      13.13   &      -7.6806 & 2.03456    &  34    &        3.09         &-11.724\\ 
		 0.53856     &  9     &      11.67   &      -8.4848  & 2.09440    &  35    &        3.00    &     -11.818\\ 
		 0.59840     & 10    &       10.50  &       -8.5764 &    2.15423    &  36    &        2.92    &     -11.805\\ 
		 0.65824     & 11    &        9.55   &      -9.4643 &  2.21407    &  37    &        2.84    &     -12.002\\ 
		 0.71808     & 12    &        8.75   &      -9.0868 &  2.27391    &  38    &        2.76    &     -11.930\\ 
		 0.77792     & 13    &        8.08   &      -9.3608 &  2.33375    &  39    &        2.69    &     -12.010\\ 
		 0.83776     & 14    &        7.50     &    -9.9332 &  2.39359    &  40    &        2.63    &     -12.124\\ 
		 0.89760     & 15    &        7.00     &    -9.6836 &  2.45343    &  41    &        2.56    &     -12.190\\ 
		 0.95744     & 16    &        6.56     &    -10.00  & 2.51327    &  42           & 2.50    &     -12.248\\ 
		 1.01728     & 17    &        6.18     &    -9.8994 & 2.57311    &  43      &      2.44    &     -12.306\\ 
		 1.07712     & 18    &        5.83     &    -10.188 & 2.63295    &  44      &      2.39    &     -12.323\\ 
		 1.13696     & 19    &        5.53     &    -10.197 & 2.69279    &  45      &      2.33    &     -12.353\\ 
		 1.19680     & 20    &        5.25     &    -10.467 &  2.75263    &  46      &      2.28    &     -12.420\\ 
		 1.25664     & 21    &        5.00     &    -10.505 &  2.81247    &  47      &      2.23    &     -12.382\\ 
		 1.31648     & 22    &        4.77    &    -10.641 &  2.87231    &  48      &      2.19    &     -12.487\\ 
		 1.37632     & 23    &        4.57     &    -10.906 &  2.93215    &  49      &    2.14      &   -12.433\\ 
		 1.43616     & 24    &        4.38     &    -10.957 &  2.99199    &  50        &    2.10    &     -12.609\\ 
		 1.49600      &25    &        4.20     &    -10.785 &   3.05183    &  51      &      2.06    &     -12.410\\ 
		 3.11167    &  52      &      2.02    &     -12.477 & & & & \\   \hline \hline
	\end{tabular}
	\end{center}
	\end{small}
	Number of observations = 105
\end{table}
%%% Gretl end

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-period-sand30.jpg}
		\caption{Plot illustrating periodogram for v30 from measurements of carbonate sand, series 51-75. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}\label{fig:A.31}
	\end{center}	
\end{figure}

%%%Begin Plot Gretl
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-polynom-trend-wo.jpg}
		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater without sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
	\end{center}
	\label{fig:A.32}
\end{figure}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-polynom-trend-with.jpg}
		\caption{Plot illustrating polynomial trend for spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
	\end{center}
	\label{fig:A.33}
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.4]{G-exp-mov-aver.jpg}
		\caption{Exponential moving average of spectral reflectance of seawater with sediments (Ligaria Beach), variable V15. \href{http://gretl.sourceforge.net/}{Gretl} modelling visualization­}
	\end{center}
	\label{fig:A.34}
\end{figure}
\pagebreak
%%%end Plot Gretl

\rowcolors{1}{OldLace}{FloralWhite}
\begin{longtable}{|r|r|r|r|r|r|r|}
	\caption{Results of the statistical analysis of spectral reflectance of \textbf{carbonate sand}, with average values (for sets 1-3). Wavelength step: 3 nm. Measured on Agia Pelagia beach, 14th October}\label{tab:A.29} \\
 		 \multicolumn{1}{|l|}{\textbf{wl}} &
  		 \multicolumn{1}{l|}{\textbf{min}} & 
  		 \multicolumn{1}{l|}{\textbf{Q1}} & 
  		 \multicolumn{1}{l|}{\textbf{mean}} & 
 		  \multicolumn{1}{l|}{\textbf{Q3}} & 
 		  \multicolumn{1}{l|}{\textbf{max}} & 		
 		  \multicolumn{1}{c|}{\textbf{median}} \\ \hline
	\endfirsthead

	\multicolumn{7}{c}{Table A.29: Spectral reflectance of \textbf{carbonate sand} \textit{(continued from previous page)}}\\
	\hline \multicolumn{1}{|l|}{\textbf{wl}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l|}{\textbf{Q1}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}				{\textbf{Q3}} & \multicolumn{1}{l|}{\textbf{max}} & \multicolumn{1}{l|}{\textbf{median}} \\ \hline
	\endhead

	402 & 0.091522 & 0.108252 & 0.119264 & 0.125199 & 0.156262 & 0.120855 \\ \hline
	405 & 0.094129 & 0.111507 & 0.126390 & 0.139560 & 0.174126 & 0.124623 \\ \hline
	408 & 0.096608 & 0.114474 & 0.129816 & 0.143488 & 0.178780 & 0.127983 \\ \hline
	412 & 0.098703 & 0.117396 & 0.133108 & 0.147164 & 0.183246 & 0.131282 \\ \hline
	415 & 0.101803 & 0.120880 & 0.137375 & 0.151792 & 0.189039 & 0.135571 \\ \hline
	418 & 0.105012 & 0.124746 & 0.141910 & 0.156937 & 0.195073 & 0.140136 \\ \hline
	422 & 0.108640 & 0.129140 & 0.147168 & 0.162721 & 0.202021 & 0.145337 \\ \hline
	425 & 0.113279 & 0.134815 & 0.153679 & 0.169999 & 0.210668 & 0.151906 \\ \hline
	428 & 0.118544 & 0.141149 & 0.160985 & 0.178059 & 0.220194 & 0.159402 \\ \hline
	432 & 0.122987 & 0.146844 & 0.167659 & 0.185518 & 0.228841 & 0.166154 \\ \hline
	435 & 0.127911 & 0.152289 & 0.174248 & 0.192960 & 0.237370 & 0.172836 \\ \hline
	438 & 0.132310 & 0.158310 & 0.181427 & 0.201169 & 0.246705 & 0.180044 \\ \hline
	442 & 0.136621 & 0.163674 & 0.187881 & 0.208310 & 0.255343 & 0.186472 \\ \hline
	445 & 0.140316 & 0.168266 & 0.193508 & 0.214558 & 0.262974 & 0.192106 \\ \hline
	448 & 0.143322 & 0.172017 & 0.198207 & 0.219703 & 0.269568 & 0.196709 \\ \hline
	452 & 0.145645 & 0.174991 & 0.201955 & 0.223770 & 0.274973 & 0.200367 \\ \hline
	455 & 0.148101 & 0.178031 & 0.205747 & 0.227926 & 0.280378 & 0.204052 \\ \hline
	459 & 0.149944 & 0.180646 & 0.208895 & 0.230917 & 0.284950 & 0.207175 \\ \hline
	462 & 0.151978 & 0.183204 & 0.212228 & 0.234461 & 0.289839 & 0.210556 \\ \hline
	465 & 0.136264 & 0.184985 & 0.214436 & 0.238727 & 0.294286 & 0.213490 \\ \hline
	469 & 0.155248 & 0.187804 & 0.217729 & 0.240237 & 0.298004 & 0.215926 \\ \hline
	472 & 0.156554 & 0.189708 & 0.220128 & 0.242769 & 0.301445 & 0.218291 \\ \hline
	475 & 0.158239 & 0.192110 & 0.223000 & 0.246279 & 0.305677 & 0.221116 \\ \hline
	479 & 0.160611 & 0.195242 & 0.226849 & 0.250810 & 0.310961 & 0.224984 \\ \hline
	482 & 0.164012 & 0.200152 & 0.232478 & 0.257229 & 0.318555 & 0.230599 \\ \hline
	485 & 0.168472 & 0.205823 & 0.238938 & 0.264573 & 0.327001 & 0.237010 \\ \hline
	489 & 0.172449 & 0.211132 & 0.245045 & 0.271446 & 0.334981 & 0.243087 \\ \hline
	492 & 0.176260 & 0.216078 & 0.250975 & 0.278269 & 0.342800 & 0.249133 \\ \hline
	495 & 0.180950 & 0.222148 & 0.258178 & 0.286469 & 0.352048 & 0.256243 \\ \hline
	499 & 0.186009 & 0.228774 & 0.265843 & 0.295200 & 0.361896 & 0.263884 \\ \hline
	502 & 0.189596 & 0.233599 & 0.271417 & 0.301563 & 0.368882 & 0.269139 \\ \hline
	505 & 0.191587 & 0.236300 & 0.274828 & 0.305379 & 0.372817 & 0.272281 \\ \hline
	509 & 0.193259 & 0.238538 & 0.277796 & 0.309446 & 0.376347 & 0.275357 \\ \hline
	512 & 0.196153 & 0.242620 & 0.282441 & 0.315177 & 0.381948 & 0.279890 \\ \hline
	515 & 0.200015 & 0.248183 & 0.288410 & 0.322126 & 0.389945 & 0.285937 \\ \hline
	519 & 0.204139 & 0.253981 & 0.294768 & 0.329524 & 0.398796 & 0.291707 \\ \hline
	522 & 0.208548 & 0.259760 & 0.301397 & 0.337066 & 0.407969 & 0.298231 \\ \hline
	526 & 0.213382 & 0.266241 & 0.308821 & 0.345426 & 0.418276 & 0.305397 \\ \hline
	529 & 0.218205 & 0.272715 & 0.316136 & 0.353749 & 0.428329 & 0.312514 \\ \hline
	532 & 0.222568 & 0.278565 & 0.322872 & 0.361518 & 0.437755 & 0.319122 \\ \hline
	536 & 0.226789 & 0.284495 & 0.329470 & 0.369010 & 0.446804 & 0.325558 \\ \hline
	539 & 0.230711 & 0.289790 & 0.335482 & 0.376082 & 0.455017 & 0.331317 \\ \hline
	542 & 0.233458 & 0.293802 & 0.340066 & 0.381589 & 0.461345 & 0.335918 \\ \hline
	546 & 0.235193 & 0.296532 & 0.343224 & 0.385560 & 0.465598 & 0.338854 \\ \hline
	549 & 0.236680 & 0.299381 & 0.345918 & 0.388909 & 0.469140 & 0.341256 \\ \hline
	552 & 0.237769 & 0.301722 & 0.348197 & 0.391858 & 0.472296 & 0.343234 \\ \hline
	556 & 0.239247 & 0.304604 & 0.350951 & 0.395307 & 0.476036 & 0.345633 \\ \hline
	559 & 0.240209 & 0.306984 & 0.353309 & 0.398254 & 0.479382 & 0.347674 \\ \hline
	562 & 0.240654 & 0.308165 & 0.354596 & 0.400054 & 0.481185 & 0.348907 \\ \hline
	566 & 0.240416 & 0.308874 & 0.355162 & 0.401224 & 0.481897 & 0.348929 \\ \hline
	569 & 0.238748 & 0.307064 & 0.353688 & 0.400427 & 0.480263 & 0.347042 \\ \hline
	572 & 0.235026 & 0.302640 & 0.349433 & 0.395980 & 0.475648 & 0.342958 \\ \hline
	576 & 0.212260 & 0.294628 & 0.340050 & 0.386378 & 0.466085 & 0.334751 \\ \hline
	579 & 0.219835 & 0.283193 & 0.329664 & 0.375185 & 0.450895 & 0.324215 \\ \hline
	583 & 0.208955 & 0.268530 & 0.314599 & 0.359047 & 0.432720 & 0.308769 \\ \hline
	586 & 0.196370 & 0.252380 & 0.296960 & 0.338330 & 0.411550 & 0.290283 \\ \hline
	589 & 0.181828 & 0.233326 & 0.276049 & 0.316609 & 0.386565 & 0.267500 \\ \hline
	593 & 0.166276 & 0.212916 & 0.253544 & 0.293790 & 0.359268 & 0.243094 \\ \hline
	596 & 0.149323 & 0.191055 & 0.229373 & 0.267668 & 0.331353 & 0.217166 \\ \hline
	599 & 	0.132182 & 0.169324 & 0.205072 & 0.241088 & 0.303079 & 0.191137 \\ \hline
	603 & 0.117895 & 0.151271 & 0.185181 & 0.219420 & 0.280528 & 0.170074 \\ \hline
	606 & 0.107449 & 0.138360 & 0.171146 & 0.203897 & 0.264634 & 0.155317 \\ \hline
	609 & 0.101008 & 0.130206 & 0.162388 & 0.194168 & 0.254754 & 0.146269 \\ \hline
	613 & 0.097084 & 0.125265 & 0.157073 & 0.188174 & 0.248485 & 0.140671 \\ \hline
	616 & 0.094383 & 0.121987 & 0.153592 & 0.184349 & 0.244702 & 0.137070 \\ \hline
	619 & 0.092105 & 0.119159 & 0.150566 & 0.180848 & 0.241217 & 0.134371 \\ \hline
	623 & 0.090017 & 0.116792 & 0.147899 & 0.177856 & 0.238265 & 0.131097 \\ \hline
	626 & 0.088364 & 0.114636 & 0.145734 & 0.175448 & 0.236222 & 0.128638 \\ \hline
	629 & 0.086562 & 0.112565 & 0.143418 & 0.173135 & 0.233625 & 0.126183 \\ \hline
	633 & 0.084448 & 0.109785 & 0.140489 & 0.170056 & 0.230251 & 0.123296 \\ \hline
	636 & 0.082208 & 0.106766 & 0.137216 & 0.166497 & 0.226502 & 0.119990 \\ \hline
	639 & 0.079925 & 0.103852 & 0.134122 & 0.162951 & 0.222721 & 0.116886 \\ \hline
	643 & 0.077879 & 0.101395 & 0.131468 & 0.159978 & 0.219899 & 0.114099 \\ \hline
	646 & 0.075563 & 0.098621 & 0.128409 & 0.156518 & 0.216573 & 0.110812 \\ \hline
	649 & 0.072328 & 0.095004 & 0.124254 & 0.151735 & 0.212143 & 0.106456 \\ \hline
	653 & 0.068455 & 0.089876 & 0.118539 & 0.145085 & 0.205574 & 0.100544 \\ \hline
	656 & 0.063138 & 0.083138 & 0.110767 & 0.135900 & 0.196204 & 0.092816 \\ \hline
	660 & 0.057899 & 0.076562 & 0.102919 & 0.126826 & 0.186379 & 0.085030 \\ \hline
	663 & 0.054237 & 0.071651 & 0.097286 & 0.120346 & 0.179417 & 0.079514 \\ \hline
	666 & 0.051717 & 0.068781 & 0.093827 & 0.116423 & 0.175195 & 0.076101 \\ \hline
	670 & 0.050172 & 0.066690 & 0.091590 & 0.113854 & 0.172561 & 0.073923 \\ \hline
	673 & 0.048823 & 0.064991 & 0.089610 & 0.111550 & 0.170358 & 0.071985 \\ \hline
	676 & 0.047230 & 0.063165 & 0.087474 & 0.109078 & 0.167774 & 0.069807 \\ \hline
	680 & 0.045818 & 0.061108 & 0.085168 & 0.106421 & 0.165049 & 0.067486 \\ \hline
	683 & 0.044262 & 0.058946 & 0.082783 & 0.103630 & 0.162579 & 0.065041 \\ \hline
	686 & 0.041986 & 0.055955 & 0.079361 & 0.099441 & 0.158648 & 0.061599 \\ \hline
	690 & 0.040081 & 0.051396 & 0.074010 & 0.092741 & 0.152034 & 0.055896 \\ \hline
	693 & 0.034783 & 0.045555 & 0.067241 & 0.084820 & 0.143195 & 0.050116 \\ \hline
	696 & 0.030094 & 0.039507 & 0.060070 & 0.076166 & 0.133838 & 0.043600 \\ \hline
	700 & 0.025499 & 0.033281 & 0.052660 & 0.066929 & 0.124164 & 0.036937 \\ \hline
	703 & 0.020752 & 0.027050 & 0.045008 & 0.057197 & 0.113843 & 0.030173 \\ \hline
	706 & 0.016243 & 0.021080 & 0.037529 & 0.047681 & 0.103394 & 0.023549 \\ \hline
	710 & 0.012303 & 0.015914 & 0.030791 & 0.038860 & 0.093244 & 0.017834 \\ \hline
	713 & 0.007533 & 0.011641 & 0.025113 & 0.031338 & 0.084822 & 0.013233 \\ \hline
	716 & 0.006371 & 0.008236 & 0.020172 & 0.024328 & 0.077202 & 0.009214 \\ \hline
	720 & 0.004247 & 0.005386 & 0.015711 & 0.018057 & 0.069074 & 0.006121 \\ \hline
	723 & 0.002773 & 0.003524 & 0.012159 & 0.012999 & 0.061638 & 0.003836 \\ \hline
	726 & 0.001812 & 0.002194 & 0.009537 & 0.009237 & 0.055408 & 0.002543 \\ \hline
	730 & 0.001248 & 0.001453 & 0.007627 & 0.006800 & 0.049434 & 0.001739 \\ \hline
	733 & 0.000801 & 0.001018 & 0.006264 & 0.004722 & 0.044850 & 0.001208 \\ \hline
	736 & 0.000192 & 0.000855 & 0.005279 & 0.003743 & 0.042097 & 0.000954 \\ \hline
	740 & 0.000471 & 0.000718 & 0.004811 & 0.003055 & 0.039099 & 0.000816 \\ \hline
	743 & 0.000506 & 0.000644 & 0.004419 & 0.002901 & 0.037332 & 0.000740 \\ \hline
	746 & 0.000211 & 0.000582 & 0.004282 & 0.002898 & 0.036089 & 0.000715 \\ \hline
	750 & 0.000318 & 0.000557 & 0.003937 & 0.002175 & 0.033962 & 0.000667 \\ \hline
\end{longtable}

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{Fig-26.jpg}
	\caption{Spectral reflectance of carbonate sand on A.Pelagia beach. Results of single measurement set made by spectroradiometer Trios-RAMSES. \href{http://www.gnuplot.info/}{Gnuplot}}
	\label{fig:A.35}
\end{figure}
\pagebreak
\newpage

\subsection{\textit{Appendix-8}. Satellite and aerial images covering research area of Crete island: selected examples}
%\appendixname{\textbf{8:\\ Satellite and aerial images covering research area of Crete island: selected examples}}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.20]{Locations_Meas.jpg}
		\caption{Locations of selected measurements, visualization on Google Earth­}\label{fig:A.36}
	\end{center}	
\end{figure}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{App-8.jpg}
		\caption{Random mosaic of selected aerial Google Earth images}\label{fig:A.37}	
	\end{center}	
\end{figure}
\pagebreak

\begin{table}[H]
	\caption{Available broadband Landsat satellite images covering the research area of Crete island}\label{tab:34}
	\centering
	\rowcolors{1}{Ivory}{GhostWhite}
	 \begin{tabular}{| p{1cm} | p{3cm} | p{2cm} | p{5cm} |}
	    \hline
		    \textbf{No} & \textbf{Image cource} & \textbf{Data} & \textbf{Name} \\ \hline \hline
		     1 & Landsat ETM+ & 2005/May/04 & WRS2p181r035L71181035-035-20050504-ETM-GLS2005 \\ \hline
		     2 & Landsat TM & 2006/Nov/07 & WRS2p181r036L5181036-036-20061107-TM-GLS2005 \\ \hline
		     3 & Landsat ETM+ & 2005/Apr/25 & WRS2p182r035L71182035-035-20050425-ETM-GLS2005 \\ \hline
		     4 & Landsat ETM+ & 2000/Jul/09 & WRS2p181r036-7dx-20000709-ETM-GLS2000 \\ \hline
		     5 & Landsat TM & 1987/Jun/10 & LandsatWRS2p183r035p183r035-5dx-19870610-TM-GLS1990 \\ \hline
		     6 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-261Mosaic-LandsatN-35N-35-35ETM-EarthSat-MrSID-19990808-20020624 \\ \hline
		     7 & Landsat ETM+/ Earth Sat & 1999/Aug/08 & 071-260Mosaic-LandsatN-35N-35-30ETM-EarthSat-MrSID-19990808-20020617 \\ \hline
		     8 & Landsat ETM+ & 2000/Jun/30 & WRS2p182r036-7x-20000630-ETM-EarthSat \\ \hline
		     9 & Landsat MSS / Earth Sat & 1975/Jul/26 & LandsatWRS1p196r35-2m-19750726-MSS-EarthSat \\ \hline
		    10 & Landsat TM / Earth Sat & 1987/Jun/10 & 012-807LandsatWRS2p183r35-5t-19870610-TM-EarthSat \\ \hline
		    11 & Landsat ETM+ & 2000/Jun/30 & LandsatWRS2p182r036-7dx-20000630-ETM-GLS2000 \\ \hline
		    12 & Landsat ETM+ & 2005/Apr/09 & LandsatWRS2p182r036L71182036-036-20050409-ETM-GLS2005 \\ \hline
	  \end{tabular}	
\end{table}
\pagebreak
\newpage

\begin{figure}[H]
	\centering
	\subfloat [Landsat 2006-11-07-TM]{\includegraphics[width=0.3\textwidth]{Landsat1.jpg}}
	\hspace{1mm}
	\subfloat [Landsat-2005-04-09]{\includegraphics[width=0.3\textwidth]{Landsat2.jpg}}
	\hspace{1mm}
	\subfloat [Mosaic-Landsat- ETM-1999-08-08]{\includegraphics[width=0.3\textwidth]{Landsat3.jpg}}
	\hspace{1mm}
	\subfloat[Landsat 2000-07-09 ETM]{\includegraphics[width=0.3\textwidth]{Landsat4.jpg}}
	\hspace{1mm}
	\subfloat [Landsat 2000-06-30 ETM]{\includegraphics[width=0.3\textwidth]{Landsat5.jpg}}
	\hspace{1mm}
	\subfloat [Landsat 1987-06-10-TM]{\includegraphics[width=0.3\textwidth]{Landsat6.jpg}}
	\hspace{1mm}
	\subfloat [Landsat-2005-05-04-ETM]{\includegraphics[width=0.3\textwidth]{Landsat7.jpg}}
	\hspace{1mm}
	\subfloat [Landsat-1975-07-26]{\includegraphics[width=0.3\textwidth]{Landsat8.jpg}}
	\hspace{1mm}
	\subfloat [Landsat TM]{\includegraphics[width=0.3\textwidth]{Landsat9.jpg}}
	\hspace{1mm}
	\subfloat [Mosaic-Landsat-ETM-1999-08-08]{\includegraphics[width=0.3\textwidth]{Landsat10.jpg}}
	\caption{Landsat imagery, Crete island. Previews}
	\label{fig:A.38}
\end{figure}
\pagebreak
\newpage

\subsection{\textit{Appendix-9}. Analysis and classification of the satellite and aerial images}

\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.23]{OpenEV-2.jpg}
		\caption{The point querying shows the selected points and their coordinates within the area of seagrass meadow (green). \href{http://openev.sourceforge.net/}{OpenEV}}\label{fig:A.39}	
	\end{center}	
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.23]{OpenEV-3.jpg}
	\caption{\textit{Logarithmic Enhancement to Raster}, applied to the aerial Google Earth image: the seagrass meadow can now be easily seen as a bright spot of purple color. \href{http://openev.sourceforge.net/}{OpenEV}}\label{fig:A.40}	
	\end{center}	
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{OpenEV-1.jpg}
		\caption{Raster properties dialog: visualization and spatial info about the image (projection UTM, zone 35, datum WGS-84, etc.)  \href{http://openev.sourceforge.net/}{OpenEV})}\label{fig:A.41}	
	\end{center}	
\end{figure}
\begin{figure}[H]
	\begin{center}
		\includegraphics[scale=0.25]{Ilwis-2.jpg}
		\caption{Color composite image composed of 3 images of Cretan shelf, Google Earth. \href{http://www.ilwis.org/}{ILWIS}}\label{fig:A.42}	
	\end{center}	
\end{figure}
\pagebreak

\subsection{\textit{Appendix-10}. Research general workflow}
\begin{figure}[H]
	\centering
	\includegraphics[scale=0.57]{UML_GIS.png}
	\caption{General methodological research approach. \href{http://inkscape.org/}{Inkscape}}
	\label{fig:1.11}
\end{figure}

\end{appendices}
\pagebreak
\clearpage

\begin{thebibliography}{999}
\addcontentsline{toc}{section}{\textbf{Bibliography}}\label{bib}
\bibliographystyle{plainnat}
\pagenumbering{alph}
\markright{Bibliography}

\bibitem{Ackleson2003}Ackleson S.\index[people]{} 2003. \emph{Light in shallow waters: A brief research review}. Limnol. Oceanogr.,
48(1, part 2), 2003, 323–328. American Society of Limnology and Oceanography, Inc. \pageref{Ackleson, 2003}
\bibitem{Alcoverro1997} %
Alcoverro T.\index[people]{Alcoverro, T.}, Romero J.\index[people]{Romero J.}, Duarte C.M.\index[people]{Duarte C.M.}, Lopez N.\index[people]{Lopez, N.} 1997. \emph{Spatial and temporal variations
in nutrient limitation of seagrass \textit{Posidonia oceanica} growth in the NW Mediterranean}.
Marine ecology Progress Series.Vol.146: 155-161. \pageref{}
\bibitem{Amoutzopoulou-Schina05}Amoutzopoulou-Schina H.\index[people]{Amoutzopoulou-Schina, H.} and Haritonidis S.\index[people]{Haritonidis, S.} 2005. \emph{Distribution and phenology of the marine phanerogam \textit{Posidonia oceanica} in the Pagassitikos Gulf, Greece}. Journal of
Biological Research 4: 203 – 211, 2005 J. \pageref{Amoutzopoulou-Schina05}
\bibitem{Arai05}
Arai K.\index[people]{Arai, K.} and Tonooka H.\index[people]{Tonooka, H.} 2005. \emph{Radiometric Performance Evaluation of \ac{ASTER} VNIR, \ac{SWIR} and \ac{TIR}} 
IEEE Transactions on Geoscience and Remote Sensing, Vol. 43, No. 12, Dec. 2005. \pageref{Arai05}
\bibitem{Ardizzone06}
Ardizzone G.D.\index[people]{Ardizzone G.D.}, Belluscio A.\index[people]{Belluscio A.} and Maiorano L.\index[people]{Maiorano, L.} 2006. \emph{Long-term change in the structure of
a \textit{Posidonia oceanica} landscape and its reference for a monitoring plan}. Marine Ecology.
ISSN 0173-9565. \pageref{Ardizzone06}
\bibitem{Ball09}
Ball D.\index[people]{Ball, D.}, Soto-Berelov M.\index[people]{Soto-Berelov, M.}, Young P.\index[people]{Young, P.} and Coots A.\index[people]{Coots, A.} 2009. \emph{Baywide Seagrass Monitoring Program - Historical Seagrass Mapping}. Fisheries Victoria Technical Report Series No. 70. \pageref{Ball09}
\bibitem{Barille09}
Barille L.\index[people]{Barille, L.}, Robin M.\index[people]{Robin, M.}, Harin N.\index[people]{Harin, N.}, Bargain A.\index[people]{Bargain, A.}, Patrick L.\index[people]{Patrick, L.} 2009. \emph{Increase in seagrass distribution at Bourgneuf Bay (France) detected by spatial remote sensing}. Aquatic Botany. \pageref{Barille09}
\bibitem{Bakran-Petricioli06} %
Bakran-Petricioli T.\index[people]{Bakran-Petricioli, T.}, Antonic O.\index[people]{Antonic, O.}, Bukovec D.\index[people]{Bukovec, D.}, Petricioli D.\index[people]{Petricioli, D.}, Janekovic I.\index[people]{Janekovic, I.}, Krizan J.\index[people]{Krizan, J.}, Kusan V.\index[people]{Kusan, V.} and Dujmovic S.\index[people]{Dujmovic, S.} 2006. \emph{Modelling spatial distribution of the Croatian marine
benthic habitats}. Ecological Modelling 191, 96–105. \pageref{Bakran-Petricioli06}
\bibitem{Bell99}
Bell S.S.\index[people]{Bell S.S.}, Robbins B.D.\index[people]{Robbins B.D.} and Jensen S.L.\index[people]{Jensen S.L.} 1999. \emph{Gap Dynamics in a Seagrass Landscape}.
Springer-Verlag. Ecosystems (1999) 2: 493–504. \pageref{Bell99}
\bibitem{Bierwirth93}
Bierwirth P. N.\index[people]{Bierwirth, P. N.}, Lee T. J.\index[people]{Lee, T. J.} and Burne R. V.\index[people]{Burne, R.V.} 1993. \emph{Shalow sea-floor reflectance and water
depth derived by unmixing multispectral imagery}. Photogrammetric Engineering and Remote
Sensing, vol. 59, No. 3, March, pp. 331-338. \pageref{Bierwirth93}
\bibitem{Blake00}
Blake S.\index[people]{Blake, S.}, Roob R.\index[people]{Roob, R.} and Patterson E.\index[people]{Patterson, E.} 2000. \emph{Seagrass mapping of Victoria's Minor Inlets}.
Marine and Freshwater Resources Institute Report No. 28. \pageref{Blake00}
\bibitem{Borum04}
Borum J.\index[people]{Borum, J.}, Duarte C.M.\index[people]{Duarte C.M.}, Krause-Jensen D.\index[people]{Krause-Jensen, D.} and Greve T.M.\index[people]{Greve, T.M} 2004. \emph{European seagrasses: an introduction to monitoring and management}. The M and MS project. ISBN: 87-89143-21-3.
Internet version: Available at \href{http://www.seagrasses.org}{Seagrasses.org} \pageref{Borum04}
\bibitem{Brown02}
Brown C.J.\index[people]{Brown, C.J.}, Cooper K.M.\index[people]{Cooper, K.M.}, Meadows W.J.\index[people]{Meadows, W.J.}, Limpenny D.S.\index[people]{Limpenny, D.S.}, Rees H.L.\index[people]{Rees, H.L.} 2002. \emph{Small scale mapping of sea bed assemblages in the eastern English Channel using sonar and remote
sampling techniques}. Estuarine Coastal and Shelf Science, 54, 263-278. \pageref{Brown02}
\bibitem{Boudouresque89}
Boudouresque C.F.\index[people]{Boudouresque C.F.}, Meinesz, A.\index[people]{Meinesz A.}, Fresi, E.\index[people]{Fresi, E.} and Gravez, V.\index[people]{Gravez, V}, 1989. \emph{2nd International Workshop on \textit{Posidonia oceanica} Beds}. G.I.S. Posidonie, Oceanica, 2, 322 pp. \pageref{Boudouresque89}
\bibitem{Bharathi03}
Bharathi A.M.\index[people]{Bharathi A.M.}, Kumar S.C.\index[people]{Kumar S.C.}, Kalliappan S.\index[people]{Kalliappan S.} 2003. \emph{Target Identification in multispectral images using \textit{in-situ} hyperspectral reflectance}. The Geospatial Resource Portal. Remote Sensing.
\href{www.gisdevelopment.net}{gisdevelopment} \pageref{Bharathi03}
\bibitem{Buia91}
Buia M.C.\index[people]{Buia M.C.} and Mazzella L.\index[people]{Mazzella L.} 1991. \emph{Reproductive phenology of the Mediterranean
seagrasses \textit{Posidonia oceanica} (L.) Delile, Cymodocea nodosa ( Ucria ) Aschers., and
Zostera noltii Hornem}. Aquatic Botany, 40 (1991) 343-362. Elsevier Science Publishers
B.V., Amsterdam. \pageref{Buia91}
\bibitem{Butler87}
Butler\index[people]{Butler} 1987. \emph{Marine resource mapping: an introductory manual}. \pageref{Butler87}
\bibitem{Cappo95}
Cappo M.\index[people]{Cappo, M.}, Alongi D.M.\index[people]{Alongi D.M.}, Williams D.\index[people]{Williams D.} and Duke N.\index[people]{Duke N.} 1995. \emph{A Review and Synthesis of Australian Fisheries Habitat Research: Major threats, issues and gaps in knowledge of
coastal and marine fisheries habitats}. Fisheries Research and Development Corporation. \pageref{Cappo95}
\bibitem{Carter08}
Carter G.A.\index[people]{Carter G.A.}, Criss G.A.\index[people]{Criss G.A.}, Blossom G.A.\index[people]{Blossom G.A.} and Biber P.D.\index[people]{Biber P.D.} 2008. \emph{Decadal-Scale Changes in Coverage on the Mississippi Barrier Islands, Northern Gulf of Mexico}.
\href{http://precedings.nature.com/documents/3669/version/1/files/npre20093669-1.pdf}{precedings.nature.com} \pageref{Carter08}
\bibitem{Calvo03}
Calvo S.\index[people]{Calvo S.}, Ciraolo G.\index[people]{Ciraolo G.} and Loggia G. la\index[people]{Loggia G. la}. 2003. \emph{Monitoring \textit{Posidonia oceanica} meadows in a Mediterranean coastal lagoon (Stagnone, Italy) by means of neural network and
ISODATA classification methods}. INT. J. REMOTE SENSING, 2003, VOL. 24, NO. 13,
2703–2716 \pageref{Calvo03}
\bibitem{Cambridge09}
Cambridge M.L.\index[people]{Cambridge M.L.}, Kendrick G.A.\index[people]{Kendrick G.A.} 2009. \emph{Contrasting responses of seagrass transplants
(Posidonia australis) to nitrogen, phosphorus and iron addition in an estuary and a coastal
embayment}. Journal of Experimental Marine Biology and Ecology. 371, 34–41. \pageref{Cambridge09}
\bibitem{Choo06}
Choo, C.K.\index[people]{Choo, C.K.} 2006. \emph{Save our Seahorses}. First Edition. \href{www.sosmalaysia.org}{sosmalaysia} \pageref{Choo06}
\bibitem{Clark99}
Clark, R.N.\index[people]{Clark, R.N.}, 1999. \emph{Spectroscopy of rocks and minerals and principles of spectroscopy}.
In: A.N. Rencz (ed.). Manual of Remote Sensing, Chapter 1, John Wiley and Sons, New York. 3-58. \pageref{Choo06}
\bibitem{Connell78}
Connell J.H.\index[people]{Connell J.H.} 1978. \emph{Diversity in tropical rain forests and coral reefs}. Science 199:1302–
10. \pageref{Connell78}
\bibitem{Dehouck08}
Dehouck A.\index[people]{Dehouck A.}, Martiny N.\index[people]{Martiny N.}, Froidefond J.-M.\index[people]{Froidefond J.-M.}, S\'en\'echal N.\index[people]{S\'en\'echal N.} 2008. \emph{In-Water Reflectance Spectra Measured On Board a Jet-Ski Across a Complex Nearshore Zone of Case-2 Waters during the ECORS Experiment}. Manuscript published in "Ocean Optics XIX, Barga: Italie
(2008)" \pageref{Dehouck08}
\bibitem{Dekker05a}
Dekker A.G.\index[people]{Dekker A.G.}, Brando V.E.\index[people]{Brando V.E.}, Anstee J.M.\index[people]{Anstee J.M.} 2005. \emph{Retrospective seagrass change detection in a shallow coastal tidal Australian lake}. Remote Sensing of Environment 97:415-433. \pageref{Dekker05a}
\bibitem{Dekker05b}
Dekker A.\index[people]{Dekker A.G.}, Brando V.E.\index[people]{Brando V.E.}, Anstee J.M.\index[people]{Anstee J.M.} 2005. \emph{Remote Sensing of Seagrass Ecosystems: Use of Spaceborne and Airborne Sensors}. SVNL083-Larkum. \pageref{05b}
\bibitem{DiCarlo04}
Di Carlo G.\index[people]{Di Carlo G.} 2004. \emph{The natural recolonisation process of the seagrass Posidonia
oceanica (L.) Delile after the introduction of the Italo-Algerian methane pipeline in the SW
Mediterranean sea}.\pageref{DiCarlo04}
\bibitem{Diaz-Almela06}
Diaz-Almela E.\index[people]{Diaz-Almela E.}, Marb\`a N.\index[people]{Marb\`a N.}, Alvarez E.\index[people]{Alvarez E.}, Santiago R.\index[people]{Santiago R.}, Holmer M.\index[people]{Holmer M.}, Grau A.\index[people]{Grau A.}, Danovaro R.\index[people]{Danovaro R.}, Argyrou M.\index[people]{Argyrou M.}, Karakassis I.\index[people]{Karakassis I.}, Duarte C.M.\index[people]{Duarte C.M.} 2006. \emph{Benthic inputs as predictors of seagrass (\textit{Posidonia oceanica}) fish farm-induced decline}. \href{http://arxiv.org/ftp/qbio/
papers/0611/0611006.pdf}{arxiv.org} \pageref{Diaz-Almela06}
\bibitem{Dierssen03}
Dierssen H.\index[people]{} and Zimmerman R.C.\index[people]{} 2003. \emph{Ocean color remote sensing of seagrass and
bathymetry in the Bahamas Banks by high-resolution airborne imagery}. Limnol. Oceanogr.,
48(1, part 2), 2003, 444–455, by the American Society of Limnology and Oceanography. \pageref{Dierssen03}
\bibitem{Dixon05}
Dixon R.\index[people]{}, editor. 2005. \emph{Watching the Seagrass Grow – a guide for community seagrass
monitoring in NSW (2nd Ed)}. Reference manual . The Community Environment Network,
Ourimbah. \pageref{Dixon05}
\bibitem{Duarte02}
Duarte, C.M.\index[people]{Duarte C.M.} 2002. \emph{The future of seagrass meadows}. Environmental Conservation 29 (2), 192–206. \pageref{Duarte02}
\bibitem{Dumay02}
Dumay O.\index[people]{Dumay O.}, Fernandez C.\index[people]{Fernandez C.} and Pergent G.\index[people]{Pergent G.} 2002. \emph{Primary production and vegetative cycle in \textit{Posidonia oceanica} when in competition with the green algae Caulerpa taxifolia nd
Caulerpa racemosa}. J.Mar. Biol. Ass. UK, 82, 379-387. \pageref{Dumay02}
\bibitem{Dural10}
Dural B.\index[people]{Dural B.} 2010. \emph{Phenological observations on \textit{Posidonia oceanica} (L.) Delile medows
along the coast of Akkum (Sıgacık Bay, Aegean Sea, Turkey)}. J. Black Sea/Mediterranean
Environment, Vol. 16 (1): 133-144. \pageref{Dural10}
\bibitem{Edwards00}
Edwards A.J.\index[people]{Edwards A.J.} (ed.) 2000. \emph{Remote Sensing handbook for tropical coastal management}.
\href{http://www.unesco.org/csi/pub/source/rs.htm}{UNESCO} \pageref{Edwards00}
\bibitem{Ferguson97}
Ferguson R.L.\index[people]{Ferguson R.L.} and Karl Korfmacher K.\index[people]{Korfmacher Karl K.} 1997 \emph{Remote sensing and GIS analysis of seagrass meadows in North Carolina, USA}. Aquatic Botany 58 (1997) 241-258. \pageref{Ferguson97}
\bibitem{Fletcher09}
Fletcher R.S.\index[people]{Fletcher R.S.}, Pulich Jr. W.\index[people]{Pulich Jr. W.} and Hardegree B.\index[people]{Hardegree B.} 2009. \emph{A Semiautomated Approach for Monitoring Landscape Changes in Texas Seagrass Beds from Aerial Photography}. Coastal
Education and Research Foundation, doi: 10.2112/07-0882.1 \pageref{Fletcher09}
\bibitem{Fornes06}
Fornes A.\index[people]{Fornes A.}, Basterretxea G.\index[people]{Basterretxea G.}, Orfila A.\index[people]{Orfila A.}, Jordi A.\index[people]{Jordi A.}, Alvarez A.\index[people]{Alvarez A.} and Tintore J.\index[people]{Tintore J.} 2006. \emph{Mapping
\textit{Posidonia oceanica} from IKONOS}. ISPRS Journal of Photogrammetry and Remote Sensing
60 (2006) 315–322. \pageref{Fornes06}
\bibitem{Francour06}
Francour, P.\index[people]{Francour P.}, Magreau, J.F.\index[people]{Magreau, J.F.}, Mannoni, P.A.\index[people]{Mannoni, P.A.}, Cottalorda, J.M.\index[people]{Cottalorda, J.M}, Gratiot, J.\index[people]{Gratiot, J.}, 2006.\emph{Management guide for Marine Protected Areas of the Mediterranean Sea, Permanent Ecological Moorings}. Universite de Nice-Sophia Antipolis and Parc National de Port-Cros,
Nice, 68 pp. \pageref{Francour06}
\bibitem{Francour99}
Francour P.\index[people]{Francour P.}, Ganteaume A.\index[people]{Ganteaume A.} and M. Poulain\index[people]{Poulain M.}. 1999. \emph{Effects of boats anchoring in \textit{Posidonia oceanica} seagrass beds in the Port-Cross National Park (north-western
Mediterranean Sea)}. Aquatic conservation: Marine and Freshwater Ecosystems: 9: 391-400. \pageref{Francour99}
\bibitem{Fyfe03}
Fyfe S.K.\index[people]{Fyfe S.K.} 2003. \emph{Spatial and temporal variation in spectral reflectance: Are seagrass
species spectrally distinct?} Limnology and Oceanography 48: 464–479. \pageref{Fyfe03}
\bibitem{Fyfe04}
Fyfe S.K.\index[people]{Fyfe S.K.} 2004. \emph{Hyperspectral studies of New South Wales with particular emphasis on
the detection of light stress in Eelgrass Zostera capricorni. PhD Thesis}. The University of
Wollongong. Research Online, \href{http://ro.uow.edu.au/theses/472/}{University of Wollongong Thesis Collection} \pageref{Fyfe03}
\bibitem{Gege04}
Gege P.\index[people]{Gege P.} 2004. \emph{The water color simulator WASI: an integrating software tool for analysis
and simulation of optical in situ spectra}. Computers and Geosciences 30 (2004) 523–532. \pageref{Gege04}
\bibitem{Gege05}
Gege, P.\index[people]{Gege P.} 2005. \emph{The Water Colour Simulator WASI. User manual for version 3}. DLR
Internal Report IB 564-01/05, 83 pp. \pageref{Gege05}
\bibitem{Gobert06}
Gobert S.\index[people]{Gobert S.}, Cambridge, M. L.\index[people]{Cambridge M.L.}, Velimirov, B.\index[people]{Velimirov B.}, Pergent, G.\index[people]{Pergent G.}, Lepoint, G.\index[people]{Lepoint G.}, Bouquegneau J.-M.\index[people]{Bouquegneau J.-M.}, Dauby, P.\index[people]{Dauby, P.}, Pergent-Martini, C.\index[people]{Pergent-Martini C.}, Walker, D. I.\index[people]{Walker, D.I.}. 2006. Biology of Posidonia. In Larkum A.W.D.\index[people]{Larkum A.W.D}, Orth, R.J.\index[people]{Orth R.J.}, Duarte C.M.\index[people]{Duarte C.M.} (eds.), \emph{Seagrasses: Biology, ecology and conservation}. Springer Verlag, Berlin, Pages 387–408. \pageref{Gobert06}
\bibitem{Gobert01}
Gobert S.\index[people]{Gobert S.}, Defawe O.\index[people]{Defawe O.}, Lepoint G.\index[people]{Lepoint G.}, Demoulin V.\index[people]{Demoulin V.} and Bouquegneau J.-M.\index[people]{Bouquegneau J.-M.} 2001. \emph{Anthesis effects on \textit{Posidonia oceanica} (L.) Delile phenology in the Bay of Calvi (Corsica, Mediterranean Sea)}. Hydrobiologia 455: 121–125. \pageref{Gobert01}
\bibitem{Green96}
Green, E.P.\index[people]{Green E.P.}, Mumby P.J.\index[people]{Mumby P.J.}, Edwards  A.J.\index[people]{Edwards A.J.}, and Clark C.D.\index[people]{Clark C.D.}, 1996. \emph{A review of remote sensing for the assessment and management of tropical coastal resources}. Coast. Manag.,
24(1):1-40. \pageref{Green96}
\bibitem{Green03}
Green E.P.\index[people]{Green E.P.}, Short F.T.\index[people]{Short F.T.} 2003. \emph{World Atlas of Seagrasses}. University of California Press,
Berkeley, USA. \pageref{Green03}
\bibitem{Guidetti08}
Guidetti P.\index[people]{Guidetti P.} 2008. \emph{Population dynamics of the seagrass \textit{Posidonia oceanica} for
evaluating meadow health status: a study case using reconstruction techniques}. Laboratory
of Zoology and Marine Biology, DiSTeBA, CoNISMa, University of Lecce. \pageref{Guidetti08}
\bibitem{Gullstroom06}
Gullstroom M.\index[people]{Gullstroom M.}, Lunden B.\index[people]{Lunden B.}, Bodin M.\index[people]{Bodin M.}, Kangwe J.\index[people]{Kangwe J.}, Ohman M.C.\index[people]{Ohman M.C.}, Mtolera M.S.P.\index[people]{Mtolera M.S.P.} and Bjork M.\index[people]{Bjork M.} 2006. \emph{Assessment of changes in the seagrass-dominated submerged vegetation of tropical Chwaka Bay (Zanzibar) using satellite remote sensing}. Estuarine, Coastal and Shelf
Science 67 (2006) 399-408. \pageref{Gullstroom06}
\bibitem{Haag08}
Haag S.M.\index[people]{Haag S.M.}, Kennish M.\index[people]{Kennish M.}, Sakowicz G.P.\index[people]{Sakowicz G.P.} 2008. \emph{Seagrass Habitat Characterization in Estuarine Waters of the Jacques Cousteau National Estuarine Research Reserve Using
Underwater Videographic Imaging Techniques}. Journal of Coastal Research, SI(55), 171–
179. West Palm Beach (Florida), ISSN 0749-0208. \pageref{Haag08}
\bibitem{DenHartog70}
Den Hartog, C.\index[people]{Den Hartog C.} 1970. \emph{The Seagrasses of the World}. North Holland Publishing Company,
Amsterdam, 275 pp. 31 plates. \pageref{DenHartog70}
\bibitem{DenHartog77}
Den Hartog, C.\index[people]{Den Hartog C.}, 1977. \emph{Structure, function, and classification in seagrass communities}.
In: C.P. McRoy\index[people]{McRoy, C.P.} and C. Helfferich\index[people]{Helfferich C.} (Editors), Seagrass Ecosystems: a Scientific Perspective.
Marcel, New York, pp. 89-121. \pageref{DenHartog77}
\bibitem{Heege03}
Heege T.\index[people]{Heege T.}, Bogner A.\index[people]{Bogner A.}, Pinnel N.\index[people]{Pinnel N.} 2003. \emph{Mapping of submerged aquatic vegetation with a physically based process chain}. Proc. Remote Sensing, September 8-12 2--3/ Barcelona,
Spain. Ed. by Kramer E.\index[people]{Kramer E.} The International Society for Optical Engineering, Vol.5233. \pageref{Heege03}
\bibitem{Hirose04}
Hirose K.\index[people]{Hirose K.}, Syoji M.\index[people]{Syoji M.}, Hang H.T.M.\index[people]{Hang H.T.M.}, Anh N.H.\index[people]{Anh N.H.}, Triet T.\index[people]{Triet T} and Nam V.N.\index[people]{Nam V.N.} 2004. \emph{Satellite data application for mangrove management}. International Symposium on Geoinfomatics for Spatial Infrastructure Development in Earth and Allied Sciences. \pageref{Hirose04} 
\bibitem{Hochberg03a}
Hochberg, E.J.\index[people]{Hochberg E.J.}, Atkinson, M.J.\index[people]{Atkinson, M.J.} 2003a. \emph{Capabilities of remote sensors to classify coral,
algae, and sand as pure and mixed spectra}. Remote Sensing of Environment 85, 174–189. \pageref{Hochberg03a}
\bibitem{Hochberg03b}
Hochberg, E.J.\index[people]{Hochberg E.J.}, Atkinson, M.J.\index[people]{Atkinson, M.J.}, Andr\'efou\"et, S.\index[people]{Andr\'efou\"et S.}, 2003b. \emph{Spectral reflectance of coral reef bottom-types worldwide and implications for coral reef remote sensing}. Remote Sensing of
Environment 85, 159–173. \pageref{Hochberg03b}
\bibitem{Hogarth07}
Hogarth P.\index[people]{Hogarth P.} 2007. \emph{The Biology of Mangroves and Seagrasses}. Oxford University Press. \pageref{Hogarth07}
\bibitem{Holmer09}
Holmer M.\index[people]{Holmer M.}, Marb\`a N.\index[people]{Marb\`a N.}, Lamote M.\index[people]{Lamote M.} and Duarte C.\index[people]{Duarte C.M.} 2009\emph{Deterioration of Sediment Quality in Seagrass Meadows (\textit{Posidonia oceanica}) Invaded by Macroalgae (Caulerpa sp.)}. Estuaries
and Coasts (2009) 32:456–466 DOI 10.1007/s12237-009-9133-4 \pageref{Hogarth07}
\bibitem{Holmer08}
Holmer M.\index[people]{Holmer M.}, Argyrou M.\index[people]{Argyrou M.}, Dalsgaard T.\index[people]{Dalsgaard T.}, Danovaro R.\index[people]{Danovaro R.}, Diaz-Almela E.\index[people]{Diaz-Almela E.}, Duarte C.\index[people]{Duarte C.M.}, Frederiksen M.\index[people]{Frederiksen M.}, Grau A.\index[people]{Grau A.}, Karakassis I.\index[people]{Karakassis I.}, Marb\`a N.\index[people]{Marb\`a N.}, Mirto S.\index[people]{Mirto S.}, P\'erez M.\index[people]{P\'erez M}, Pusceddu A.\index[people]{Pusceddu A.}, Tsapakis M.\index[people]{Tsapakis M.}, 2008
\emph{Effects of fish farm waste on \textit{Posidonia oceanica} meadows: Synthesis and provision of
monitoring and management tools}. Marine Pollution Bulletin 56 (2008) 1618–1629. \pageref{Holmer08}
\bibitem{Iverson86}
Iverson R.L.\index[people]{Iverson R.L.} and Bittaker H.F.\index[people]{Bittaker H.F.} 1986. \emph{Seagrass Distribution and Abundance in Eastern
Gulf of Mexico Coastal Waters}. Estuarine, Coastal and Shelf Science, 22, 577-602. \pageref{Iverson86}
\bibitem{Jackson07}
Jackson J.\index[people]{Jackson J.}, Nemeth D.J.\index[people]{Nemeth D.J.} 2007. \emph{A New Method to Describe Seagrass Habitat Sampled
During Fisheries-Independent Monitoring}. Estuaries and Coasts Vol. 30, No. 1, p. 171–178 February 2007 \pageref{Jackson07}
\bibitem{Jensen01}
Jensen S.\index[people]{Jensen S.L.} and Bell S.S.\index[people]{Bell S.S.} 2001. \emph{Seagrass growth and patch dynamics: cross-scale
morphological plasticity}. Plant Ecology 155: 201–217, 2001. \pageref{Jensen01}
\bibitem{Jensen95}
Jensen J.R.\index[people]{Jensen J.R.}, Rutchey K.\index[people]{Rutchey K.}, Koch M.S.\index[people]{Koch M.S.}, Narumalani S.\index[people]{Narumalani S.} 1995. \emph{Inland wetland change detection in the everglades water conservation area 2a using a time-series of normalized
remotely-sensed data}. Photogrammetric Eng Remote Sensing 61: 199–209 \pageref{Jensen95}
\bibitem{Jerlov52}
Jerlov N.G.\index[people]{Jerlov N.G.} 1952. \emph{Influence of suspended and dissolved matter on the transparency of
sea water}. Wiley online Library. TellusVolume 5, Issue 1, Published online: 2010. \pageref{Jerlov52}
\bibitem{Kalinowski04}
Kalinowski A.\index[people]{Kalinowski A.} and Oliver S.\index[people]{Oliver S.} \emph{2004 ASTER Mineral Index Processing Manual}. Remote
Sensing Applications. Geoscience Australia \pageref{Kalinowski04}
\bibitem{Kelly80}
Kelly, M.G.\index[people]{Kelly, M.G.}, 1980. \emph{Remote sensing of seagrass beds, Handbook of Seagrass Biology: An
Ecosystem Perspective} (R.C. Phillips\index[people]{Phillips, R.C.} and C.P. McRoy\index[people]{McRoy, C.P.}, editors), Garland STPM Press, New
York, N.Y., pp. 69-86. \pageref{Kelly80}
\bibitem{Kendrick00}
Kendrick G.A.\index[people]{Kendrick G.A.}, Hegge B.J.\index[people]{Hegge B.J.}, Wyllie A.\index[people]{Wyllie A.}, Davidson A.\index[people]{Davidson A.}, Lord D.A.\index[people]{Lord D.A.}. 2000. \emph{Changes in Seagrass Cover on Success and Parmelia Banks, Western Australia Between 1965 and 1995}. Estuarine, Coastal and Shelf Science, 50:341-353. \pageref{Kendrick00}
\bibitem{Kirk94}
Kirk J.T.O.\index[people]{Kirk J.T.O.} 1994. \emph{Light and photosynthesis in aquatic ecosystems}, pp 1–509. University
Press, Cambridge, UK. \pageref{Kirk94}
\bibitem{Kirkman96}
Kirkman H.\index[people]{Kirkman H.} 1996. \emph{Baseline and Monitoring Methods for Seagrass Meadows}. Journal of
Environmental Management (1996) 47, 191–201. \pageref{Kirkman96}
\bibitem{Kuo89}
Kuo J.\index[people]{Kuo J.}, McComb A.J.\index[people]{McComb A.J.}, 1989. \emph{Seagrass taxonomy, structure and development. In: Larkum
A.W.D.\index[people]{Larkum A.W.D.}, McComb A.J.\index[people]{McComb A.J.}, Shepherd J.\index[people]{Shepherd J.}, eds. Biology of seagrasses: a treatise on the biology of seagrasses with special reference to the Australian region}. Elsevier, The Netherlands: 6-73. \pageref{Kuo89}
\bibitem{Leoni06}
Leoni V.\index[people]{Leoni V.}, Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Vela A.\index[people]{Vela A.}, Pergent G.\index[people]{Pergent G.} 2006. \emph{Morphological responses of \textit{Posidonia oceanica} to experimental nutrient enrichment of the canopy water}. Journal of Experimental Marine Biology and Ecology xx (2006) xxx–xxx. \pageref{Leoni06}
\bibitem{Leoni07}
Leoni V.\index[people]{Leoni V.}, Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Vela A.\index[people]{Vela A.}, Pergent G.\index[people]{Pergent G.} 2007. \emph{Physiological responses of \textit{{Posidonia oceanica}} to experimental nutrient enrichment of the canopy water}. Journal of Experimental Marine Biology and Ecology 349, 73–83. \pageref{Leoni07}
\bibitem{Leoni08}
Leoni V.\index[people]{Leoni V.}, Vela A.\index[people]{Vela A.}, Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pergent G.\index[people]{Pergent G.} 2008. \emph{Effects of experimental reduction of light and nutrient enrichments (N and P) on seagrasses: a review}.
Aquatic Conservation: Marine and Freshwater Ecosystems, 18: 202–220. \pageref{Leoni08}
\bibitem{Lewis01}
Lewis M.\index[people]{Lewis M.} 2001. \emph{Discriminating Vegetation with Hyperspectral Imagery - What is
Possible?} \href{http://www.ieeexplore.ieee.org/stamp/stamp.jsp?arnumber=00978199}{ieeexplore.ieee.org} Proceedings,
International Geoscience and Remote Sensing Symposium, 9-13 July 2001, Sydney, Australia. ISBN 0-7803-7031-7 \pageref{Lewis01}
\bibitem{Louchard03}
Louchard E.M.\index[people]{Louchard E.M.}, Reid R.P.\index[people]{Reid R.P.}, Stephens F.C.\index[people]{Stephens F.C.}, Davis C.O.\index[people]{Davis C.O.}, Leathers R.A.\index[people]{Leathers R.A.}, Downes T.V.\index[people]{Downes T.V.} 2003. \emph{Optical remote sensing of benthic habitats and bathymetry in coastal environments at Lee Stocking Island, Bahamas: A comparative spectral classification approach}. Limnology
and Oceanography (2003) 48:511-521. \pageref{Louchard03}
\bibitem{Luque04}
Luque A.\index[people]{Luque A.} and Templado J.A.\index[people]{Templado J.A.} 2004. \emph{Praderas y bosques marinos de Andalucia
Consejeria de Medio Ambiente, Junta de Andalucia} (in press). \pageref{Luque04}
\bibitem{Lyzenga81}
Lyzenga D.R.\index[people]{Lyzenga D.R.} 1981. \emph{Remote sensing of bottom reflectance and water attenuation in
shallow water using aircraft and Landsat data}. International Journal of Remote Sensing 2,
71-82. \pageref{Lyzenga81}
\bibitem{Maffione97}
Maffione R.A.\index[people]{Maffione R.A.} and Dana D.R.\index[people]{Dana D.R.} 1997 \emph{Instruments and methods for measuring the
backward-scattering coefficient of ocean waters}. Optical Society of America \pageref{Maffione97}
\bibitem{Maffione01a}
Maffione R.A.\index[people]{Maffione R.A.} 2001a. \emph{Evolution and Revolution in Measuring Ocean Optical Properties}.
Oceanography, Vol.14, No.3. \pageref{Maffione01a}
\bibitem{Maffione01b}
Maffione R.A.\index[people]{Maffione R.A.} 2001b.\emph{Hyperspectral Optical Properties, Remote Sensing, and
Underwater Visibility}. Hydro-Optics, Biology, and Instrumentation Laboratories. \pageref{Maffione01b}
\bibitem{Malthus03}
Malthus T.J.\index[people]{Malthus T.J.} and Karpouzli E.\index[people]{Karpouzli E.} 2003. \emph{Integrating field and high spatial resolution
satellite based methods for monitoring shallow submersed aquatic habitats in the Sound of Eriskay, Scotland, UK}. Int Journal Remote Sensing 24: 2585–2593. \pageref{Malthus03}
\bibitem{Matarrese08}
Matarrese R.\index[people]{Matarrese R.}, Acquaro M.\index[people]{Acquaro M.}, Morea A.\index[people]{Morea A.}, Tijani K.\index[people]{Tijani K.}, Chiaradia M.T.\index[people]{Chiaradia M.T.} 2008. \emph{Applications of remote sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}. Geoscience and Remote Sensing Symposium. ISBN:978-1-4244-2807-6. \pageref{Matarrese08}
\bibitem{Matarrese06}
Matarrese R.\index[people]{Matarrese R.}, Acquaro M.\index[people]{Acquaro M.}, Morea A.\index[people]{Morea A.}, Khalid T.\index[people]{Khalid T.}, Chiaradia M.T.\index[people]{Chiaradia M.T.} 2006 \emph{Applications of remote sensing techniques for mapping \textit{{Posidonia oceanica}} meadows}.
\href{http://www.igarss08.org/Abstracts/pdfs/3888.pdf}{International Geoscience and Remote Sensing Symposium} \pageref{Matarrese06}
\bibitem{McKenzie02}
McKenzie L.J.\index[people]{McKenzie L.J.}, Campbell S.J.\index[people]{Campbell S.J.} 2002. \emph{Manual for Community (citizen) Monitoring of
Seagrass Habitat}. Western Pacific Edition. Marine Plant Ecology Group, QDPI, Northern Fisheries Centre, Cairns. 43 pp. \pageref{McKenzie02}
\bibitem{McKenzie03}
McKenzie L.J.\index[people]{McKenzie L.J.}, Campbell S.J.\index[people]{Campbell S.J.} and Roder C.A.\index[people]{Roder C.A.} 2003. \emph{Seagrass-Watch: Manual for Mapping and Monitoring Seagrass Resources by Community (citizen) Volunteers}. 2nd Edition. QFS, NFC, Cairns. 100pp. \pageref{McKenzie03}
\bibitem{McKenzieetal03}
McKenzie L.J.\index[people]{McKenzie L.J.} 2003. \emph{Guidelines for the rapid assessment of seagrass habitats in the
Western Pacific}. Marine Plant Ecology Group, QDPI, Northern Fisheries Centre, Cairns, \href{http://www.seagrasswatch.org}{seagrasswatch} \pageref{McKenzieetal03}
\bibitem{McKenzie06}
McKenzie L.J.\index[people]{McKenzie L.J.}, Campbell S.J.\index[people]{Campbell S.J.} and F.Lasi\index[people]{Lasi, F.} 2006. \emph{Seagrasses and Mangroves}. In: Green A.P.\index[people]{Green A.P.}, Lokani W.\index[people]{Lokani W.}, Atu P.\index[people]{Atu P.}, Ramonia P.\index[people]{Ramonia P.}, Thomas P.\index[people]{Thomas P.} and Almanmy J.\index[people]{Almanmy J.} (eds). 2006 \emph{Solomon Islands Marine Assessment. Technical report of survey conducted May 13 to June 17 2004}. The Nature Conservancy, Indo-Pacific Resource Centre. Report No 1/06 \pageref{McKenzie06}
\bibitem{McKenzie07}
McKenzie L.J.\index[people]{McKenzie L.J.} and Yoshida R.\index[people]{Yoshida R.} 2007. \emph{Seagrass-Watch: Guidelines for Monitoring
Seagrass Habitats in the Fiji Islands. Proceeding of a training workshop. Corpus Christi Teachers College Laucala Bay, Suva, Fiji, 16th June 2007}. Seagrass-Watch HQ Cairns, 42pp. \pageref{McKenzie07}
\bibitem{McKenzie08}
McKenzie L.J.\index[people]{McKenzie L.J.} 2008. \emph{Seagrass-Watch. Proceedings of a Workshop for Mapping and
Monitoring Seagrass Habitats in North East Arnhem Land, Northern Territory. 18-20 October 2008}. Seagrass-Watch HQ, Cairns, 49pp. \pageref{McKenzie08}
\bibitem{McKenzie09}
McKenzie L.J.\index[people]{McKenzie L.J.}, Yoshida R.\index[people]{Yoshida R.} 2009. Seagrass Watch. Proceedings of a Workshop for
Monitoring Seagrass Habitats in Indonesia. The Nature Conservancy, Coral Triangle Centre, Sanur, Bali, Indonesia. \pageref{McKenzie09}
\bibitem{Meinesz91}
Meinesz, A.\index[people]{Meinesz A.}, Lefevre, J.R.\index[people]{Lefevre, J.R.}, Astier, J.M.\index[people]{Astier, J.M.}, 1991. \emph{Impact of coastal development on the infralittoral zone along the south-eastern Mediterranean shore of continental France}. Marine Pollution Bullettin 23, 343–347. \pageref{Meinesz91}
\bibitem{Mellors09a}
Mellors J.\index[people]{Mellors J.} and McKenzie L.\index[people]{McKenzie L.J.} 2009(a). \emph{Seagrass-Watch: Guidelines for Monitoring
Seagrass Habitats. Proceedings of a workshop held on Thursday Island, Torres Strait 4-8.03.2009}. \pageref{Mellors09a}
\bibitem{Mellors09b}
Mellors J.\index[people]{Mellors J.} and McKenzie L.\index[people]{McKenzie L.J.} 2009(b). \emph{Seagrass-Watch: Seagrass Monitoring Guidelines
for Torres Strait Communities. Meer Island. Proceedings of a workshop held on Meer Island, Torres Strait 14-15 September 2009}. 58 pp. \pageref{Mellors09b}
\bibitem{Micheli05}
Micheli C.\index[people]{Micheli C.}, Paganin P.\index[people]{Paganin P.}, Peirano A.\index[people]{Peirano A.}, Caye, G.\index[people]{Caye, G.}, Meinesz A.\index[people]{Meinesz A.} and Bianchi C.N.\index[people]{Bianchi C.N.} 2005. \emph{Genetic variability of \textit{{Posidonia oceanica}} (L.) Delile in relation to local factors and biogeographic patterns}. Aquatic Botany 82 (2005) 210–221 \pageref{Micheli05}
\bibitem{Montefalcone09}
Montefalcone M.\index[people]{Montefalcone M.} 2009.\emph{Ecosystem health assessment using the Mediterranean seagrass
\textit{Posidonia oceanica}: A review}. Ecological indicators 595 – 604 \pageref{Montefalcone09}
\bibitem{Montefalcone10}
Montefalcone M.\index[people]{Montefalcone M.}, Albertelli G.\index[people]{Albertelli G.}, Morri C.\index[people]{Morri C.} and Bianchi C.N.\index[people]{Bianchi C.N.} 2010. \emph{Patterns of wide-scale substitution within meadows of the seagrass \textit{{Posidonia oceanica}} in NW Mediterranean Sea:
invaders are stronger than natives}. Aquatic Conserv: Mar. Freshw. Ecosyst. 20: 507–515. \pageref{Montefalcone10}
\bibitem{Montefalcone06}
Montefalcone M.\index[people]{Montefalcone M.}, Albertelli G.\index[people]{Albertelli G.}, Bianchi C.N.\index[people]{Bianchi C.N.}, Mariani M.\index[people]{Mariani M.}, Morri C.\index[people]{Morri C.} 2006. \emph{A new synthetic index and a protocol for monitoring the status of \textit{{Posidonia oceanica}} meadows: a case study at Sanremo (Ligurian Sea, NW Mediterranean)}. Aquatic Conservation: Marine and Freshwater Ecosystems, 16, 29–42. \pageref{Montefalcone06}
\bibitem{Mount03}
Mount R.\index[people]{Mount R.} 2003. \emph{The application of digital aerial photography to shallow water seabed
mapping and monitoring – how deep can you see? Coastal GIS 2003: and integrated approach to Australian coastal issues, 7th - 8th July 2003}. Wollongong, University of Wollongong, pp. 139-156. \pageref{Mount03}
\bibitem{Mumby99}
Mumby P.J.\index[people]{Mumby P.J.}, Green E.P.\index[people]{Green E.P.}, Edwards A.J.\index[people]{Edwards A.J.} and Clark C.D.\index[people]{Clark C.D.} 1999. \emph{The cost-effectiveness of remote sensing for tropical coastal resources assessment and management}. Journal of Environmental Management Volume 55, Issue 3, March 1999, Pages 157-166 \pageref{Mumby99}
\bibitem{Mumby02}
Mumby, P.J.\index[people]{Mumby P.J.} and Edwards A.J.\index[people]{Edwards A.J.} 2002. \emph{Mapping marine environments with IKONOS
imagery: Enhanced spatial resolution can deliver greater thematic accuracy}. Remote Sensing of Environment, 82, 248–257 \pageref{Mumby02}
\bibitem{Mumby04}
Mumby P.J.\index[people]{Mumby P.J.}, Skirving W.\index[people]{Skirving W.}, Strong A.E.\index[people]{Strong A.E.}, Hardy J.T.\index[people]{Hardy J.T.}, LeDrew E.F.\index[people]{LeDrew E.F.}, Hochberg E.J.\index[people]{Hochberg E.J.}, Stumpf R.P.\index[people]{Stumpf R.P.} and David L.T.\index[people]{David L.T.} 2004. \emph{Remote sensing of coral reefs and their physical environment}. Marine Pollution Bulletin, 48:219-228. \pageref{Mumby04}
\bibitem{Nagelkerken00}
Nagelkerken I.\index[people]{Nagelkerken I.}, van der Velde G.\index[people]{van der Velde G.}, Gorissen M.W.\index[people]{Gorissen M.W.}, Meijer G.J.\index[people]{Meijer G.J}, van’t Hofc T.\index[people]{van’t Hofc T.} and Den Hartog C.\index[people]{Den Hartog C.} 2000. \emph{Importance of Mangroves, Seagrass Beds and the Shallow Coral Reef as a Nursery for Important Coral Reef Fishes, Using a Visual Census Technique}. Estuarine,
Coastal and Shelf Science 51, 31–44 doi:10.1006/ecss.2000.0617 \pageref{Nagelkerken00}
\bibitem{Noralez10}
Noralez S.\index[people]{Noralez S.} 2010. \emph{Exploring the Limits of Mapping Seagrass. A case study for P. oceanica}. MSc Thesis. ITC. Netherlands. \pageref{Noralez10}
\bibitem{Norris97}
Norris J.G.\index[people]{Norris J.G.}, Wyllie-Echeverria S.\index[people]{Wyllie-Echeverria S.}, Mumford T.\index[people]{Mumford T.}, Bailey A.\index[people]{Bailey A.} and Turner T.\index[people]{Turner T.} 1997. \emph{Estimating basal area coverage of subtidal seagrass beds using underwater videography}.
Aquatic Botany 58 (1997) 269-287 \pageref{Noralez10}
\bibitem{SPOT} 
\emph{SPOT, Official Website} \href{http://www.spotimage.com/web/en/172-spot-images.php}{SPOT} \pageref{SPOT}
\bibitem{Orth06}
Orth R.J.\index[people]{Orth R.J.}, Carruthers T.J.B.\index[people]{Carruthers T.J.B.}, Dennison W.C.\index[people]{Dennison W.C.}, Duarte C.M.\index[people]{Duarte C.M.}, Fourqurean J.W.\index[people]{Fourqurean J.W.}, Heck Jr.K. L.\index[people]{Heck Jr.K. L.}, Hughes A. R.\index[people]{Hughes A. R.}, Kendrick G. A.\index[people]{Kendrick G.A.}, Kenworthy W. J.\index[people]{Kenworthy W. J.}, Olyarnik S.\index[people]{Olyarnik S.}, Short F.T.\index[people]{Short F.T}, Waycott M.\index[people]{Waycott M.} and Williams S. L.\index[people]{Williams S.L.} 2006. \emph{A Global Crisis for Seagrass Ecosystems}. Vol. 56 No. 12, BioScience 987. \pageref{Orth06}
\bibitem{Palandro03}
Palandro D.\index[people]{Palandro D.}, Andr\'efou\"et S.\index[people]{Andr\'efou\"et S.}, Muller-Karger F. E.\index[people]{Muller-Karger F. E.}, Dustan Ph.\index[people]{Dustan Ph}, Hu Ch.\index[people]{Hu Ch.}, and Hallock P.\index[people]{Hallock P.} 2003. \emph{Detection of changes in coral reef communities using Landsat-5 TM and Landsat-7 ETM+ data}. Can. J. Remote Sensing, Vol. 29, No. 2, pp. 201–209. \pageref{Palandro03}
\bibitem{Pasqualini98a}
Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.} and Pergent, G.\index[people]{Pergent G.} 1998(a). \emph{Use of remote sensing for the characterization of the Mediterranean coastal environment – the case of Posidonia
oceanica}. Journal of Coastal Conservation 4: 59-66, 1998, EUCC; Opulus Press Uppsala. \pageref{Pasqualini98a}
\bibitem{Pasqualini98b}
Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Clabaut P.\index[people]{Clabaut P.}, Pergent G.\index[people]{Pergent G.} 1998(b). \emph{Mapping of \textit{{Posidonia oceanica}} using aerial photographs and side scan sonar: Application off the
island of Corsica (France)}. Estuarine Coastal Shelf Sci 47:359–367. \pageref{Pasqualini98b}
\bibitem{Pasqualini99}
Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pergent G.\index[people]{Pergent G.} 1999. \emph{Environmental impact identification along the Corsican coast (Mediterranean Sea) using image processing}. Aquatic Botany,
65:311-320. \pageref{Pasqualini99}
\bibitem{Pasqualini01}
Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Clabaut P.\index[people]{Clabaut P.}, Marteel H.\index[people]{Marteel H.}, and Pergent G.\index[people]{Pergent G.} 2001. \emph{Integration of Aerial Remote Sensing, Photogrammetry, and GIS Technologies in Seagrass
Mapping}. Photogrammetric Engineering and Remote Sensing Vol. 67, No. 1, January 2001, pp. 99-105. \pageref{Pasqualini01}
\bibitem{Pasqualini05}
Pasqualini V.\index[people]{Pasqualini V.}, Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pergent G.\index[people]{Pergent G.}, Agreila M.\index[people]{Agreila M.}, Skoufas G.\index[people]{Skoufas G.}, Sourbes L.\index[people]{Sourbes L.} and Tsirika A.\index[people]{Tsirika A.} 2005. \emph{Use of SPOT 5 for mapping seagrasses: An application to Posidonia oceanica}. Remote Sensing of Environment, 94:39-45. \pageref{Pasqualini05}
\bibitem{Peneva08}
Peneva E.\index[people]{Peneva E.I.}, Griffith J.A.\index[people]{Griffith J.A.} and Carter G.A.\index[people]{Carter G.A.} 2008. \emph{Seagrass Mapping in the Northern Gulf of Mexico using Airborne Hyperspectral Imagery: A Comparison of Classification Methods}. Journal of Coastal Research, Number 244:850-856. 2008. Published By: Coastal Education and Research Foundation DOI: 10.2112/06-0764.1 \pageref{Peneva08}
\bibitem{Peneva07}
Peneva E.I.\index[people]{Peneva E.I.}, Griffith J.A.\index[people]{Griffith J.A.} and Carter G.A.\index[people]{Carter G.A.} 2007. \emph{Mapping seagrass distribution on Horn Island}, MS using hyperspectral imagery.
\href{ftp://ftp.ecn.purdue.edu/jshan/proceedings/asprs2005/Files/0134.pdf}{ecn.purdue.edu} \pageref{Peneva07}
\bibitem{Pergent90}
Pergent G.\index[people]{Pergent G.} 1990. \emph{Lepidochronological analysis of the seagrass \textit{{Posidonia oceanica}} (L.)
Delile: a standardized approach}. Aquatic Botany, 37 (1990) 39-54 39 Elsevier Science Publishers B.V., Amsterdam -- Printed in The Netherlands \pageref{Pergent90}
\bibitem{Pergent-Martini06}
Pergent-Martini C.\index[people]{Pergent-Martini C.}, Pasqualini V.\index[people]{Pasqualini V.}, Ferrat L.\index[people]{Ferrat L.} and Pergent G.\index[people]{Pergent G.} 2006. \emph{Ecological Data in Integrated Coastal Zone Management: Case Study of \textit{{Posidonia oceanica}} Meadows Along
the Corsican Coastline (Mediterranean Sea)}. Environ Manage (2006) 38:889–895 DOI
10.1007/s00267-005-0056-y. \pageref{Pergent-Martini06}
\bibitem{Pergent-Martini05}
Pergent-Martini C.\index[people]{Pergent-Martini C.}, Leoni V.\index[people]{Leoni V.}, Pasqualini V.\index[people]{Pasqualini V.}, Ardizzone G.D.\index[people]{Ardizzone G.D.}, Balestri E.\index[people]{Balestri E.}, Bedini R.\index[people]{Bedini R.}, Belluscio A.\index[people]{Belluscio A.}, Belsher T.\index[people]{Belsher T.}, Borg J.\index[people]{Borg J.}, Boudouresque C.F.\index[people]{Boudouresque C.F.}, Boumaza S.\index[people]{Boumaza S.}, Bouquegneau J.-M.\index[people]{Bouquegneau J.-M.}, Buia M.C.\index[people]{Buia M.C.}, Calvo S.\index[people]{Calvo S.}, Cebrian J.\index[people]{Cebrian J.}, Charbonnel E.\index[people]{Charbonnel E.}, Cinelli F.\index[people]{Cinelli F.}, Cossu A.\index[people]{Cossu A.}, Di Maida G.\index[people]{Di Maida G.}, Dural B.\index[people]{Dural B.}, Francour P.\index[people]{Francour P.}, Gobert S.\index[people]{Gobert S.}, Lepoint G.\index[people]{Lepoint G.}, Meinesz A.\index[people]{Meinesz A.}, Molenaar H.\index[people]{Molenaar H.}, Mansour H.M.\index[people]{Mansour H.M.}, Panayotidis P.\index[people]{Panayotidis P.}, Peirano A.\index[people]{Peirano A.}, Pergent G.\index[people]{Pergent G.}, Piazzi L.\index[people]{Piazzi L.}, Pirrotta M.\index[people]{Pirrotta M.}, Relini G.\index[people]{Relini G}, Romero J.\index[people]{Romero J.}, Sanchez-Lizaso J.L.\index[people]{Sanchez-Lizaso J.L.}, Semroud R.\index[people]{Semroud R.}, Shembri P.\index[people]{Shembri P.}, Shili A.\index[people]{Shili A.}, Tomasello A.\index[people]{Tomasello A.}, Velimirov B.\index[people]{Velimirov B.} 2005. \emph{Descriptors of \textit{{Posidonia oceanica}} meadows: Use and application}. Ecological Indicators 5, 213–230. \pageref{Pergent-Martini05}
\bibitem{Phinn08}
Phinn S.\index[people]{Phinn S.}, Roelfsema C.\index[people]{Roelfsema C.}, Dekker A.\index[people]{Dekker A.G.}, Brando V.E.\index[people]{Brando V.E.}, Anstee J.M.\index[people]{Anstee J.M.} 2008. \emph{Mapping seagrass species, cover and biomass in shallow waters: An assessment of satellite multi-spectral and airborne
hyperspectral imaging systems in Moreton Bay (Australia)}. Remote Sensing of Environment, 112:3413-3425. \pageref{Phinn08}
\bibitem{Ralph05}
Ralph P.J.\index[people]{Ralph P.J.}, Macinnis-Ng C.M.O.\index[people]{Macinnis-Ng C.M.O.}, Frankart C.\index[people]{Frankart C.} 2005. \emph{Fluorescence imaging application: effect of leaf age on seagrass photokinetics}. Aquatic Botany 81 (2005) 69–84. \pageref{Ralph05}
\bibitem{Rasib97}
Rasib A.W.\index[people]{Rasib A.W.} and Hashim M.\index[people]{Hashim M.} 1997. \emph{Mapping Seagrass From Satellite Remote Sensing
Data. Proceedings of Asian Conference on Remote Sensing}. \href{http://eprints.utm.my/2161/}{eprints.utm.my} \pageref{Rasib97}
\bibitem{Ribed02}
Ribed P.S.\index[people]{Ribed P.S.} 2002. \emph{Submarine vegetation studies using remote sensing: importance in beach measurements}. Journal of Environmental Protection and Ecology, 3 No 4, 953-958 (2002) \pageref{Ribed02}
\bibitem{Richardson06}
Richardson L.L.\index[people]{Richardson L.L.}, LeDrew Ellesworth F.\index[people]{LeDrew E.F.} 2006. \emph{Remote Sensing of Aquatic Costal
Ecosystems Processes}. Springer. Science and Management Applications. \pageref{Richardson06}
\bibitem{Robbins94}
Robbins B.D.\index[people]{Robbins B.D.}, Bell S.S.\index[people]{Bell S.S.} 1994. \emph{Seagrass landscapes: a terrestrial approach to the marine
subtidal environment}. TREE vol.9, no 8, doi:10.1016/0169-5347(94)90041-8 \pageref{Robbins94}
\bibitem{Sabins97}
Sabins F.\index[people]{Sabins F.} 1997. \emph{Remote Sensing, Principles and Interpretation}. W.H. Freeman and
Company, New York. \pageref{Sabins97}
\bibitem{Schmidt03}
Schmidt K.\index[people]{Schmidt K.} and Skidmore A.K.\index[people]{Skidmore A.K.} 2003. \emph{Spectral discrimination of vegetation types in a
coastal wetland}. Remote Sensing of Environment. 85:92-108. \pageref{Schmidt03}
\bibitem{Schmieder97}
Schmieder K.\index[people]{Schmieder K.} 1997. \emph{Littoral zone - GIS of Lake Constance: a useful tool in lake
monitoring and autecological studies with submersed macrophytes}. Aquatic Botany 58
(1997) 333-346 \pageref{Schmieder97}
\bibitem{Schultz08}
Schultz S.T.\index[people]{Schultz S.T.} 2008. \emph{Seagrass monitoring by underwater videography: Disturbance
regimes, sampling design, and statistical power}. Aquatic Botany 88 (2008) 228–238. \pageref{Schultz08}
\bibitem{Short01}
Short F.T.\index[people]{Short F.T.} and Coles R.G.\index[people]{Coles R.G.} (eds). 2001. \emph{Global Seagrass Research Methods}, p 482.
Elsevier Publishing Co., Amsterdam \pageref{Short01}
\bibitem{Shortis07}
Shortis M.\index[people]{Shortis M.}, Harvey E.\index[people]{Harvey E.}, Seager J.\index[people]{Seager J.} 2007. \emph{A Review of the Status and Trends in Underwater Videometric Measurement}. SPIE Conference 6491, Videometrics IX,IS and T/SPIE Electronic Imaging, San Jose, California, U.S.A., 28 January–1 February 2007 \pageref{Shortis07}
\bibitem{Stephens03}
Stephens F.C.\index[people]{Stephens F.C.}, Louchard E.M.\index[people]{Louchard E.M.}, Reid R.P.\index[people]{Reid R.P.}, Maffione R.A.\index[people]{Maffione R.A.} 2003. \emph{Effects of microalgal communities on reflectance spectra of carbonate sediments in subtidal optically shallow
marine environments}. Limnology and Oceanography, 48:535-546. \pageref{Stephens03}
\bibitem{Thorhaug07}
Thorhaug, A.\index[people]{Thorhaug, A.}, Richardson, A. D.\index[people]{Richardson, A.D.} and Berlyn, G.P.\index[people]{Berlyn, G.P.} 2007. \emph{Spectral reflectance of the seagrasses: Thalassia testudinum, Halodule wrightii, Syringodium filiforme and five marine
algae}. International Journal of Remote Sensing, 28: 7, 1487 — 1501. \pageref{Thorhaug07}
\bibitem{Toole00}
Toole D.A.\index[people]{Toole D.A.}, Siegel D.A.\index[people]{Siegel D.A.}, Menzies D.W.\index[people]{Menzies D.W.}, Neumann M.J.\index[people]{Neumann M.J.} and Smith R.C.\index[people]{Smith R.C.} 2000. \emph{Remote-sensing reflectance determinations in the coastal ocean environment: impact of instrumental characteristics and environmental variability}. Applied Optics 39(3), 456-469. \pageref{Toole00}
\bibitem{Vahtmae06}
Vahtmae E.\index[people]{Vahtmae E.}, Kutser T.\index[people]{Kutser T.}, Martin G.\index[people]{Martin G.}, Kotta J.\index[people]{Kotta J.}, 2006. \emph{Feasibility of hyperspectral remote sensing for mapping benthic macroalgal cover in turbid coastal waters - a Baltic Sea case
study. Remote Sensing of Environment} (2006) 101:342-351. \pageref{Vahtmae06}
\bibitem{Vahtmae07}
Vahtmae E.\index[people]{Vahtmae E.} and Kutser T.\index[people]{Kutser T.} 2007. \emph{Mapping bottom type and water depth in shallow
coastal waters with satellite remote sensing}. Journal of Coastal Research, SI 50 (Proceedings of the 9th International Coastal Symposium), 185 – 189. Gold Coast, Australia, ISSN 0749.0208 \pageref{Vahtmae07}
\bibitem{Villegas06}
Villegas P. P.\index[people]{Villegas P. P.} 2006. \emph{Magnitude of herbivory in \textit{Posidonia oceanica} (L.) Delile and
factors responsible for spatial variation}. Available on web: \href{http://www.tesisenxarxa.net/}{tesisenxarxa} \pageref{Villegas06}
\bibitem{Wabnitz08}
Wabnitz C.C.\index[people]{Wabnitz C.C.}, Andr\'efou\"et S.\index[people]{Andr\'efou\"et S.}, Torres-Pulliza D.\index[people]{Torres-Pulliza D.}, M\"uller-Karger F. E.\index[people]{M\"uller-Karger F. E.} and Kramer P.A.\index[people]{Kramer P.A.}, 2008. \emph{Regional-scale seagrass habitat mapping in the Wider Caribbean region using Landsat sensors: Applications to conservation and ecology}. Remote Sensing of Environment Vol. 112, Issue 8, 15 August 2008, Pages 3455-3467 Earth Observations for Marine and Coastal Biodiversity and Ecosystems Special Issue. \pageref{Wabnitz08}
\bibitem{Walker89}
Walker, D.I.\index[people]{Walker, D.I.}, 1989. \emph{Methods for monitoring seagrass habitat}. Proceedings Rep. Workshop, 20-22 June 1988, Melbourne, Australia, VIMS Working Paper 18, pp. 1-26. \pageref{Walker89}
\bibitem{Werdell03}
Werdell P.J.\index[people]{Werdell P.J.} and Roesler C.S.\index[people]{Roesler C.S.} 2003. \emph{Remote assessment of benthic substrate composition
in shallow waters using multispectral reflectance}. Limnol. Oceanogr., 48 (1, part 2), 557–567 \pageref{Werdell03}
\bibitem{Yang10}
 Yang Ch.\index[people]{Yang Ch.}, Yang D.\index[people]{Yang D.}, Cao W.\index[people]{Cao W.}, Zhao J.\index[people]{Zhao J.}, Wang G.\index[people]{Wang G.}, Sun Zh.\index[people]{Sun Zh.}, Xu Zh.\index[people]{Xu Zh.}, Kumar M. S. R.\index[people]{Kumar M. S. R.} 2010. 
\emph{Analysis of seagrass reflectivity by using a water column correction algorithm}. International Journal of Remote Sensing. 31: 17, 4595 — 4608. \pageref{Yang10}

\end{thebibliography}

%\cleardoublepage
\newpage
\singlespacing
%\section*{List of People}
%\markright{List of People}
%\addcontentsline{toc}{section}{List of People}
%\indexsetup{level=\section,toclevel=section}
\printindex[people]
\pagebreak

\cleardoublepage
\newpage
%\phantomsection
%\addcontentsline{toc}{section}{Index}
\printindex[concepts]

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%% End Polina Lemenkova's MSc Thesis Document %%%%%%%%%%%%%%%%%%%%%%%%%%%%